---
title: "DHS Unified Dataset Pipeline — Complete Self-Contained"
subtitle: "Incremental checkpointing, geographic linking, 6-study merge, auto new-survey detection"
author: "Joseph Akuze — Wellcome Accelerator Award"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

<!--
# ============================================================================
# DHS UNIFIED DATASET PIPELINE v7.2 — COMPLETE SELF-CONTAINED
# ============================================================================
#
# PURPOSE:
#   Downloads, processes, and harmonizes DHS Birth Recode + Individual Recode
#   data for all Sub-Saharan African countries into a unified dataset matching
#   the 13-domain variable schema (outcomes, demographics, SES, etc.).
#   THEN merges DHS data with the existing 6-study unified dataset to produce
#   a single merged output for analysis and environmental pipelines.
#
# KEY FEATURES:
#   1. Runs end-to-end from a single .Rmd file
#   2. INCREMENTAL CHECKPOINTING: detects new surveys, processes ONLY those,
#      merges with existing checkpoints, re-saves
#   3. Geographic data (GE shapefiles) linked BEFORE all exports
#   4. MERGES DHS with 6-study unified dataset (NCOPS, ALERT, PRECISE,
#      PTBi, EN-INDEPTH, WHOMCS) into a single merged file
#   5. Outputs: RDS, CSV, Stata (.dta) + 10% sample — all with GPS coords
#   6. Dropbox = primary; OneDrive = optional secondary copy
#
# HOW TO RUN:
#   - First time: Run ALL chunks in order (Steps 1-4 will take several hours
#     for downloads, Step 7 takes 6-10 hours for processing).
#   - Subsequent runs: Steps 1-6 define functions; Step 7 loads checkpoints
#     in seconds and only processes NEW surveys if any.
#   - To force full reprocessing: Run the "clear-checkpoints" chunk at bottom.
#
# INCREMENTAL UPDATES:
#   When new DHS surveys become available:
#   1. Re-run Steps 3-4 (query + download new surveys — rdhs caches existing)
#   2. Re-run Step 7 — it will detect new surveys not in checkpoints,
#      process only those, merge with existing, and re-save
#   3. Re-run Steps 8-10 to rebuild final outputs + merged dataset
#
# ============================================================================
-->


# Step 1: Install and Load Packages

```{r install-packages, eval=FALSE}
# ============================================================================
# INSTALL REQUIRED PACKAGES (run once)
# ============================================================================
install.packages(c(
  "rdhs",          # DHS API client and dataset management
  "haven",         # Read Stata/SPSS/SAS files
  "tidyverse",     # Data manipulation and visualization
  "labelled",      # Handle value labels from survey data
  "survey",        # Survey-weighted analysis
  "sf",            # Spatial data for GPS coordinates
  "glue",          # String interpolation
  "janitor",       # Data cleaning
  "data.table",    # Fast data manipulation
  "openxlsx",      # Excel output
  "countrycode"    # Country code conversions
))
```

```{r load-packages, message=FALSE, warning=FALSE}
# ============================================================================
# LOAD PACKAGES
# ============================================================================
suppressPackageStartupMessages({
  library(rdhs)
  library(haven)
  library(tidyverse)
  library(labelled)
  library(glue)
  library(janitor)
  library(data.table)
  library(openxlsx)
  library(countrycode)
})

# Install sf if missing (needed for geographic linking)
if (!requireNamespace("sf", quietly = TRUE)) {
  install.packages("sf")
}

cat("All packages loaded successfully\n")
cat(glue("rdhs version: {packageVersion('rdhs')}\n"))
cat(glue("R version: {R.version.string}\n"))
cat(glue("Date: {Sys.time()}\n"))
```


# Step 1b: Directory Setup

```{r directory-setup}
# ============================================================================
# PROJECT DIRECTORY SETUP (v6.0)
# ============================================================================
# Dropbox = PRIMARY for pipeline work AND outputs
# OneDrive = OPTIONAL secondary copy
#
# Layout:
#   DROPBOX:
#     .../unified_dataset_pipeline/
#     ├── data/dhs/
#     │   ├── cache/               <- rdhs downloaded .rds files
#     │   │   └── datasets/        <- actual .rds files live here
#     │   ├── geographic/          <- GE shapefiles (.shp, .dbf, etc.)
#     │   ├── inventories/         <- survey/dataset inventory CSVs
#     │   ├── processing_logs/     <- processing logs
#     │   └── checkpoints/         <- Step 7 intermediate results (KEY!)
#     └── outputs/dhs/             <- FINAL unified outputs
#
#   ONEDRIVE (optional copy):
#     .../Datasets/master/Unified Dataset/
# ============================================================================

user <- Sys.info()[["user"]]
cat(glue("Detected user: {user}\n"))

# --- CONFIGURATION: Set your local paths below ---
# Add your username and paths, or set dropbox_base to the project root.
if (user == "josep") {
  dropbox_base <- "C:/Users/josep/Dropbox/Predictive Models for SB and NND/Minimalist/unified_dataset_pipeline"
  onedrive_data <- "C:/Users/josep/OneDrive - London School of Hygiene and Tropical Medicine/LSHTM Grants and Consultancies/Wellcome_Accelerator grant_minorities/Datasets"

} else if (user == "eidejwai") {
  dropbox_base <- "C:/Users/eidejwai/Dropbox/Predictive Models for SB and NND/Minimalist/unified_dataset_pipeline"
  onedrive_data <- "C:/Users/eidejwai/OneDrive - London School of Hygiene and Tropical Medicine/LSHTM Grants and Consultancies/Wellcome_Accelerator grant_minorities/Datasets"

} else {
  # Default: assume running from the project root
  dropbox_base <- getwd()
  onedrive_data <- NULL
  cat("NOTE: Unknown user. Using working directory as project root.\n")
  cat("      Set dropbox_base manually if paths do not resolve.\n")
}

# --- Control flag: set FALSE to skip OneDrive copy ---
copy_to_onedrive <- TRUE

dirs <- list(
  # DROPBOX: Pipeline base
  pipeline     = dropbox_base,

  # DROPBOX: DHS working directories
  dhs          = file.path(dropbox_base, "data", "dhs"),
  dhs_cache    = file.path(dropbox_base, "data", "dhs", "cache"),
  dhs_geo      = file.path(dropbox_base, "data", "dhs", "geographic"),
  dhs_inv      = file.path(dropbox_base, "data", "dhs", "inventories"),
  dhs_logs     = file.path(dropbox_base, "data", "dhs", "processing_logs"),
  dhs_checkpts = file.path(dropbox_base, "data", "dhs", "checkpoints"),

  # DROPBOX: Primary output directory
  output       = file.path(dropbox_base, "outputs", "dhs"),

  # ONEDRIVE: Study data directories (read-only source for main R notebook)
  data_root    = onedrive_data,
  master       = file.path(onedrive_data, "master"),
  ncops        = file.path(onedrive_data, "ncops"),
  alert        = file.path(onedrive_data, "alert"),
  precise      = file.path(onedrive_data, "precise"),
  ptbi         = file.path(onedrive_data, "ptbi"),
  enindepth    = file.path(onedrive_data, "en-indepth"),
  whomcs       = file.path(onedrive_data, "whomcs"),

  # ONEDRIVE: Secondary output location (optional copy)
  output_onedrive = file.path(onedrive_data, "master", "Unified Dataset")
)

# Create Dropbox directories
for (d in c(dirs$dhs, dirs$dhs_cache, dirs$dhs_geo, dirs$dhs_inv,
            dirs$dhs_logs, dirs$dhs_checkpts, dirs$output)) {
  if (!dir.exists(d)) {
    dir.create(d, recursive = TRUE)
    cat(glue("Created: {d}\n"))
  }
}

# Create OneDrive output dir only if copy is enabled
if (copy_to_onedrive && !dir.exists(dirs$output_onedrive)) {
  dir.create(dirs$output_onedrive, recursive = TRUE)
}

# Verify
cat("\n=== DIRECTORY VERIFICATION ===\n")
cat("--- Dropbox (PRIMARY) ---\n")
for (nm in c("pipeline", "dhs", "dhs_cache", "dhs_checkpts", "output")) {
  ok <- if (dir.exists(dirs[[nm]])) "OK" else "MISSING"
  cat(glue("  {nm}: [{ok}] {dirs[[nm]]}\n"))
}
cat("\n--- OneDrive (secondary copy) ---\n")
for (nm in c("data_root", "master", "output_onedrive")) {
  ok <- if (dir.exists(dirs[[nm]])) "OK" else "MISSING"
  cat(glue("  {nm}: [{ok}] {dirs[[nm]]}\n"))
}
cat(glue("\nCopy to OneDrive: {copy_to_onedrive}\n"))

# Checkpoint status
chk_files <- list.files(dirs$dhs_checkpts, pattern = "\\.rds$")
if (length(chk_files) > 0) {
  cat(glue("\n>> CHECKPOINTS FOUND ({length(chk_files)} files) - Step 7 will use cached data\n"))
  cat(glue("   Location: {dirs$dhs_checkpts}\n"))
  cat(glue("   Last modified: {file.info(file.path(dirs$dhs_checkpts, chk_files[1]))$mtime}\n"))
} else {
  cat("\n>> NO CHECKPOINTS - Step 7 will process ALL surveys (6-10 hours)\n")
}
```


# Step 2: Configure DHS API Credentials

```{r configure-rdhs, eval=FALSE}
# ============================================================================
# CONFIGURE DHS API CREDENTIALS (run ONCE)
# ============================================================================
# Register at: https://dhsprogram.com/data/new-user-registration.cfm
# ============================================================================

# Create config file if it doesn't exist
rdhs_config_file <- file.path(dirs$dhs, "rdhs.json")
if (!file.exists(rdhs_config_file)) file.create(rdhs_config_file)

set_rdhs_config(
  email = "joseph.waiswa@lshtm.ac.uk",
  project = "Predictive Modelling for stillbirths and neonatal deaths in Sub-Saharan Africa",
  config_path = rdhs_config_file,
  cache_path = dirs$dhs_cache,
  global = FALSE,
  verbose_download = TRUE
)

cat("DHS API credentials configured successfully.\n")
cat(glue("Config saved at: {rdhs_config_file}\n"))
cat(glue("Cache path: {dirs$dhs_cache}\n"))
```


# Step 3: Identify Sub-Saharan African DHS Surveys

```{r define-ssa-countries}
# ============================================================================
# DEFINE SUB-SAHARAN AFRICAN COUNTRIES (DHS country codes)
# ============================================================================

ssa_country_codes <- c(
  "AO",  # Angola
  "BJ",  # Benin
  "BF",  # Burkina Faso
  "BU",  # Burundi
  "CM",  # Cameroon
  "CV",  # Cape Verde
  "CF",  # Central African Republic
  "TD",  # Chad
  "KM",  # Comoros
  "CD",  # Congo Democratic Republic
  "CG",  # Congo (Brazzaville)
  "CI",  # Cote d'Ivoire
  "ER",  # Eritrea
  "ET",  # Ethiopia
  "GA",  # Gabon
  "GM",  # Gambia
  "GH",  # Ghana
  "GN",  # Guinea
  "GU",  # Guinea-Bissau
  "KE",  # Kenya
  "LS",  # Lesotho
  "LB",  # Liberia
  "MD",  # Madagascar
  "MW",  # Malawi
  "ML",  # Mali
  "MR",  # Mauritania
  "MZ",  # Mozambique
  "NM",  # Namibia
  "NI",  # Niger
  "NG",  # Nigeria
  "RW",  # Rwanda
  "ST",  # Sao Tome and Principe
  "SN",  # Senegal
  "SL",  # Sierra Leone
  "ZA",  # South Africa
  "SD",  # Sudan
  "SZ",  # Eswatini (Swaziland)
  "TZ",  # Tanzania
  "TG",  # Togo
  "UG",  # Uganda
  "ZM",  # Zambia
  "ZW"   # Zimbabwe
)

cat(glue("{length(ssa_country_codes)} SSA countries defined\n"))
```

```{r query-surveys, eval=FALSE}
# ============================================================================
# QUERY ALL AVAILABLE SSA DHS SURVEYS
# ============================================================================

all_surveys <- dhs_surveys()

ssa_surveys <- all_surveys %>%
  filter(
    DHS_CountryCode %in% ssa_country_codes,
    SurveyType == "DHS"
  ) %>%
  arrange(CountryName, SurveyYear)

cat(glue("\n=== SSA DHS SURVEY INVENTORY ===\n"))
cat(glue("Total SSA DHS surveys found: {nrow(ssa_surveys)}\n"))
cat(glue("Countries: {n_distinct(ssa_surveys$CountryName)}\n"))
cat(glue("Year range: {min(ssa_surveys$SurveyYear)} - {max(ssa_surveys$SurveyYear)}\n\n"))

survey_summary <- ssa_surveys %>%
  group_by(CountryName, DHS_CountryCode) %>%
  summarise(
    n_surveys = n(),
    years = paste(SurveyYear, collapse = ", "),
    .groups = "drop"
  ) %>%
  arrange(desc(n_surveys))

print(survey_summary, n = 50)

write.csv(ssa_surveys, file.path(dirs$dhs_inv, "dhs_ssa_survey_inventory.csv"), row.names = FALSE)
```

```{r query-datasets, eval=FALSE}
# ============================================================================
# IDENTIFY REQUIRED DATASET FILES
# ============================================================================

all_datasets <- dhs_datasets()

ssa_datasets <- all_datasets %>%
  filter(
    DHS_CountryCode %in% ssa_country_codes,
    SurveyType == "DHS",
    FileType %in% c(
      "Births Recode",
      "Individual Recode",
      "Household Recode",
      "Geographic Data",
      "Pregnancy Recode"
    ),
    # v7.2 FIX: Geographic Data files are shapefiles, NOT Stata .dta format.
    # The old filter (FileFormat == "Stata dataset (.dta)") excluded all GE files,
    # which is why no GPS coordinates were available for DHS records.
    # Fix: include Geographic Data regardless of FileFormat.
    (FileFormat == "Stata dataset (.dta)" | FileType == "Geographic Data")
  )

cat(glue("\n=== DATASET FILES AVAILABLE ===\n"))
cat(glue("Total dataset files: {nrow(ssa_datasets)}\n\n"))
ssa_datasets %>% count(FileType) %>% arrange(desc(n)) %>% print()

# Which surveys have both BR + IR (minimum for analysis)
survey_file_coverage <- ssa_datasets %>%
  group_by(SurveyId, CountryName, SurveyYear) %>%
  summarise(
    has_BR = any(FileType == "Births Recode"),
    has_IR = any(FileType == "Individual Recode"),
    has_GE = any(FileType == "Geographic Data"),
    has_GR = any(FileType == "Pregnancy Recode"),
    n_files = n(),
    .groups = "drop"
  ) %>%
  arrange(CountryName, SurveyYear)

complete_surveys <- survey_file_coverage %>%
  filter(has_BR & has_IR)

cat(glue("\nSurveys with both BR and IR: {nrow(complete_surveys)}\n"))
cat(glue("Surveys also with GPS data:  {sum(complete_surveys$has_GE)}\n"))
cat(glue("Surveys with Pregnancy Recode: {sum(complete_surveys$has_GR)}\n"))

write.csv(ssa_datasets, file.path(dirs$dhs_inv, "dhs_ssa_dataset_inventory.csv"), row.names = FALSE)
write.csv(complete_surveys, file.path(dirs$dhs_inv, "dhs_complete_surveys.csv"), row.names = FALSE)
```


# Step 4: Download DHS Datasets

```{r download-datasets, eval=FALSE}
# ============================================================================
# DOWNLOAD DATASETS (rdhs caches — re-running is safe & skips existing)
# ============================================================================

download_files <- ssa_datasets %>%
  filter(
    SurveyId %in% complete_surveys$SurveyId,
    FileType %in% c("Births Recode", "Individual Recode", "Geographic Data")
  )

cat(glue("Downloading {nrow(download_files)} dataset files...\n"))
cat("This may take several hours on first run. rdhs caches downloads.\n\n")

downloaded <- get_datasets(download_files$FileName)

cat(glue("\nDownload complete. {length(downloaded)} files in cache.\n"))
cat(glue("Cache location: {dirs$dhs_cache}\n"))
```


## Step 4b: Extract GE Shapefiles to Geographic Directory (v7.2 addition)

```{r extract-ge-files, eval=FALSE}
# ============================================================================
# v7.2 FIX: COPY GE SHAPEFILES FROM rdhs CACHE TO GEOGRAPHIC DIRECTORY
# ============================================================================
# rdhs downloads GE files to its cache directory, but Step 8 looks for
# shapefiles in dirs$dhs_geo. This step copies them to the right location.
# ============================================================================

cat("--- Extracting GE shapefiles to geographic directory ---\n")

# Get GE file entries from the download list
ge_download_files <- download_files %>%
  filter(FileType == "Geographic Data")

if (nrow(ge_download_files) > 0) {
  cat(glue("Found {nrow(ge_download_files)} GE datasets to process\n\n"))

  n_copied <- 0

  for (i in 1:nrow(ge_download_files)) {
    ge_filename <- ge_download_files$FileName[i]
    ge_stem <- gsub("\\.ZIP$", "", ge_filename, ignore.case = TRUE)

    # rdhs stores downloaded files in cache/datasets/ directory
    # Look for extracted shapefile components
    cache_datasets_dir <- file.path(dirs$dhs_cache, "datasets")

    # Search for shapefile components matching this GE file stem
    # GE filenames follow pattern: XXGE##FL (e.g., ETGE81FL)
    ge_pattern <- gsub("DT", "FL", ge_stem)  # Sometimes naming varies

    # Search in cache for any shapefile components
    shp_components <- list.files(
      cache_datasets_dir,
      pattern = paste0(ge_stem, "\\.(shp|shx|dbf|prj|cpg|sbn|sbx)$"),
      full.names = TRUE, ignore.case = TRUE, recursive = TRUE
    )

    # Also check if rdhs saved it as .rds (it reads shapefiles too)
    rds_path <- file.path(cache_datasets_dir, paste0(ge_stem, ".rds"))

    if (length(shp_components) > 0) {
      # Copy shapefile components to geographic directory
      for (src in shp_components) {
        dst <- file.path(dirs$dhs_geo, basename(src))
        file.copy(src, dst, overwrite = TRUE)
      }
      n_copied <- n_copied + 1
      cat(glue("  {ge_stem}: copied {length(shp_components)} files\n"))

    } else if (file.exists(rds_path)) {
      # rdhs may have read the shapefile and saved as .rds
      # Read it and write as shapefile to geographic directory
      tryCatch({
        geo_data <- readRDS(rds_path)
        if (inherits(geo_data, "sf") || inherits(geo_data, "SpatialPointsDataFrame")) {
          if (!inherits(geo_data, "sf")) geo_data <- sf::st_as_sf(geo_data)
          shp_out <- file.path(dirs$dhs_geo, paste0(ge_stem, ".shp"))
          sf::st_write(geo_data, shp_out, delete_dsn = TRUE, quiet = TRUE)
          n_copied <- n_copied + 1
          cat(glue("  {ge_stem}: converted from .rds to .shp\n"))
        }
      }, error = function(e) {
        cat(glue("  {ge_stem}: could not convert - {e$message}\n"))
      })
    } else {
      cat(glue("  {ge_stem}: not found in cache\n"))
    }
  }

  cat(glue("\nGE extraction complete: {n_copied}/{nrow(ge_download_files)} files copied\n"))

  # Verify
  shp_in_geo <- list.files(dirs$dhs_geo, pattern = "\\.shp$", recursive = TRUE)
  cat(glue("Shapefiles now in geographic directory: {length(shp_in_geo)}\n"))

} else {
  cat("No GE files in download list. Check Step 3 query results.\n")
}
```


# Step 5: Define Helper Functions

```{r helper-functions}
# ============================================================================
# CORE HELPER FUNCTIONS
# ============================================================================

# --- Safe numeric extraction ---
safe_extract <- function(df, varname, default = NA_real_) {
  if (varname %in% names(df)) {
    x <- df[[varname]]
    if (is.labelled(x)) x <- zap_labels(x)
    as.numeric(x)
  } else {
    rep(default, nrow(df))
  }
}

# --- Safe character extraction ---
safe_extract_chr <- function(df, varname, default = NA_character_) {
  if (varname %in% names(df)) {
    x <- df[[varname]]
    if (is.labelled(x)) {
      as.character(as_factor(x))
    } else {
      as.character(x)
    }
  } else {
    rep(default, nrow(df))
  }
}

# --- DHS country code to full name ---
dhs_country_name <- function(code2) {
  dhs_map <- c(
    "AO"="Angola", "BJ"="Benin", "BF"="Burkina Faso", "BU"="Burundi",
    "CM"="Cameroon", "CV"="Cape Verde", "CF"="Central African Republic",
    "TD"="Chad", "KM"="Comoros", "CD"="Congo Democratic Republic",
    "CG"="Congo", "CI"="Cote d'Ivoire", "ER"="Eritrea", "ET"="Ethiopia",
    "GA"="Gabon", "GM"="Gambia", "GH"="Ghana", "GN"="Guinea",
    "GU"="Guinea-Bissau", "KE"="Kenya", "LS"="Lesotho", "LB"="Liberia",
    "MD"="Madagascar", "MW"="Malawi", "ML"="Mali", "MR"="Mauritania",
    "MZ"="Mozambique", "NM"="Namibia", "NI"="Niger", "NG"="Nigeria",
    "RW"="Rwanda", "ST"="Sao Tome and Principe", "SN"="Senegal",
    "SL"="Sierra Leone", "ZA"="South Africa", "SD"="Sudan",
    "SZ"="Eswatini", "TZ"="Tanzania", "TG"="Togo", "UG"="Uganda",
    "ZM"="Zambia", "ZW"="Zimbabwe"
  )
  code2_clean <- substr(code2, 1, 2)
  unname(dhs_map[code2_clean])
}

# --- Find .rds file in rdhs cache from ZIP filename ---
find_rds <- function(zip_filename, cache_dir) {
  stem <- gsub("\\.ZIP$", "", zip_filename, ignore.case = TRUE)
  # rdhs stores in datasets/ subfolder
  rds_path <- file.path(cache_dir, "datasets", paste0(stem, ".rds"))
  if (file.exists(rds_path)) return(rds_path)
  # Case-insensitive fallback
  all_rds <- list.files(file.path(cache_dir, "datasets"), pattern = "\\.rds$", full.names = TRUE)
  match_idx <- which(toupper(basename(all_rds)) == toupper(paste0(stem, ".rds")))
  if (length(match_idx) > 0) return(all_rds[match_idx[1]])
  return(NULL)
}

cat("Helper functions defined\n")
```


# Step 6: Define Processing Functions

## 6a: Classify DHS Survey

```{r classify-survey}
# ============================================================================
# CLASSIFY DHS SURVEY TYPE
# ============================================================================
# Determines: module type, pregnancy history availability, stillbirth method

classify_dhs_survey <- function(ir_data, survey_id, survey_year) {
  ir_names <- names(ir_data)
  
  # Check for pregnancy history (p32 = birth outcome)
  has_preg_history <- any(grepl("^p32_", ir_names))
  preg_history_type <- "none"
  
  if (has_preg_history) {
    if (survey_year >= 2018) {
      preg_history_type <- "dhs8_standard"
    } else {
      preg_history_type <- "pre_dhs8_experimental"
    }
  }
  
  # Check for contraceptive calendar
  has_calendar <- "vcal_1" %in% ir_names
  
  # Determine module and stillbirth method
  if (has_preg_history) {
    module <- "Full pregnancy history"
    sb_method <- "pregnancy_history"
  } else if (has_calendar) {
    module <- "Birth history + calendar"
    sb_method <- "calendar"
  } else {
    module <- "Birth history only"
    sb_method <- "none"
  }
  
  cat(glue("  Survey: {survey_id} ({survey_year})\n"))
  cat(glue("    Module: {module}\n"))
  cat(glue("    Pregnancy history: {has_preg_history}"))
  if (has_preg_history) cat(glue(" ({length(grep('^p32_', ir_names))} pregnancy slots)"))
  cat("\n")
  cat(glue("    Calendar: {has_calendar}\n"))
  cat(glue("    Stillbirth method: {sb_method}\n"))
  
  return(list(
    module = module,
    has_preg_history = has_preg_history,
    preg_history_type = preg_history_type,
    has_calendar = has_calendar,
    sb_method = sb_method
  ))
}

cat("Survey classification function defined\n")
```

## 6b: Process Birth Recode File

```{r process-br-file}
# ============================================================================
# PROCESS BIRTH RECODE (BR) FILE — One record per live birth
# ============================================================================

process_br_file <- function(br_data, survey_id, survey_year, module = "Birth history + calendar") {
  
  cat(glue("  Processing BR file: {survey_id} ({nrow(br_data)} births)\n"))
  
  br_num <- br_data %>% mutate(across(where(is.labelled), zap_labels))
  
  result <- tibble(
    # --- IDENTIFIERS ---
    study_source = "DHS",
    study_design = "Population-based",
    module = module,
    study_id = paste0("DHS_", survey_id, "_", seq_len(nrow(br_num))),
    dhs_survey_id = survey_id,
    studyyear = survey_year,
    dhs_cluster = safe_extract(br_num, "v001"),
    dhs_caseid = if ("caseid" %in% names(br_num)) as.character(br_num$caseid) else NA_character_,
    dhs_weight = safe_extract(br_num, "v005") / 1000000,
    
    # --- SECTION 1: OUTCOMES ---
    out_livebirth = "Yes",
    out_stillbirth = "No",
    
    out_nnd = {
      b6_raw <- safe_extract(br_num, "b6")
      b7_raw <- safe_extract(br_num, "b7")
      case_when(
        !is.na(b6_raw) & b6_raw >= 100 & b6_raw <= 128 ~ "Yes",  # Died day 0-28
        !is.na(b7_raw) & b7_raw == 0 ~ "Yes",                     # Died in month 0
        b5_alive <- safe_extract(br_num, "b5"),
        !is.na(b5_alive) & b5_alive == 0 & !is.na(b7_raw) & b7_raw == 0 ~ "Yes",
        TRUE ~ "No"
      )
    },
    
    out_nnd_early = {
      b6_raw <- safe_extract(br_num, "b6")
      case_when(
        !is.na(b6_raw) & b6_raw >= 100 & b6_raw <= 107 ~ "Yes",
        TRUE ~ NA_character_
      )
    },
    
    out_nnd_late = {
      b6_raw <- safe_extract(br_num, "b6")
      case_when(
        !is.na(b6_raw) & b6_raw >= 108 & b6_raw <= 128 ~ "Yes",
        TRUE ~ NA_character_
      )
    },
    
    out_infant_sex = case_when(
      safe_extract(br_num, "b4") == 1 ~ "Male",
      safe_extract(br_num, "b4") == 2 ~ "Female",
      TRUE ~ NA_character_
    ),
    
    out_birthweight_g = {
      bw <- safe_extract(br_num, "m19")
      if_else(!is.na(bw) & bw >= 500 & bw < 9000, bw, NA_real_)
    },
    
    out_ga_weeks = NA_real_,  # Enriched later from IR data
    out_ga_method = NA_character_,
    
    out_dob = safe_extract(br_num, "b3"),  # CMC date
    
    out_ageatdeath = {
      b6_raw <- safe_extract(br_num, "b6")
      case_when(
        !is.na(b6_raw) & b6_raw >= 100 & b6_raw <= 128 ~ b6_raw - 100,
        TRUE ~ NA_real_
      )
    },
    
    out_multiple = case_when(
      safe_extract(br_num, "b0") == 0 ~ "Singleton",
      safe_extract(br_num, "b0") >= 1 ~ "Multiple",
      TRUE ~ NA_character_
    ),
    
    out_apgar_1min = NA_real_,
    out_apgar_5min = NA_real_,
    out_apgar_10min = NA_real_,
    out_sga = NA_character_,
    
    # --- SECTION 2: MATERNAL DEMOGRAPHICS ---
    mat_age = safe_extract(br_num, "v012"),
    
    mat_education = case_when(
      safe_extract(br_num, "v106") == 0 ~ "None",
      safe_extract(br_num, "v106") == 1 ~ "Primary",
      safe_extract(br_num, "v106") == 2 ~ "Secondary",
      safe_extract(br_num, "v106") == 3 ~ "Higher",
      TRUE ~ NA_character_
    ),
    
    mat_education_years = safe_extract(br_num, "v133"),
    
    mat_marital_status = case_when(
      safe_extract(br_num, "v501") == 0 ~ "Single",
      safe_extract(br_num, "v501") %in% c(1, 2) ~ "Married-Cohabiting",
      safe_extract(br_num, "v501") == 3 ~ "Widowed",
      safe_extract(br_num, "v501") %in% c(4, 5) ~ "Divorced-Separated",
      TRUE ~ NA_character_
    ),
    
    mat_religion = safe_extract_chr(br_num, "v130"),
    
    mat_occupation = safe_extract_chr(br_num, "v717"),
    
    mat_urban_rural = case_when(
      safe_extract(br_num, "v025") == 1 ~ "Urban",
      safe_extract(br_num, "v025") == 2 ~ "Rural",
      TRUE ~ NA_character_
    ),
    
    mat_country = sapply(safe_extract_chr(br_num, "v000"), function(x) {
      nm <- dhs_country_name(x)
      if (is.na(nm)) x else nm
    }, USE.NAMES = FALSE),
    
    mat_parity = safe_extract(br_num, "v220"),
    mat_gravidity = safe_extract(br_num, "v201"),
    
    # --- SECTION 3: HOUSEHOLD/SES ---
    hh_wealth_quintile = case_when(
      safe_extract(br_num, "v190") == 1 ~ "Poorest",
      safe_extract(br_num, "v190") == 2 ~ "Poorer",
      safe_extract(br_num, "v190") == 3 ~ "Middle",
      safe_extract(br_num, "v190") == 4 ~ "Richer",
      safe_extract(br_num, "v190") == 5 ~ "Richest",
      TRUE ~ NA_character_
    ),
    
    hh_asset_score = safe_extract(br_num, "v191"),
    
    # --- SECTION 7: ANC ---
    anc_visits = safe_extract(br_num, "m14"),
    anc_provider = safe_extract_chr(br_num, "m2a"),
    anc_timing_months = safe_extract(br_num, "m13"),
    
    # --- SECTION 8: MATERNAL HEALTH ---
    mat_height_cm = safe_extract(br_num, "v438") / 10,  # Stored as mm
    mat_weight_kg = safe_extract(br_num, "v437") / 10,  # Stored as 0.1kg
    mat_bmi = safe_extract(br_num, "v445") / 100,        # Stored as 0.01
    mat_hiv_status = safe_extract_chr(br_num, "v781"),
    mat_anaemia = safe_extract_chr(br_num, "v457"),
    
    # --- SECTION 9: DELIVERY ---
    del_mode = case_when(
      safe_extract(br_num, "m17") == 1 ~ "Caesarean",
      safe_extract(br_num, "m17") == 0 ~ "Vaginal",
      TRUE ~ NA_character_
    ),
    
    del_location = safe_extract_chr(br_num, "m15"),
    del_attendant = safe_extract_chr(br_num, "m3a"),
    
    # --- SECTION 11: NEONATAL CARE ---
    neo_breastfeeding = safe_extract_chr(br_num, "m4"),
    neo_early_bf = case_when(
      safe_extract(br_num, "m34") == 0 ~ "Immediate",
      safe_extract(br_num, "m34") <= 1 ~ "Within 1 hour",
      TRUE ~ NA_character_
    )
  )
  
  # Fix NND calculation (redo properly after tibble creation)
  b5_vals <- safe_extract(br_num, "b5")
  b6_vals <- safe_extract(br_num, "b6")
  b7_vals <- safe_extract(br_num, "b7")
  
  result$out_nnd <- case_when(
    !is.na(b6_vals) & b6_vals >= 100 & b6_vals <= 128 ~ "Yes",
    !is.na(b5_vals) & b5_vals == 0 & !is.na(b7_vals) & b7_vals == 0 ~ "Yes",
    !is.na(b5_vals) & b5_vals == 1 ~ "No",
    !is.na(b5_vals) & b5_vals == 0 & !is.na(b7_vals) & b7_vals > 0 ~ "No",
    TRUE ~ NA_character_
  )
  
  result$out_nnd_early <- case_when(
    !is.na(b6_vals) & b6_vals >= 100 & b6_vals <= 107 ~ "Yes",
    result$out_nnd == "Yes" & (is.na(b6_vals) | b6_vals > 107) ~ "No",
    result$out_nnd == "No" ~ "No",
    TRUE ~ NA_character_
  )
  
  result$out_nnd_late <- case_when(
    !is.na(b6_vals) & b6_vals >= 108 & b6_vals <= 128 ~ "Yes",
    result$out_nnd == "Yes" & (is.na(b6_vals) | b6_vals < 108) ~ "No",
    result$out_nnd == "No" ~ "No",
    TRUE ~ NA_character_
  )
  
  # Derive: stillbirth flags (all No for BR file live births)
  result$out_stillbirth_28wks <- "No"
  result$out_stillbirth_20wks <- "No"
  
  # Derive: perinatal death
  result$out_perinatal_death <- case_when(
    result$out_nnd_early == "Yes" ~ "Yes",
    result$out_nnd_early == "No" ~ "No",
    TRUE ~ NA_character_
  )
  
  cat(glue("  Extracted {nrow(result)} live births, {sum(result$out_nnd == 'Yes', na.rm=TRUE)} NNDs\n"))
  
  return(result)
}

cat("process_br_file function defined\n")
```

## 6c: Extract Stillbirths from Pregnancy History

```{r extract-stillbirths-preg-history}
# ============================================================================
# EXTRACT STILLBIRTHS FROM PREGNANCY HISTORY (DHS-8 or pre-DHS-8 with p32)
# ============================================================================

extract_stillbirths_preg_history <- function(ir_data, survey_id, survey_year, module = "Full pregnancy history") {
  
  cat(glue("  Extracting pregnancy history stillbirths: {survey_id}\n"))
  
  has_p32 <- any(grepl("^p32_", names(ir_data)))
  if (!has_p32) {
    cat("  No pregnancy history (p32) variables found.\n")
    return(NULL)
  }
  
  ir_num <- ir_data %>% mutate(across(where(is.labelled), zap_labels))
  
  # Find all pregnancy slots
  p32_vars <- sort(grep("^p32_", names(ir_num), value = TRUE))
  p20_vars <- sort(grep("^p20_", names(ir_num), value = TRUE))
  
  sb_records <- list()
  
  for (slot_idx in seq_along(p32_vars)) {
    p32_var <- p32_vars[slot_idx]
    p20_var <- if (slot_idx <= length(p20_vars)) p20_vars[slot_idx] else NULL
    
    # p32 == 2 means "born dead" (stillbirth)
    p32_vals <- ir_num[[p32_var]]
    is_stillbirth <- !is.na(p32_vals) & p32_vals == 2
    
    if (sum(is_stillbirth) == 0) next
    
    sb_subset <- ir_num[is_stillbirth, ]
    
    # Get GA from p20 (pregnancy duration)
    # Coding: 100-150 = weeks (subtract 100), 200-211 = months (subtract 200)
    ga_weeks <- NA_real_
    if (!is.null(p20_var) && p20_var %in% names(sb_subset)) {
      p20_vals <- sb_subset[[p20_var]]
      ga_weeks <- case_when(
        !is.na(p20_vals) & p20_vals >= 100 & p20_vals <= 150 ~ p20_vals - 100,
        !is.na(p20_vals) & p20_vals >= 200 & p20_vals <= 211 ~ (p20_vals - 200) * 4.33,
        TRUE ~ NA_real_
      )
    }
    
    sb_records[[length(sb_records) + 1]] <- tibble(
      study_source = "DHS",
      study_design = "Population-based",
      module = module,
      study_id = paste0("DHS_", survey_id, "_PH_SB_", slot_idx, "_", seq_len(nrow(sb_subset))),
      dhs_survey_id = survey_id,
      studyyear = survey_year,
      dhs_cluster = safe_extract(sb_subset, "v001"),
      dhs_caseid = if ("caseid" %in% names(sb_subset)) as.character(sb_subset$caseid) else NA_character_,
      dhs_weight = safe_extract(sb_subset, "v005") / 1000000,
      
      out_stillbirth = "Yes",
      out_livebirth = "No",
      out_nnd = NA_character_,
      out_nnd_early = NA_character_,
      out_nnd_late = NA_character_,
      out_ga_weeks = ga_weeks,
      out_stillbirth_28wks = if_else(!is.na(ga_weeks) & ga_weeks >= 28, "Yes",
                                     if_else(!is.na(ga_weeks), "No", NA_character_)),
      out_stillbirth_20wks = if_else(!is.na(ga_weeks) & ga_weeks >= 20, "Yes",
                                     if_else(!is.na(ga_weeks), "No", NA_character_)),
      out_infant_sex = NA_character_,
      out_birthweight_g = NA_real_,
      out_dob = NA,
      out_multiple = NA_character_,
      out_apgar_1min = NA_real_,
      out_apgar_5min = NA_real_,
      out_apgar_10min = NA_real_,
      out_ageatdeath = NA_real_,
      out_sga = NA_character_,
      out_perinatal_death = if_else(!is.na(ga_weeks) & ga_weeks >= 28, "Yes", NA_character_),
      
      mat_age = safe_extract(sb_subset, "v012"),
      mat_education = case_when(
        safe_extract(sb_subset, "v106") == 0 ~ "None",
        safe_extract(sb_subset, "v106") == 1 ~ "Primary",
        safe_extract(sb_subset, "v106") == 2 ~ "Secondary",
        safe_extract(sb_subset, "v106") == 3 ~ "Higher",
        TRUE ~ NA_character_
      ),
      mat_marital_status = case_when(
        safe_extract(sb_subset, "v501") == 0 ~ "Single",
        safe_extract(sb_subset, "v501") %in% c(1, 2) ~ "Married-Cohabiting",
        safe_extract(sb_subset, "v501") == 3 ~ "Widowed",
        safe_extract(sb_subset, "v501") %in% c(4, 5) ~ "Divorced-Separated",
        TRUE ~ NA_character_
      ),
      mat_urban_rural = case_when(
        safe_extract(sb_subset, "v025") == 1 ~ "Urban",
        safe_extract(sb_subset, "v025") == 2 ~ "Rural",
        TRUE ~ NA_character_
      ),
      mat_country = sapply(safe_extract_chr(sb_subset, "v000"), function(x) {
        nm <- dhs_country_name(x)
        if (is.na(nm)) x else nm
      }, USE.NAMES = FALSE),
      
      hh_wealth_quintile = case_when(
        safe_extract(sb_subset, "v190") == 1 ~ "Poorest",
        safe_extract(sb_subset, "v190") == 2 ~ "Poorer",
        safe_extract(sb_subset, "v190") == 3 ~ "Middle",
        safe_extract(sb_subset, "v190") == 4 ~ "Richer",
        safe_extract(sb_subset, "v190") == 5 ~ "Richest",
        TRUE ~ NA_character_
      ),
      hh_asset_score = safe_extract(sb_subset, "v191")
    )
  }
  
  if (length(sb_records) > 0) {
    result <- bind_rows(sb_records)
    cat(glue("  Pregnancy history stillbirths extracted: {nrow(result)}\n"))
    return(result)
  } else {
    cat("  No pregnancy history stillbirths found\n")
    return(NULL)
  }
}

cat("extract_stillbirths_preg_history function defined\n")
```

## 6d: Extract Stillbirths from Calendar

```{r extract-stillbirths-calendar}
# ============================================================================
# EXTRACT STILLBIRTHS FROM CONTRACEPTIVE CALENDAR (DHS-7 and earlier)
# ============================================================================
# vcal_1: Character string of monthly codes, most recent month first
# "B" = birth, "T" = termination, "P" = pregnant
# Stillbirth = "T" preceded by 7+ months of "P"
# ============================================================================

extract_stillbirths_calendar <- function(ir_data, survey_id, survey_year) {
  
  cat(glue("  Extracting calendar stillbirths: {survey_id}\n"))
  
  if (!"vcal_1" %in% names(ir_data)) {
    cat("  No contraceptive calendar (vcal_1) found.\n")
    return(NULL)
  }
  
  ir_num <- ir_data %>% mutate(across(where(is.labelled), zap_labels))
  
  sb_records <- list()
  
  for (i in seq_len(nrow(ir_num))) {
    cal <- as.character(ir_num$vcal_1[i])
    if (is.na(cal) || nchar(cal) < 12) next
    
    t_positions <- gregexpr("T", cal)[[1]]
    if (t_positions[1] == -1) next
    
    for (t_pos in t_positions) {
      preg_months <- 0
      check_pos <- t_pos + 1
      while (check_pos <= nchar(cal) && substr(cal, check_pos, check_pos) == "P") {
        preg_months <- preg_months + 1
        check_pos <- check_pos + 1
      }
      
      # Stillbirth = termination after 6+ months of P (7+ months total pregnancy)
      if (preg_months >= 6) {
        sb_records[[length(sb_records) + 1]] <- tibble(
          study_source = "DHS",
          study_design = "Population-based",
          module = "Birth history + calendar",
          study_id = paste0("DHS_", survey_id, "_CAL_SB_", length(sb_records) + 1),
          dhs_survey_id = survey_id,
          studyyear = survey_year,
          dhs_cluster = safe_extract(ir_num[i, ], "v001"),
          dhs_caseid = if ("caseid" %in% names(ir_num)) as.character(ir_num$caseid[i]) else NA_character_,
          dhs_weight = safe_extract(ir_num[i, ], "v005") / 1000000,
          
          out_stillbirth = "Yes",
          out_livebirth = "No",
          out_nnd = NA_character_,
          out_nnd_early = NA_character_,
          out_nnd_late = NA_character_,
          out_ga_weeks = (preg_months + 1) * 4.33,
          out_stillbirth_28wks = if_else(preg_months >= 6, "Yes", "No"),
          out_stillbirth_20wks = "Yes",
          out_infant_sex = NA_character_,
          out_birthweight_g = NA_real_,
          out_dob = NA,
          out_multiple = NA_character_,
          out_apgar_1min = NA_real_,
          out_apgar_5min = NA_real_,
          out_apgar_10min = NA_real_,
          out_ageatdeath = NA_real_,
          out_sga = NA_character_,
          out_perinatal_death = if_else(preg_months >= 6, "Yes", NA_character_),
          
          mat_age = safe_extract(ir_num[i, ], "v012"),
          mat_education = case_when(
            safe_extract(ir_num[i, ], "v106") == 0 ~ "None",
            safe_extract(ir_num[i, ], "v106") == 1 ~ "Primary",
            safe_extract(ir_num[i, ], "v106") == 2 ~ "Secondary",
            safe_extract(ir_num[i, ], "v106") == 3 ~ "Higher",
            TRUE ~ NA_character_
          ),
          mat_marital_status = case_when(
            safe_extract(ir_num[i, ], "v501") == 0 ~ "Single",
            safe_extract(ir_num[i, ], "v501") %in% c(1, 2) ~ "Married-Cohabiting",
            safe_extract(ir_num[i, ], "v501") == 3 ~ "Widowed",
            safe_extract(ir_num[i, ], "v501") %in% c(4, 5) ~ "Divorced-Separated",
            TRUE ~ NA_character_
          ),
          mat_urban_rural = case_when(
            safe_extract(ir_num[i, ], "v025") == 1 ~ "Urban",
            safe_extract(ir_num[i, ], "v025") == 2 ~ "Rural",
            TRUE ~ NA_character_
          ),
          mat_country = {
            v000 <- safe_extract_chr(ir_num[i, ], "v000")
            nm <- dhs_country_name(v000)
            if (is.na(nm)) v000 else nm
          },
          
          hh_wealth_quintile = case_when(
            safe_extract(ir_num[i, ], "v190") == 1 ~ "Poorest",
            safe_extract(ir_num[i, ], "v190") == 2 ~ "Poorer",
            safe_extract(ir_num[i, ], "v190") == 3 ~ "Middle",
            safe_extract(ir_num[i, ], "v190") == 4 ~ "Richer",
            safe_extract(ir_num[i, ], "v190") == 5 ~ "Richest",
            TRUE ~ NA_character_
          ),
          hh_asset_score = safe_extract(ir_num[i, ], "v191")
        )
      }
    }
  }
  
  if (length(sb_records) > 0) {
    result <- bind_rows(sb_records)
    cat(glue("  Calendar stillbirths extracted: {nrow(result)}\n"))
    return(result)
  } else {
    cat("  No calendar stillbirths found\n")
    return(NULL)
  }
}

cat("extract_stillbirths_calendar function defined\n")
```

## 6e: Enrich BR with GA from IR

```{r enrich-ga}
# ============================================================================
# ENRICH LIVE BIRTHS WITH GA FROM PREGNANCY HISTORY
# ============================================================================

enrich_br_ga_from_preg_history <- function(br_processed, ir_data) {
  cat("  Enriching BR with GA from pregnancy history...\n")
  
  ir_num <- ir_data %>% mutate(across(where(is.labelled), zap_labels))
  p20_vars <- sort(grep("^p20_", names(ir_num), value = TRUE))
  p32_vars <- sort(grep("^p32_", names(ir_num), value = TRUE))
  
  if (length(p20_vars) == 0 || length(p32_vars) == 0) {
    cat("    No p20/p32 variables found\n")
    return(br_processed)
  }
  
  # For each live birth in p32 (p32==1), extract p20 as GA
  # This is approximate — matching by caseid and birth order
  ga_enriched <- 0
  # (Placeholder: full matching logic would go here)
  
  cat(glue("    GA enrichment from preg history: {ga_enriched} records\n"))
  return(br_processed)
}

enrich_br_ga_from_calendar <- function(br_processed, ir_data) {
  cat("  Enriching BR with GA from calendar (limited)...\n")
  # Calendar doesn't directly provide GA for live births
  return(br_processed)
}

cat("GA enrichment functions defined\n")
```


# Step 7: Batch Process with Incremental Checkpointing

```{r batch-process}
# ============================================================================
# BATCH PROCESS ALL DOWNLOADED SSA SURVEYS — INCREMENTAL CHECKPOINTING
# ============================================================================
#
# HOW IT WORKS:
#   1. Loads checkpoints if they exist (named lists, one element per survey)
#   2. Compares checkpoint survey IDs against the current complete_surveys list
#   3. If ALL surveys are in checkpoints → loads in seconds (FAST PATH)
#   4. If NEW surveys detected → processes ONLY the new ones (INCREMENTAL)
#   5. If NO checkpoints → processes everything (FULL PATH, 6-10 hours)
#   6. Saves/updates checkpoints after processing
#
# TO FORCE FULL REPROCESSING:
#   Run the "clear-checkpoints" chunk at the bottom of this file.
# ============================================================================

# --- Define checkpoint file paths ---
checkpoint_br  <- file.path(dirs$dhs_checkpts, "all_br_processed.rds")
checkpoint_sb  <- file.path(dirs$dhs_checkpts, "all_stillbirths.rds")
checkpoint_log <- file.path(dirs$dhs_checkpts, "processing_log.rds")

# --- Reload inventories from disk if not in memory ---
if (!exists("complete_surveys")) {
  inv_path <- file.path(dirs$dhs_inv, "dhs_complete_surveys.csv")
  if (file.exists(inv_path)) {
    complete_surveys <- read.csv(inv_path, stringsAsFactors = FALSE)
    cat(glue("Reloaded complete_surveys from disk: {nrow(complete_surveys)} surveys\n"))
  } else {
    stop("No complete_surveys found. Run Steps 3-4 first (query + download).")
  }
}

if (!exists("ssa_datasets")) {
  inv_path2 <- file.path(dirs$dhs_inv, "dhs_ssa_dataset_inventory.csv")
  if (file.exists(inv_path2)) {
    ssa_datasets <- read.csv(inv_path2, stringsAsFactors = FALSE)
    cat(glue("Reloaded ssa_datasets from disk: {nrow(ssa_datasets)} datasets\n"))
  } else {
    stop("No ssa_datasets inventory found. Run Step 3 first.")
  }
}

survey_list <- complete_surveys

# --- Check for existing checkpoints ---
checkpoints_exist <- all(
  file.exists(checkpoint_br),
  file.exists(checkpoint_sb),
  file.exists(checkpoint_log)
)

if (checkpoints_exist) {
  # Load existing checkpoints
  cat(strrep("=", 60), "\n")
  cat("LOADING EXISTING CHECKPOINTS\n")
  cat(strrep("=", 60), "\n")
  cat(glue("  Checkpoint date: {file.info(checkpoint_br)$mtime}\n\n"))
  
  all_br_processed <- readRDS(checkpoint_br)
  all_stillbirths  <- readRDS(checkpoint_sb)
  processing_log   <- readRDS(checkpoint_log)
  
  cached_surveys <- names(all_br_processed)
  target_surveys <- survey_list$SurveyId
  new_surveys    <- setdiff(target_surveys, cached_surveys)
  
  cat(glue("  Cached surveys: {length(cached_surveys)}\n"))
  cat(glue("  Target surveys: {length(target_surveys)}\n"))
  cat(glue("  NEW surveys to process: {length(new_surveys)}\n\n"))
  
  if (length(new_surveys) == 0) {
    # =======================================================================
    # FAST PATH: All surveys already cached — load and combine
    # =======================================================================
    cat("ALL SURVEYS CACHED — skipping processing entirely\n\n")
    
  } else {
    # =======================================================================
    # INCREMENTAL PATH: Process only new surveys, merge with existing
    # =======================================================================
    cat(strrep("=", 60), "\n")
    cat(glue("INCREMENTAL UPDATE: Processing {length(new_surveys)} new surveys\n"))
    cat(strrep("=", 60), "\n\n")
    
    new_survey_list <- survey_list %>% filter(SurveyId %in% new_surveys)
    
    n_rds <- length(list.files(file.path(dirs$dhs_cache, "datasets"), pattern = "\\.rds$"))
    cat(glue("Found {n_rds} .rds files in cache\n\n"))
    
    for (i in seq_len(nrow(new_survey_list))) {
      sid     <- new_survey_list$SurveyId[i]
      syear   <- new_survey_list$SurveyYear[i]
      country <- new_survey_list$CountryName[i]
      
      cat(glue("\n{strrep('=', 60)}\n"))
      cat(glue("[NEW] Processing: {country} {syear} ({sid}) [{i}/{nrow(new_survey_list)}]\n"))
      cat(glue("{strrep('=', 60)}\n"))
      
      tryCatch({
        # Load IR
        ir_filename <- ssa_datasets %>%
          filter(SurveyId == sid, FileType == "Individual Recode") %>%
          pull(FileName)
        
        ir_data <- NULL
        survey_class <- list(module = "Birth history only", sb_method = "none")
        
        if (length(ir_filename) > 0) {
          ir_path <- find_rds(ir_filename[1], dirs$dhs_cache)
          if (!is.null(ir_path)) {
            cat(glue("  Loading IR: {basename(ir_path)}\n"))
            ir_data <- readRDS(ir_path)
            survey_class <- classify_dhs_survey(ir_data, sid, syear)
          }
        }
        
        # Load and process BR
        br_filename <- ssa_datasets %>%
          filter(SurveyId == sid, FileType == "Births Recode") %>%
          pull(FileName)
        
        if (length(br_filename) > 0) {
          br_path <- find_rds(br_filename[1], dirs$dhs_cache)
          if (!is.null(br_path)) {
            cat(glue("  Loading BR: {basename(br_path)}\n"))
            br_data <- readRDS(br_path)
            br_processed <- process_br_file(br_data, sid, syear, module = survey_class$module)
            
            if (!is.null(ir_data)) {
              if (survey_class$sb_method == "pregnancy_history") {
                br_processed <- enrich_br_ga_from_preg_history(br_processed, ir_data)
              } else if (survey_class$sb_method == "calendar") {
                br_processed <- enrich_br_ga_from_calendar(br_processed, ir_data)
              }
            }
            
            all_br_processed[[sid]] <- br_processed
          }
        }
        
        # Extract stillbirths
        if (!is.null(ir_data)) {
          if (survey_class$sb_method == "pregnancy_history") {
            sb <- extract_stillbirths_preg_history(ir_data, sid, syear, module = survey_class$module)
            if (!is.null(sb)) all_stillbirths[[sid]] <- sb
          } else if (survey_class$sb_method == "calendar") {
            sb_cal <- extract_stillbirths_calendar(ir_data, sid, syear)
            if (!is.null(sb_cal)) all_stillbirths[[sid]] <- sb_cal
          }
        }
        
        processing_log[[sid]] <- list(
          survey_id = sid, country = country, year = syear,
          status = "success", module = survey_class$module,
          sb_method = survey_class$sb_method,
          n_births = if (!is.null(all_br_processed[[sid]])) nrow(all_br_processed[[sid]]) else 0,
          n_stillbirths = if (!is.null(all_stillbirths[[sid]])) nrow(all_stillbirths[[sid]]) else 0
        )
        
      }, error = function(e) {
        cat(glue("  ERROR: {e$message}\n"))
        processing_log[[sid]] <<- list(
          survey_id = sid, country = country, year = syear,
          status = "error", error_msg = e$message,
          module = NA_character_, sb_method = NA_character_
        )
      })
      
      if (exists("ir_data")) rm(ir_data)
      if (exists("br_data")) rm(br_data)
      gc(verbose = FALSE)
    }
    
    # Save UPDATED checkpoints
    cat(glue("\n\nSaving updated checkpoints ({length(all_br_processed)} surveys)...\n"))
    saveRDS(all_br_processed, checkpoint_br)
    saveRDS(all_stillbirths, checkpoint_sb)
    saveRDS(processing_log, checkpoint_log)
    cat("Updated checkpoints saved.\n")
  }
  
} else {
  # =========================================================================
  # FULL PATH: No checkpoints — process ALL surveys (6-10 hours)
  # =========================================================================
  cat(strrep("=", 60), "\n")
  cat("NO CHECKPOINTS FOUND — PROCESSING ALL SURVEYS\n")
  cat("This will take several hours. Results will be checkpointed.\n")
  cat(strrep("=", 60), "\n\n")
  
  all_br_processed <- list()
  all_stillbirths  <- list()
  processing_log   <- list()
  
  n_rds <- length(list.files(file.path(dirs$dhs_cache, "datasets"), pattern = "\\.rds$"))
  cat(glue("Found {n_rds} .rds files in cache/datasets/\n"))
  cat(glue("Processing {nrow(survey_list)} surveys\n\n"))
  
  for (i in seq_len(nrow(survey_list))) {
    sid     <- survey_list$SurveyId[i]
    syear   <- survey_list$SurveyYear[i]
    country <- survey_list$CountryName[i]
    
    cat(glue("\n{strrep('=', 60)}\n"))
    cat(glue("Processing: {country} {syear} ({sid}) [{i}/{nrow(survey_list)}]\n"))
    cat(glue("{strrep('=', 60)}\n"))
    
    tryCatch({
      # Load IR
      ir_filename <- ssa_datasets %>%
        filter(SurveyId == sid, FileType == "Individual Recode") %>%
        pull(FileName)
      
      ir_data <- NULL
      survey_class <- list(module = "Birth history only", sb_method = "none")
      
      if (length(ir_filename) > 0) {
        ir_path <- find_rds(ir_filename[1], dirs$dhs_cache)
        if (!is.null(ir_path)) {
          cat(glue("  Loading IR: {basename(ir_path)}\n"))
          ir_data <- readRDS(ir_path)
          survey_class <- classify_dhs_survey(ir_data, sid, syear)
        }
      }
      
      # Load and process BR
      br_filename <- ssa_datasets %>%
        filter(SurveyId == sid, FileType == "Births Recode") %>%
        pull(FileName)
      
      if (length(br_filename) > 0) {
        br_path <- find_rds(br_filename[1], dirs$dhs_cache)
        if (!is.null(br_path)) {
          cat(glue("  Loading BR: {basename(br_path)}\n"))
          br_data <- readRDS(br_path)
          br_processed <- process_br_file(br_data, sid, syear, module = survey_class$module)
          
          if (!is.null(ir_data)) {
            if (survey_class$sb_method == "pregnancy_history") {
              br_processed <- enrich_br_ga_from_preg_history(br_processed, ir_data)
            } else if (survey_class$sb_method == "calendar") {
              br_processed <- enrich_br_ga_from_calendar(br_processed, ir_data)
            }
          }
          
          all_br_processed[[sid]] <- br_processed
        }
      }
      
      # Extract stillbirths
      if (!is.null(ir_data)) {
        if (survey_class$sb_method == "pregnancy_history") {
          sb <- extract_stillbirths_preg_history(ir_data, sid, syear, module = survey_class$module)
          if (!is.null(sb)) all_stillbirths[[sid]] <- sb
        } else if (survey_class$sb_method == "calendar") {
          sb_cal <- extract_stillbirths_calendar(ir_data, sid, syear)
          if (!is.null(sb_cal)) all_stillbirths[[sid]] <- sb_cal
        }
      }
      
      processing_log[[sid]] <- list(
        survey_id = sid, country = country, year = syear,
        status = "success", module = survey_class$module,
        sb_method = survey_class$sb_method,
        n_births = if (!is.null(all_br_processed[[sid]])) nrow(all_br_processed[[sid]]) else 0,
        n_stillbirths = if (!is.null(all_stillbirths[[sid]])) nrow(all_stillbirths[[sid]]) else 0
      )
      
    }, error = function(e) {
      cat(glue("  ERROR: {e$message}\n"))
      processing_log[[sid]] <<- list(
        survey_id = sid, country = country, year = syear,
        status = "error", error_msg = e$message,
        module = NA_character_, sb_method = NA_character_
      )
    })
    
    if (exists("ir_data")) rm(ir_data)
    if (exists("br_data")) rm(br_data)
    gc(verbose = FALSE)
  }
  
  # Save checkpoints
  cat(glue("\n\nSaving checkpoints to: {dirs$dhs_checkpts}\n"))
  saveRDS(all_br_processed, checkpoint_br)
  saveRDS(all_stillbirths, checkpoint_sb)
  saveRDS(processing_log, checkpoint_log)
  cat("Checkpoints saved. Next run will load these instead of reprocessing.\n")
}

# --- COMBINE (runs in all paths) ---
dhs_livebirths  <- bind_rows(all_br_processed)
dhs_stillbirths <- bind_rows(all_stillbirths)
log_df          <- bind_rows(processing_log)

# --- SUMMARY ---
cat(glue("\n\n{strrep('=', 60)}\n"))
cat("PROCESSING SUMMARY\n")
cat(glue("{strrep('=', 60)}\n"))
cat(glue("Total live births: {nrow(dhs_livebirths)}\n"))
cat(glue("Total stillbirths: {nrow(dhs_stillbirths)}\n"))
cat(glue("Surveys successful: {sum(log_df$status == 'success')}\n"))
cat(glue("Surveys with errors: {sum(log_df$status == 'error')}\n\n"))

cat("Module classification:\n")
log_df %>% filter(status == "success") %>% count(module) %>% print()
```


# Steps 8-9: Combine, Link Geographic Data, and Save

```{r combine-link-geo-save}
# ============================================================================
# COMBINE LIVE BIRTHS + STILLBIRTHS, LINK GEOGRAPHIC DATA, THEN SAVE
# ============================================================================

# --- 1. COMBINE ---
cat(strrep("=", 60), "\n")
cat("COMBINING AND SAVING DHS UNIFIED DATASET\n")
cat(strrep("=", 60), "\n\n")

# Align columns
missing_in_sb <- setdiff(names(dhs_livebirths), names(dhs_stillbirths))
missing_in_lb <- setdiff(names(dhs_stillbirths), names(dhs_livebirths))
for (col in missing_in_sb) dhs_stillbirths[[col]] <- NA
for (col in missing_in_lb) dhs_livebirths[[col]] <- NA

dhs_unified <- bind_rows(dhs_livebirths, dhs_stillbirths)

# Generate unified IDs and perinatal death
dhs_unified <- dhs_unified %>%
  mutate(
    unified_id = paste0("DHS_", row_number()),
    out_perinatal_death = case_when(
      out_stillbirth_28wks == "Yes" | out_nnd_early == "Yes" ~ "Yes",
      out_stillbirth_28wks == "No" & (out_nnd_early == "No" | is.na(out_nnd_early)) ~ "No",
      TRUE ~ NA_character_
    )
  )

cat(glue("Combined dataset: {nrow(dhs_unified)} records\n"))
cat(glue("  Live births: {sum(dhs_unified$out_livebirth == 'Yes', na.rm=TRUE)}\n"))
cat(glue("  Stillbirths: {sum(dhs_unified$out_stillbirth == 'Yes', na.rm=TRUE)}\n"))
cat(glue("  Neonatal deaths: {sum(dhs_unified$out_nnd == 'Yes', na.rm=TRUE)}\n\n"))


# --- 2. LINK GEOGRAPHIC DATA ---
cat("--- Linking geographic data ---\n")

ge_files <- list.files(dirs$dhs_geo, pattern = "\\.(shp|dbf)$",
                       full.names = TRUE, recursive = TRUE)
ge_shp_files <- ge_files[grepl("\\.shp$", ge_files)]

if (length(ge_shp_files) > 0) {
  cat(glue("Found {length(ge_shp_files)} GE shapefiles\n"))
  
  if (!requireNamespace("sf", quietly = TRUE)) install.packages("sf")
  
  all_geo <- list()
  
  for (ge_path in ge_shp_files) {
    tryCatch({
      geo <- sf::st_read(ge_path, quiet = TRUE)
      
      geo_df <- geo %>%
        sf::st_drop_geometry() %>%
        select(any_of(c("DHSCLUST", "LATNUM", "LONGNUM", "URBAN_RURA",
                         "DHSID", "DHSREGCO", "DHSREGNA", "ALT_GPS", "ALT_DEM"))) %>%
        rename_with(~case_when(
          . == "DHSCLUST"    ~ "dhs_cluster",
          . == "LATNUM"      ~ "env_latitude",
          . == "LONGNUM"     ~ "env_longitude",
          . == "URBAN_RURA"  ~ "geo_urban_rural",
          . == "ALT_DEM"     ~ "env_altitude_m",
          TRUE ~ .
        )) %>%
        mutate(
          ge_source_file = basename(ge_path),
          ge_country_code = substr(basename(ge_path), 1, 2)
        )
      
      if ("env_latitude" %in% names(geo_df)) {
        geo_df <- geo_df %>%
          filter(!is.na(env_latitude) & env_latitude != 0)
      }
      
      all_geo[[length(all_geo) + 1]] <- geo_df
    }, error = function(e) {
      cat(glue("  Warning: Could not read {basename(ge_path)}: {e$message}\n"))
    })
  }
  
  if (length(all_geo) > 0) {
    geo_combined <- bind_rows(all_geo) %>%
      distinct(dhs_cluster, ge_country_code, .keep_all = TRUE)
    
    dhs_unified <- dhs_unified %>%
      mutate(.ge_country_code = substr(dhs_survey_id, 1, 2)) %>%
      left_join(
        geo_combined,
        by = c("dhs_cluster" = "dhs_cluster", ".ge_country_code" = "ge_country_code")
      ) %>%
      select(-.ge_country_code)
    
    n_geo <- sum(!is.na(dhs_unified$env_latitude))
    cat(glue("Geographic data linked: {n_geo}/{nrow(dhs_unified)} records ({round(n_geo/nrow(dhs_unified)*100,1)}%)\n\n"))
  } else {
    cat("No valid geographic data loaded.\n")
    if (!"env_latitude" %in% names(dhs_unified)) dhs_unified$env_latitude <- NA_real_
    if (!"env_longitude" %in% names(dhs_unified)) dhs_unified$env_longitude <- NA_real_
  }

} else {
  cat("No GE shapefiles found in geographic directory.\n")
  cat(glue("  Looked in: {dirs$dhs_geo}\n"))
  cat("  Download GE files and place them in this directory.\n\n")
  if (!"env_latitude" %in% names(dhs_unified)) dhs_unified$env_latitude <- NA_real_
  if (!"env_longitude" %in% names(dhs_unified)) dhs_unified$env_longitude <- NA_real_
}

# --- v7.2 FIX: Also store coordinates as loc_latitude/loc_longitude ---
# The environmental pipeline (v3.2) reads loc_latitude/loc_longitude, not
# env_latitude/env_longitude. Copy DHS GPS coordinates to both column names
# so coordinates are available regardless of which name is used downstream.
if ("env_latitude" %in% names(dhs_unified)) {
  dhs_unified$loc_latitude <- dhs_unified$env_latitude
  dhs_unified$loc_longitude <- dhs_unified$env_longitude
  n_with_loc <- sum(!is.na(dhs_unified$loc_latitude))
  cat(glue("v7.2: Copied env_latitude/longitude to loc_latitude/longitude ({n_with_loc} records)\n\n"))
} else {
  dhs_unified$loc_latitude <- NA_real_
  dhs_unified$loc_longitude <- NA_real_
}


# --- 3. SAVE TO DROPBOX (PRIMARY) ---
cat("--- Saving to Dropbox (primary) ---\n")

if (!dir.exists(dirs$output)) dir.create(dirs$output, recursive = TRUE)

saveRDS(dhs_unified, file.path(dirs$output, "dhs_unified_dataset.rds"))
cat(glue("  Saved: dhs_unified_dataset.rds\n"))

write_csv(dhs_unified, file.path(dirs$output, "dhs_unified_dataset.csv"))
cat(glue("  Saved: dhs_unified_dataset.csv\n"))

# Stata (truncate long variable names)
dhs_stata <- dhs_unified
long_names <- names(dhs_stata)[nchar(names(dhs_stata)) > 32]
if (length(long_names) > 0) {
  cat(glue("  Truncating {length(long_names)} variable names for Stata compatibility\n"))
  names(dhs_stata)[nchar(names(dhs_stata)) > 32] <- substr(
    names(dhs_stata)[nchar(names(dhs_stata)) > 32], 1, 32
  )
}
haven::write_dta(dhs_stata, file.path(dirs$output, "dhs_unified_dataset.dta"), version = 14)
cat(glue("  Saved: dhs_unified_dataset.dta\n"))
rm(dhs_stata)

# 10% sample (stratified by country, seed for reproducibility)
set.seed(42)
dhs_sample <- dhs_unified %>%
  group_by(mat_country) %>%
  sample_frac(0.1) %>%
  ungroup()

saveRDS(dhs_sample, file.path(dirs$output, "dhs_unified_10pct_sample.rds"))
write_csv(dhs_sample, file.path(dirs$output, "dhs_unified_10pct_sample.csv"))
cat(glue("  Saved 10% sample: {nrow(dhs_sample)} records\n"))

# Processing log
write_csv(log_df, file.path(dirs$dhs_logs, "dhs_processing_log.csv"))
cat(glue("  Saved log: dhs_processing_log.csv\n"))


# --- 4. OPTIONAL: COPY TO ONEDRIVE ---
if (copy_to_onedrive && dir.exists(dirs$output_onedrive)) {
  cat("\n--- Copying to OneDrive (secondary) ---\n")
  
  files_to_copy <- c(
    "dhs_unified_dataset.rds",
    "dhs_unified_dataset.csv",
    "dhs_unified_dataset.dta",
    "dhs_unified_10pct_sample.rds",
    "dhs_unified_10pct_sample.csv"
  )
  
  for (f in files_to_copy) {
    src <- file.path(dirs$output, f)
    dst <- file.path(dirs$output_onedrive, f)
    if (file.exists(src)) {
      file.copy(src, dst, overwrite = TRUE)
      cat(glue("  Copied: {f}\n"))
    }
  }
  
  cat(glue("  OneDrive copy complete: {dirs$output_onedrive}\n"))
} else if (copy_to_onedrive) {
  cat("\nOneDrive directory not found - skipping copy.\n")
}


# --- 5. FINAL SUMMARY ---
cat(glue("\n\n{strrep('=', 60)}\n"))
cat("DHS UNIFIED DATASET - FINAL SUMMARY\n")
cat(glue("{strrep('=', 60)}\n\n"))

cat(glue("Total records: {nrow(dhs_unified)}\n"))
cat(glue("  Live births:     {sum(dhs_unified$out_livebirth == 'Yes', na.rm=TRUE)}\n"))
cat(glue("  Stillbirths:     {sum(dhs_unified$out_stillbirth == 'Yes', na.rm=TRUE)}\n"))
cat(glue("  Neonatal deaths: {sum(dhs_unified$out_nnd == 'Yes', na.rm=TRUE)}\n"))
cat(glue("  Countries:       {n_distinct(dhs_unified$mat_country)}\n"))
cat(glue("  Surveys:         {n_distinct(dhs_unified$dhs_survey_id)}\n"))
cat(glue("  With GPS coords: {sum(!is.na(dhs_unified$env_latitude))}\n\n"))

cat("Module distribution:\n")
dhs_unified %>% count(module) %>% print()

cat(glue("\nPrimary output: {dirs$output}\n"))
if (copy_to_onedrive) cat(glue("OneDrive copy:  {dirs$output_onedrive}\n"))

cat("\n\nPipeline ready for Step 10: Merge with 6-study unified dataset.\n")
```


# Step 10: Merge DHS with 6-Study Unified Dataset

```{r merge-with-6study}
# ============================================================================
# STEP 10: MERGE DHS WITH 6-STUDY UNIFIED DATASET
# ============================================================================
# Loads the existing 6-study unified dataset (NCOPS, ALERT, PRECISE,
# PTBi, EN-INDEPTH, WHOMCS) and appends DHS data to create a single
# merged output file for analysis and environmental pipelines.
# ============================================================================

MERGED_VERSION <- "10.16"
```


```{r merge-with-6study}
```


```{r merge-with-6study}
```


```{r merge-with-6study}
# Load DHS unified dataset from saved RDS (if not already in memory)
if (!exists("dhs_unified")) {
  dhs_rds_candidates <- c(
    file.path(gsub("/data$", "", dirs$pipeline), "output", "dhs_unified_dataset.rds"),
    file.path(dirs$pipeline, "output", "dhs_unified_dataset.rds"),
    file.path(gsub("/data$", "", dirs$pipeline), "output", "intermediate", "dhs_unified_dataset.rds"),
    file.path(dirs$output_onedrive, "dhs_unified_dataset.rds")
  )
  dhs_rds_path <- NULL
  for (p in dhs_rds_candidates) {
    if (file.exists(p)) { dhs_rds_path <- p; break }
  }
  if (is.null(dhs_rds_path)) stop("Cannot find dhs_unified_dataset.rds - re-run Steps 1-9 first")
  cat(glue("Loading DHS dataset from: {dhs_rds_path}\n\n"))
  dhs_unified <- readRDS(dhs_rds_path)
}

cat(strrep("=", 60), "\n")
cat("STEP 10: MERGE DHS WITH 6-STUDY UNIFIED DATASET\n")
cat(strrep("=", 60), "\n\n")

# --- 1. LOCATE THE 6-STUDY UNIFIED DATASET ---
base_pipeline <- gsub("/data$", "", dirs$pipeline)

versions_to_try <- c(MERGED_VERSION, "10.15", "10.14", "10.13", "10.12",
                     "10.11", "10.10", "10.9", "10.8")
versions_to_try <- unique(versions_to_try)

six_study_candidates <- c(
  file.path(base_pipeline, "output", paste0("unified_dataset_v", versions_to_try, ".rds")),
  file.path(dirs$pipeline, "output", paste0("unified_dataset_v", versions_to_try, ".rds")),
  file.path(base_pipeline, "output", "intermediate", "phase3_unified.rds"),
  file.path(dirs$pipeline, "intermediate", "phase3_unified.rds"),
  file.path(dirs$output_onedrive, paste0("unified_dataset_v", versions_to_try[1:3], ".rds"))
)

cat("Searching for 6-study unified dataset...\n")
cat(glue("  base_pipeline:     {base_pipeline}\n"))
cat(glue("  dirs$pipeline:     {dirs$pipeline}\n"))
cat(glue("  dirs$output_onedrive: {dirs$output_onedrive}\n\n"))

six_study_path <- NULL
for (p in six_study_candidates) {
  if (file.exists(p)) { six_study_path <- p; break }
}

# Recursive fallback
if (is.null(six_study_path)) {
  cat("Standard paths failed. Trying recursive search...\n")
  recursive_hits <- list.files(
    path = base_pipeline,
    pattern = "^unified_dataset_v10.*\\.rds$",
    recursive = TRUE, full.names = TRUE
  )
  if (length(recursive_hits) > 0) {
    six_study_path <- sort(recursive_hits, decreasing = TRUE)[1]
    cat(glue("  Found via recursive search: {six_study_path}\n"))
  }
}

if (is.null(six_study_path)) {
  cat("WARNING: No 6-study unified dataset found!\n")
  cat("Searched locations:\n")
  for (p in unique(six_study_candidates[1:6])) cat(glue("  {p}\n"))
  cat("\nSkipping merge - DHS standalone outputs are still available.\n")
  
} else {
  
  cat(glue("Found 6-study dataset: {basename(six_study_path)}\n"))
  cat(glue("  Path: {six_study_path}\n"))
  cat(glue("  Size: {round(file.info(six_study_path)$size / 1024^2, 1)} MB\n\n"))
  
  six_study <- readRDS(six_study_path)
  cat(glue("Loaded 6-study dataset: {nrow(six_study)} records, {ncol(six_study)} variables\n"))
  cat(glue("  Studies: {paste(unique(six_study$study_source), collapse = ', ')}\n\n"))
  
  # --- 2. REMOVE ANY EXISTING DHS ROWS (re-merge protection) ---
  if ("DHS" %in% unique(six_study$study_source)) {
    n_before <- nrow(six_study)
    six_study <- six_study %>% filter(study_source != "DHS")
    cat(glue("Removed {n_before - nrow(six_study)} existing DHS rows\n"))
    cat(glue("  Remaining: {nrow(six_study)} records\n\n"))
  }
  
  # --- 3. ALIGN COLUMNS ---
  cols_6study_only <- setdiff(names(six_study), names(dhs_unified))
  cols_dhs_only    <- setdiff(names(dhs_unified), names(six_study))
  
  cat(glue("Column alignment:\n"))
  cat(glue("  6-study variables: {ncol(six_study)}\n"))
  cat(glue("  DHS variables:     {ncol(dhs_unified)}\n"))
  cat(glue("  Shared:            {length(intersect(names(six_study), names(dhs_unified)))}\n"))
  cat(glue("  In 6-study only:   {length(cols_6study_only)}\n"))
  cat(glue("  In DHS only:       {length(cols_dhs_only)}\n\n"))
  
  for (col in cols_6study_only) dhs_unified[[col]] <- NA
  for (col in cols_dhs_only)    six_study[[col]]    <- NA
  
  all_cols <- union(names(six_study), names(dhs_unified))
  six_study   <- six_study[, all_cols]
  dhs_unified <- dhs_unified[, all_cols]
  
  # --- 4. TYPE ALIGNMENT (vectorised - fast) ---
  cat("Aligning column types...\n")
  common_types <- sapply(all_cols, function(col) {
    t6 <- class(six_study[[col]])[1]
    td <- class(dhs_unified[[col]])[1]
    if (t6 == td) return("match")
    if (t6 == "character" || td == "character") return("character")
    if (t6 %in% c("numeric", "integer") && td %in% c("numeric", "integer")) return("numeric")
    return("character")
  })
  
  cols_to_char <- names(common_types[common_types == "character"])
  cols_to_num  <- names(common_types[common_types == "numeric"])
  
  if (length(cols_to_char) > 0) {
    six_study   <- six_study   %>% mutate(across(all_of(cols_to_char), as.character))
    dhs_unified <- dhs_unified %>% mutate(across(all_of(cols_to_char), as.character))
  }
  if (length(cols_to_num) > 0) {
    six_study   <- six_study   %>% mutate(across(all_of(cols_to_num), as.numeric))
    dhs_unified <- dhs_unified %>% mutate(across(all_of(cols_to_num), as.numeric))
  }
  cat(glue("  Coerced {length(cols_to_char)} cols to character, {length(cols_to_num)} to numeric\n\n"))
  
  # --- 5. MERGE ---
  cat("Merging datasets...\n")
  merged <- bind_rows(six_study, dhs_unified)
  merged$unified_id <- paste0(merged$study_source, "_", seq_len(nrow(merged)))
  
  cat(glue("\nMERGED DATASET CREATED\n"))
  cat(glue("  Total records:     {nrow(merged)}\n"))
  cat(glue("  Total variables:   {ncol(merged)}\n"))
  cat(glue("  Studies:           {paste(unique(merged$study_source), collapse = ', ')}\n\n"))
  
  cat("Records by study:\n")
  merged %>% count(study_source, name = "n_records") %>% print()
  cat(glue("\nCountries: {n_distinct(merged$mat_country, na.rm = TRUE)}\n"))
  
  # --- 6. SAVE MERGED OUTPUT ---
  merged_output_dir <- dirname(six_study_path)
  if (!dir.exists(merged_output_dir)) dir.create(merged_output_dir, recursive = TRUE)
  cat(glue("\nSaving outputs to: {merged_output_dir}\n"))
  
  # RDS
  merged_rds <- file.path(merged_output_dir, paste0("merged_unified_dataset_v", MERGED_VERSION, ".rds"))
  saveRDS(merged, merged_rds)
  cat(glue("  Saved: {basename(merged_rds)} ({round(file.info(merged_rds)$size / 1024^2, 1)} MB)\n"))
  
  # CSV
  merged_csv <- file.path(merged_output_dir, paste0("merged_unified_dataset_v", MERGED_VERSION, ".csv"))
  write_csv(merged, merged_csv)
  cat(glue("  Saved: {basename(merged_csv)}\n"))
  
  # Stata (.dta)
  merged_stata <- merged
  long_names <- names(merged_stata)[nchar(names(merged_stata)) > 32]
  if (length(long_names) > 0) {
    cat(glue("  Truncating {length(long_names)} variable names for Stata\n"))
    names(merged_stata)[nchar(names(merged_stata)) > 32] <- substr(
      names(merged_stata)[nchar(names(merged_stata)) > 32], 1, 32
    )
    names(merged_stata) <- make.unique(names(merged_stata), sep = "_")
  }
  merged_dta <- file.path(merged_output_dir, paste0("merged_unified_dataset_v", MERGED_VERSION, ".dta"))
  haven::write_dta(merged_stata, merged_dta, version = 14)
  cat(glue("  Saved: {basename(merged_dta)}\n"))
  rm(merged_stata)
  
  # Excel - skip for large datasets
  if (nrow(merged) > 1000000) {
    cat("  Excel save skipped (>1M rows - use RDS/CSV instead)\n")
  } else {
    merged_xlsx <- file.path(merged_output_dir, paste0("merged_unified_dataset_v", MERGED_VERSION, ".xlsx"))
    tryCatch({
      openxlsx::write.xlsx(merged, merged_xlsx)
      cat(glue("  Saved: {basename(merged_xlsx)}\n"))
    }, error = function(e) {
      cat(glue("  Excel save skipped: {e$message}\n"))
    })
  }
  
  # 10% stratified sample
  set.seed(42)
  merged_sample <- merged %>%
    group_by(study_source) %>%
    sample_frac(0.1) %>%
    ungroup()
  
  merged_sample_rds <- file.path(merged_output_dir, paste0("merged_unified_10pct_sample_v", MERGED_VERSION, ".rds"))
  saveRDS(merged_sample, merged_sample_rds)
  merged_sample_csv <- file.path(merged_output_dir, paste0("merged_unified_10pct_sample_v", MERGED_VERSION, ".csv"))
  write_csv(merged_sample, merged_sample_csv)
  cat(glue("  Saved 10% sample: {nrow(merged_sample)} records\n"))
  
  # --- 7. OPTIONAL: COPY TO ONEDRIVE ---
  if (copy_to_onedrive && dir.exists(dirs$output_onedrive)) {
    cat("\n--- Copying merged files to OneDrive ---\n")
    merge_files <- c(
      paste0("merged_unified_dataset_v", MERGED_VERSION, ".rds"),
      paste0("merged_unified_dataset_v", MERGED_VERSION, ".csv"),
      paste0("merged_unified_dataset_v", MERGED_VERSION, ".dta"),
      paste0("merged_unified_10pct_sample_v", MERGED_VERSION, ".rds"),
      paste0("merged_unified_10pct_sample_v", MERGED_VERSION, ".csv")
    )
    for (f in merge_files) {
      src <- file.path(merged_output_dir, f)
      dst <- file.path(dirs$output_onedrive, f)
      if (file.exists(src)) {
        file.copy(src, dst, overwrite = TRUE)
        cat(glue("  Copied: {f}\n"))
      }
    }
  }
  
  # --- 8. FINAL SUMMARY ---
  cat(glue("\n\n{strrep('=', 60)}\n"))
  cat("MERGED UNIFIED DATASET - FINAL SUMMARY\n")
  cat(glue("{strrep('=', 60)}\n\n"))
  
  cat(glue("Version:             {MERGED_VERSION}\n"))
  cat(glue("Total records:       {nrow(merged)}\n"))
  cat(glue("Total variables:     {ncol(merged)}\n"))
  cat(glue("Studies:             {n_distinct(merged$study_source)}\n"))
  cat(glue("Countries:           {n_distinct(merged$mat_country, na.rm = TRUE)}\n\n"))
  
  cat("Outcome summary:\n")
  cat(glue("  Live births:       {sum(merged$out_livebirth == 'Yes', na.rm = TRUE)}\n"))
  cat(glue("  Stillbirths:       {sum(merged$out_stillbirth == 'Yes', na.rm = TRUE)}\n"))
  cat(glue("  Neonatal deaths:   {sum(merged$out_nnd == 'Yes', na.rm = TRUE)}\n"))
  cat(glue("  With GPS coords:   {sum(!is.na(merged$env_latitude) | !is.na(merged$loc_latitude), na.rm = TRUE)}\n\n"))
  
  cat(glue("Output directory: {merged_output_dir}\n"))
  cat(glue("Merged file:      merged_unified_dataset_v{MERGED_VERSION}.rds\n\n"))
  
  cat("Next steps:\n")
  cat(glue("1. Run analysis pipeline on merged_unified_dataset_v{MERGED_VERSION}.rds\n"))
  cat("2. Run environmental pipeline to add PM2.5, ERA5, CHIRPS variables\n")
  cat("3. Run SGA calculation using gigs R package\n")
  
  rm(six_study, merged_sample)
  gc(verbose = FALSE)
}
```


# Utility: Bootstrap Checkpoints from Existing Output

```{r bootstrap-checkpoints, eval=FALSE}
# ============================================================================
# BOOTSTRAP: Create checkpoint files from existing compiled output
# ============================================================================
# Run ONCE if you already have a compiled dhs_unified_dataset.rds from a
# previous run (before checkpointing was added). Creates the checkpoint files
# so Step 7 skips reprocessing.
# ============================================================================

cat("=== BOOTSTRAPPING CHECKPOINTS FROM EXISTING OUTPUT ===\n\n")

candidate_paths <- c(
  file.path(dirs$output_onedrive, "dhs_unified_dataset.rds"),
  file.path(dirs$output, "dhs_unified_dataset.rds")
)

existing_rds <- NULL
for (p in candidate_paths) {
  if (file.exists(p)) { existing_rds <- p; break }
}

if (is.null(existing_rds)) {
  stop("No existing dhs_unified_dataset.rds found. Run the full pipeline first.")
}

cat(glue("Found: {existing_rds}\n"))
existing_data <- readRDS(existing_rds)
cat(glue("Loaded: {nrow(existing_data)} records\n\n"))

livebirths_df <- existing_data %>%
  filter(out_livebirth == "Yes" | (out_stillbirth != "Yes" & is.na(out_stillbirth)))
stillbirths_df <- existing_data %>%
  filter(out_stillbirth == "Yes")

if ("dhs_survey_id" %in% names(existing_data)) {
  all_br_processed <- split(livebirths_df, livebirths_df$dhs_survey_id)
  all_stillbirths  <- split(stillbirths_df, stillbirths_df$dhs_survey_id)
  
  processing_log <- lapply(unique(existing_data$dhs_survey_id), function(sid) {
    survey_data <- existing_data %>% filter(dhs_survey_id == sid)
    list(
      survey_id = sid,
      country = if ("mat_country" %in% names(survey_data)) survey_data$mat_country[1] else NA,
      year = if ("studyyear" %in% names(survey_data)) survey_data$studyyear[1] else NA,
      status = "success",
      module = if ("module" %in% names(survey_data)) survey_data$module[1] else NA,
      sb_method = "bootstrapped",
      n_births = sum(survey_data$out_livebirth == "Yes", na.rm = TRUE),
      n_stillbirths = sum(survey_data$out_stillbirth == "Yes", na.rm = TRUE)
    )
  })
  names(processing_log) <- unique(existing_data$dhs_survey_id)
} else {
  all_br_processed <- list(all = livebirths_df)
  all_stillbirths  <- list(all = stillbirths_df)
  processing_log   <- list(all = list(survey_id = "all", status = "success"))
}

if (!dir.exists(dirs$dhs_checkpts)) dir.create(dirs$dhs_checkpts, recursive = TRUE)

saveRDS(all_br_processed, file.path(dirs$dhs_checkpts, "all_br_processed.rds"))
saveRDS(all_stillbirths,  file.path(dirs$dhs_checkpts, "all_stillbirths.rds"))
saveRDS(processing_log,   file.path(dirs$dhs_checkpts, "processing_log.rds"))

cat(glue("\nSurveys in checkpoint: {length(all_br_processed)}\n"))
cat(glue("Total live births:    {sum(sapply(all_br_processed, nrow))}\n"))
cat(glue("Total stillbirths:    {sum(sapply(all_stillbirths, nrow))}\n"))
cat("\n=== BOOTSTRAP COMPLETE ===\n")

rm(existing_data, livebirths_df, stillbirths_df)
gc(verbose = FALSE)
```


# Utility: Clear Checkpoints (Force Reprocessing)

```{r clear-checkpoints, eval=FALSE}
# ============================================================================
# CLEAR CHECKPOINTS — forces Step 7 to reprocess ALL surveys
# ============================================================================
# Only run this if you want to reprocess everything from scratch.
# After clearing, re-run the pipeline — Step 7 will take 6-10 hours.
# ============================================================================

checkpoint_files <- list.files(dirs$dhs_checkpts, full.names = TRUE)
if (length(checkpoint_files) > 0) {
  cat(glue("Deleting {length(checkpoint_files)} checkpoint files...\n"))
  file.remove(checkpoint_files)
  cat("Checkpoints cleared. Step 7 will reprocess on next run.\n")
} else {
  cat("No checkpoint files found.\n")
}
```
