---
title: "Integrated Environmental Data Pipeline v3.3"
author: "Joseph Akuze (LSHTM)"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
version: '3.3'
subtitle: "Complete Environmental Data Extraction for Stillbirth and Neonatal Death Prediction"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 6
)

# ============================================================================
# REQUIRED PACKAGES
# ============================================================================
suppressPackageStartupMessages({
  library(tidyverse)
  library(terra)
  library(sf)
  library(lubridate)
  library(ncdf4)
  library(haven)
  library(tools)
  library(knitr)
  library(janitor)
})

# Try to load elevatr (for SRTM download) - optional
has_elevatr <- require(elevatr, quietly = TRUE)

VERSION <- "3.3"
```

# Overview

This integrated pipeline performs **complete environmental data processing** in a single workflow:

**Part A: Environmental Linkage Creation (from v2.0)**

1. Load unified dataset
2. Create facility and country geocode references
3. Extract geographic-temporal data
4. Geocoding and coordinate enhancement
5. Calculate temporal windows (trimesters, periconception)
6. Create environmental variable placeholders

**Part B: Environmental Data Extraction**

7. Extract elevation and slope (SRTM)
8. Extract ERA5 climate data (temperature, humidity, precipitation, heat index)
9. Extract PM2.5 air quality (ACAG)
10. Classify seasons
11. Quality summary and validation
12. Export final dataset

---

# Phase 0: Configuration

```{r phase0-config}
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║         INTEGRATED ENVIRONMENTAL PIPELINE v", VERSION, "                       ║\n", sep = "")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Input: Your unified dataset (merged = 6-study + DHS; fallback to 6-study only)
# v3.3: Updated to read v10.16 (produced by DHS Pipeline v7.2)
FILE_UNIFIED <- "output/merged_unified_dataset_v10.16.rds"
FILE_UNIFIED_DTA <- "output/merged_unified_dataset_v10.16.dta"
FILE_UNIFIED_CSV <- "output/merged_unified_dataset_v10.16.csv"

# Optional: Raw PRECISE data for GPS coordinates
FILE_PRECISE_RAW <- "data/PRECISE_dataset.dta"

# Climate data directories
DIR_ERA5 <- "data/era5"
DIR_SRTM <- "data/srtm"
DIR_PM25 <- "data/pm25"
DIR_NDVI <- "data/ndvi"

# ERA5 file paths (specific files)
FILE_ERA5_TEMP <- "data/era5/data_stream-moda_stepType-avgua.nc"  # t2m, d2m
FILE_ERA5_PRECIP <- "data/era5/era5_monthly_africa.nc"            # tp

# Output directory and file name
DIR_OUTPUT <- "output/environmental"
OUTPUT_FILE <- "environmental_linkage_v3.3"

# Create directories
dir.create(DIR_OUTPUT, recursive = TRUE, showWarnings = FALSE)
dir.create(DIR_SRTM, recursive = TRUE, showWarnings = FALSE)

cat("Configuration:\n")
cat(sprintf("  Unified dataset: %s\n", FILE_UNIFIED))
cat(sprintf("  ERA5 temp file:  %s\n", FILE_ERA5_TEMP))
cat(sprintf("  ERA5 precip file: %s\n", FILE_ERA5_PRECIP))
cat(sprintf("  PM2.5 directory: %s\n", DIR_PM25))
cat(sprintf("  Output:          %s/%s.*\n", DIR_OUTPUT, OUTPUT_FILE))
```

## 0.1 Utility Functions

```{r phase0-utilities}
# ============================================================================
# DATA LOADING FUNCTION
# ============================================================================
safe_load_data <- function(filepath, encoding = "UTF-8") {
  if (!file.exists(filepath)) {
    warning(sprintf("File not found: %s", filepath))
    return(NULL)
  }
  
  ext <- tolower(tools::file_ext(filepath))
  
  tryCatch({
    df <- switch(ext,
      "csv" = readr::read_csv(filepath, show_col_types = FALSE),
      "dta" = haven::read_dta(filepath),
      "rds" = readRDS(filepath),
      readr::read_csv(filepath, show_col_types = FALSE)
    )
    message(sprintf("✓ Loaded %s: %d rows x %d cols", basename(filepath), nrow(df), ncol(df)))
    return(df)
  }, error = function(e) {
    warning(sprintf("✗ Failed to load %s: %s", filepath, e$message))
    return(NULL)
  })
}

# ============================================================================
# SAFE COLUMN ACCESS
# ============================================================================
get_col_chr <- function(df, col_name) {
  if (col_name %in% names(df)) return(as.character(df[[col_name]]))
  return(rep(NA_character_, nrow(df)))
}

get_col_num <- function(df, col_name) {
  if (col_name %in% names(df)) return(suppressWarnings(as.numeric(df[[col_name]])))
  return(rep(NA_real_, nrow(df)))
}

get_col_date <- function(df, col_name) {
  if (col_name %in% names(df)) {
    val <- df[[col_name]]
    if (inherits(val, "Date")) return(val)
    if (inherits(val, "POSIXct")) return(as.Date(val))
    return(suppressWarnings(as.Date(val)))
  }
  return(rep(as.Date(NA), nrow(df)))
}

# ============================================================================
# ERA5 EXTRACTION FUNCTION (from working script)
# ============================================================================
extract_era5_variable <- function(nc_file, var_name, coords_df) {
  
  nc <- nc_open(nc_file)
  
  lon <- ncvar_get(nc, "longitude")
  lat <- ncvar_get(nc, "latitude")
  var_data <- ncvar_get(nc, var_name)
  
  nc_close(nc)
  
  result <- rep(NA_real_, nrow(coords_df))
  
  for (i in 1:nrow(coords_df)) {
    target_lon <- coords_df$env_longitude[i]
    target_lat <- coords_df$env_latitude[i]
    
    lon_idx <- which.min(abs(lon - target_lon))
    lat_idx <- which.min(abs(lat - target_lat))
    
    # var_data is [lon, lat, time] - get mean across all times
    if (length(dim(var_data)) == 3) {
      vals <- var_data[lon_idx, lat_idx, ]
    } else {
      vals <- var_data[lon_idx, lat_idx]
    }
    
    result[i] <- mean(vals, na.rm = TRUE)
  }
  
  return(result)
}
```

---

# Part A: Environmental Linkage Creation

# Phase 1: Load Unified Dataset

```{r phase1-load}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                 PHASE 1: LOADING UNIFIED DATASET                           ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

unified_dataset <- NULL

if (file.exists(FILE_UNIFIED)) {
  unified_dataset <- safe_load_data(FILE_UNIFIED)
} else if (file.exists(FILE_UNIFIED_DTA)) {
  unified_dataset <- safe_load_data(FILE_UNIFIED_DTA)
} else if (file.exists(FILE_UNIFIED_CSV)) {
  unified_dataset <- safe_load_data(FILE_UNIFIED_CSV)
} else {
  # Fallback: try other versions (most recent first)
  fallback_paths <- c(
    "output/merged_unified_dataset_v10.16.rds",
    "output/unified_dataset_v10.16.rds",
    "output/merged_unified_dataset_v10.15.rds",
    "output/unified_dataset_v10.15.rds",
    "output/merged_unified_dataset_v10.14.rds",
    "output/unified_dataset_v10.14.rds",
    "output/merged_unified_dataset_v10.13.rds",
    "output/unified_dataset_v10.13.rds",
    "output/unified_dataset_v10.12.rds",
    "output/unified_dataset_v10.11.rds"
  )
  for (fp in fallback_paths) {
    if (file.exists(fp)) {
      cat(sprintf("  Using fallback: %s\n", fp))
      unified_dataset <- safe_load_data(fp)
      break
    }
  }
}

if (is.null(unified_dataset)) {
  stop("✗ FATAL: Could not load unified dataset!")
}

cat("\n=== UNIFIED DATASET SUMMARY ===\n")
cat(sprintf("Total records:   %s\n", format(nrow(unified_dataset), big.mark = ",")))
cat(sprintf("Total variables: %d\n", ncol(unified_dataset)))
cat(sprintf("Studies: %s\n", paste(unique(unified_dataset$study_source), collapse = ", ")))

study_summary <- unified_dataset %>%
  group_by(study_source) %>%
  summarise(n = n(), .groups = "drop")
print(kable(study_summary, caption = "Records by Study"))
```

# Phase 2: Geographic Reference Data

```{r phase2-geocodes}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                 PHASE 2: GEOGRAPHIC REFERENCE DATA                         ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Facility geocodes
facility_geocodes <- tribble(
  ~study_source, ~facility_clean, ~latitude, ~longitude, ~accuracy_level,
  "NCOPS", "mulago national referral hospital", 0.3370, 32.5766, "facility",
  "NCOPS", "kawempe national referral hospital", 0.3645, 32.5506, "facility",
  "NCOPS", "nsambya hospital", 0.3026, 32.5848, "facility",
  "NCOPS", "jinja regional referral hospital", 0.4244, 33.2041, "facility",
  "NCOPS", "mbale regional referral hospital", 1.0806, 34.1750, "facility",
  "NCOPS", "mbarara regional referral hospital", -0.6127, 30.6464, "facility",
  "NCOPS", "gulu regional referral hospital", 2.7691, 32.2989, "facility",
  "NCOPS", "hoima regional referral hospital", 1.4334, 31.3523, "facility",
  "NCOPS", "soroti regional referral hospital", 1.7111, 33.6097, "facility",
  "NCOPS", "lira regional referral hospital", 2.2494, 32.5434, "facility",
  "NCOPS", "fort portal regional referral hospital", 0.6710, 30.2750, "facility",
  "NCOPS", "arua regional referral hospital", 3.0293, 30.9067, "facility",
  "NCOPS", "kabale regional referral hospital", -1.2503, 29.9881, "facility",
  "NCOPS", "moroto regional referral hospital", 2.5288, 34.6613, "facility",
  "PTBi", "mulago national referral hospital", 0.3370, 32.5766, "facility",
  "PTBi", "kawempe national referral hospital", 0.3645, 32.5506, "facility",
  "PTBi", "jinja regional referral hospital", 0.4244, 33.2041, "facility",
  "PTBi", "masaka regional referral hospital", -0.3372, 31.7136, "facility",
  "ALERT", "jinja regional hospital", 0.4244, 33.2041, "facility",
  "ALERT", "1. jinja regional hospital", 0.4244, 33.2041, "facility",
  "ALERT", "ndanda hospital", -10.6500, 39.6167, "facility",
  "ALERT", "3.1 ndanda", -10.6500, 39.6167, "facility",
  "ALERT", "kamuzu central hospital", -13.9833, 33.7833, "facility",
  "PRECISE", "farafenni", 13.5667, -15.6000, "district",
  "PRECISE", "mariakani", -3.8591, 39.4783, "facility",
  "PRECISE", "rabai", -3.9333, 39.5500, "facility",
  "PRECISE", "manhica", -25.4033, 32.8072, "district",
  "PRECISE", "xinavane", -25.0500, 32.8000, "facility",
  "EN-INDEPTH", "matlab", 23.3500, 90.7333, "district",
  "EN-INDEPTH", "bandim", 11.8500, -15.6000, "district",
  "EN-INDEPTH", "dabat", 13.0333, 37.7500, "district",
  "EN-INDEPTH", "kintampo", 8.0500, -1.7333, "district",
  "EN-INDEPTH", "navrongo", 10.8833, -1.0833, "district",
  "EN-INDEPTH", "iganga/mayuge", 0.6500, 33.5000, "district",
  "EN-INDEPTH", "iganga", 0.6500, 33.5000, "district"
)

# Country centroids
# v3.3: Added all 42 SSA DHS countries that were previously missing
# (v3.2 had only 30 entries, missing ~29 SSA countries used in DHS pipeline)
country_centroids <- tribble(
  ~country_clean, ~latitude, ~longitude,
  # --- SSA countries (DHS pipeline) ---
  "uganda", 1.3733, 32.2903,
  "kenya", -0.0236, 37.9062,
  "tanzania", -6.3690, 34.8888,
  "malawi", -13.2543, 34.3015,
  "mozambique", -18.6657, 35.5296,
  "the gambia", 13.4432, -15.3101,
  "gambia", 13.4432, -15.3101,
  "ghana", 7.9465, -1.0232,
  "ethiopia", 9.1450, 40.4897,
  "guinea-bissau", 11.8037, -15.1804,
  "angola", -11.2027, 17.8739,
  "niger", 17.6078, 8.0817,
  "nigeria", 9.0820, 7.4898,
  "democratic republic of the congo", -4.0383, 21.7587,
  # v3.3 additions - SSA countries missing from v3.2:
  "benin", 9.3077, 2.3158,
  "burkina faso", 12.2383, -1.5616,
  "burundi", -3.3731, 29.9189,
  "cameroon", 7.3697, 12.3547,
  "cape verde", 16.5388, -23.0418,
  "central african republic", 6.6111, 20.9394,
  "chad", 15.4542, 18.7322,
  "comoros", -11.6455, 43.3333,
  "congo", -0.2280, 15.8277,
  "congo (brazzaville)", -0.2280, 15.8277,
  "cote d'ivoire", 7.5400, -5.5471,
  "ivory coast", 7.5400, -5.5471,
  "eritrea", 15.1794, 39.7823,
  "eswatini", -26.5225, 31.4659,
  "swaziland", -26.5225, 31.4659,
  "gabon", -0.8037, 11.6094,
  "guinea", 9.9456, -9.6966,
  "lesotho", -29.6100, 28.2336,
  "liberia", 6.4281, -9.4295,
  "madagascar", -18.7669, 46.8691,
  "mali", 17.5707, -3.9962,
  "mauritania", 21.0079, -10.9408,
  "namibia", -22.9576, 18.4904,
  "rwanda", -1.9403, 29.8739,
  "sao tome and principe", 0.1864, 6.6131,
  "senegal", 14.4974, -14.4524,
  "sierra leone", 8.4606, -11.7799,
  "south africa", -30.5595, 22.9375,
  "sudan", 12.8628, 30.2176,
  "togo", 8.6195, 0.8248,
  "zambia", -13.1339, 27.8493,
  "zimbabwe", -19.0154, 29.1549,
  # --- Name variations (actual names found in dataset that differ from above) ---
  # Confirmed from environmental availability diagnostic (background task bae6942):
  "congo democratic republic", -4.0383, 21.7587,
  "dem rep of the congo", -4.0383, 21.7587,
  "viet nam", 14.0583, 108.2772,
  # --- Non-SSA countries (WHOMCS and other studies) ---
  "afghanistan", 33.9391, 67.7100,
  "argentina", -38.4161, -63.6167,
  "bangladesh", 23.6850, 90.3563,
  "brazil", -14.2350, -51.9253,
  "cambodia", 12.5657, 104.9910,
  "china", 35.8617, 104.1954,
  "ecuador", -1.8312, -78.1834,
  "india", 20.5937, 78.9629,
  "japan", 36.2048, 138.2529,
  "jordan", 30.5852, 36.2384,
  "mexico", 23.6345, -102.5528,
  "nepal", 28.3949, 84.1240,
  "nicaragua", 12.8654, -85.2072,
  "pakistan", 30.3753, 69.3451,
  "peru", -9.1900, -75.0152,
  "philippines", 12.8797, 121.7740,
  "thailand", 15.8700, 100.9925,
  "vietnam", 14.0583, 108.2772,
  # v3.3 additions - non-SSA WHOMCS countries missing from v3.2:
  "lebanon", 33.8547, 35.8623,
  "mongolia", 46.8625, 103.8467,
  "occupied palestinian territory", 31.9522, 35.2332,
  "palestine", 31.9522, 35.2332,
  "paraguay", -23.4425, -58.4438,
  "qatar", 25.3548, 51.1839,
  "sri lanka", 7.8731, 80.7718
)

cat(sprintf("✓ Loaded %d facility geocodes\n", nrow(facility_geocodes)))
cat(sprintf("✓ Loaded %d country centroids\n", nrow(country_centroids)))
```

## 2.1 Load PRECISE GPS Coordinates

```{r phase2-precise}
precise_coords <- NULL

if (file.exists(FILE_PRECISE_RAW)) {
  raw_precise <- safe_load_data(FILE_PRECISE_RAW)
  
  if (!is.null(raw_precise)) {
    id_col <- if ("f2a_participant_id" %in% names(raw_precise)) "f2a_participant_id" else 
              if ("f2a_precise_id" %in% names(raw_precise)) "f2a_precise_id" else NULL
    
    if (!is.null(id_col)) {
      precise_coords <- raw_precise %>%
        select(raw_id = all_of(id_col),
               any_of(c("villagelat", "villagelong"))) %>%
        rename_with(~case_when(
          . == "villagelat" ~ "gps_latitude",
          . == "villagelong" ~ "gps_longitude",
          TRUE ~ .
        )) %>%
        filter(!is.na(gps_latitude) & !is.na(gps_longitude)) %>%
        mutate(raw_id = as.character(raw_id),
               gps_latitude = as.numeric(gps_latitude),
               gps_longitude = as.numeric(gps_longitude)) %>%
        filter(gps_latitude >= -90 & gps_latitude <= 90) %>%
        group_by(raw_id) %>%
        slice_head(n = 1) %>%
        ungroup()
      
      cat(sprintf("✓ Extracted %d GPS coordinates from PRECISE\n", nrow(precise_coords)))
    }
  }
} else {
  cat("ℹ PRECISE raw file not found - using facility geocodes only\n")
}
```

# Phase 3: Extract Geographic-Temporal Data

```{r phase3-extract}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║           PHASE 3: EXTRACT GEOGRAPHIC-TEMPORAL DATA                        ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

environmental_linkage <- unified_dataset %>%
  transmute(
    unified_id = unified_id,
    study_source = study_source,
    raw_id = raw_id,
    geo_country = get_col_chr(., "mat_country"),
    geo_facility = get_col_chr(., "mat_facility"),
    geo_district = get_col_chr(., "mat_district"),
    # v3.3 FIX: Also check env_latitude/env_longitude (used by DHS pipeline)
    # Previously only read loc_latitude/loc_longitude, which missed DHS GPS data
    geo_latitude = dplyr::coalesce(get_col_num(., "loc_latitude"), get_col_num(., "env_latitude")),
    geo_longitude = dplyr::coalesce(get_col_num(., "loc_longitude"), get_col_num(., "env_longitude")),
    geo_urban_rural = get_col_chr(., "mat_urban_rural"),
    date_delivery = get_col_date(., "out_dob"),
    date_death = get_col_date(., "out_dod"),
    conception_date_unified = get_col_date(., "conception_date"),
    ga_weeks = get_col_num(., "out_ga_weeks"),
    studyyear = get_col_num(., "studyyear"),
    out_stillbirth = get_col_chr(., "out_stillbirth"),
    out_nnd = get_col_chr(., "out_nnd")
  )

cat(sprintf("✓ Extracted %d records\n", nrow(environmental_linkage)))
```

# Phase 4: Geocoding and Coordinate Enhancement

```{r phase4-geocode}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║              PHASE 4: GEOCODING AND COORDINATE ENHANCEMENT                 ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Merge PRECISE GPS if available
if (!is.null(precise_coords) && nrow(precise_coords) > 0) {
  environmental_linkage <- environmental_linkage %>%
    mutate(raw_id = as.character(raw_id)) %>%
    left_join(precise_coords, by = "raw_id") %>%
    mutate(
      geo_latitude = case_when(
        study_source == "PRECISE" & !is.na(gps_latitude) ~ gps_latitude,
        TRUE ~ geo_latitude
      ),
      geo_longitude = case_when(
        study_source == "PRECISE" & !is.na(gps_longitude) ~ gps_longitude,
        TRUE ~ geo_longitude
      )
    ) %>%
    select(-any_of(c("gps_latitude", "gps_longitude")))
  
  cat("✓ Merged PRECISE GPS coordinates\n")
}

# Match facilities
environmental_linkage <- environmental_linkage %>%
  mutate(facility_clean = tolower(trimws(geo_facility))) %>%
  left_join(facility_geocodes %>% select(study_source, facility_clean, lat_fac = latitude, 
                                          lon_fac = longitude, accuracy_fac = accuracy_level),
            by = c("study_source", "facility_clean"), relationship = "many-to-one")

# Match countries
environmental_linkage <- environmental_linkage %>%
  mutate(country_clean = tolower(trimws(geo_country))) %>%
  left_join(country_centroids %>% select(country_clean, lat_ctr = latitude, lon_ctr = longitude),
            by = "country_clean", relationship = "many-to-one")

# Prioritize coordinates
environmental_linkage <- environmental_linkage %>%
  mutate(
    env_latitude = case_when(
      !is.na(geo_latitude) & geo_latitude >= -90 & geo_latitude <= 90 ~ geo_latitude,
      !is.na(lat_fac) ~ lat_fac,
      !is.na(lat_ctr) ~ lat_ctr,
      TRUE ~ NA_real_
    ),
    env_longitude = case_when(
      !is.na(geo_longitude) & geo_longitude >= -180 & geo_longitude <= 180 ~ geo_longitude,
      !is.na(lon_fac) ~ lon_fac,
      !is.na(lon_ctr) ~ lon_ctr,
      TRUE ~ NA_real_
    ),
    coordinate_source = case_when(
      !is.na(geo_latitude) & geo_latitude >= -90 ~ "direct_gps",
      !is.na(lat_fac) ~ accuracy_fac,
      !is.na(lat_ctr) ~ "country_centroid",
      TRUE ~ "missing"
    )
  ) %>%
  select(-facility_clean, -country_clean, -lat_fac, -lon_fac, -accuracy_fac, -lat_ctr, -lon_ctr)

cat("\n=== COORDINATE SOURCE SUMMARY ===\n")
coord_summary <- environmental_linkage %>%
  count(study_source, coordinate_source) %>%
  pivot_wider(names_from = coordinate_source, values_from = n, values_fill = 0)
print(kable(coord_summary))

# v3.3: Diagnostic - show coordinate coverage by country
cat("\n=== COORDINATE COVERAGE BY COUNTRY (v3.3 diagnostic) ===\n")
country_coord_summary <- environmental_linkage %>%
  group_by(geo_country, study_source) %>%
  summarise(
    n_total = n(),
    n_with_coords = sum(!is.na(env_latitude)),
    pct_with_coords = round(100 * mean(!is.na(env_latitude)), 1),
    coord_source = paste(unique(coordinate_source[!is.na(env_latitude)]), collapse = "/"),
    .groups = "drop"
  ) %>%
  arrange(pct_with_coords, desc(n_total))
print(kable(country_coord_summary, caption = "Countries with lowest coordinate coverage (for debugging)"))
```

# Phase 5: Calculate Temporal Windows

```{r phase5-temporal}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                PHASE 5: CALCULATE TEMPORAL WINDOWS                         ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

environmental_linkage <- environmental_linkage %>%
  mutate(
    date_conception = case_when(
      !is.na(conception_date_unified) ~ conception_date_unified,
      !is.na(date_delivery) & !is.na(ga_weeks) ~ date_delivery - (ga_weeks * 7),
      TRUE ~ as.Date(NA)
    ),
    gestational_age_days = as.numeric(date_delivery - date_conception),
    gestational_age_weeks = gestational_age_days / 7,
    trimester1_start = date_conception,
    trimester1_end = date_conception + 91,
    trimester2_start = date_conception + 92,
    trimester2_end = date_conception + 182,
    trimester3_start = date_conception + 183,
    trimester3_end = date_delivery,
    periconception_start = date_conception - 30,
    periconception_end = date_conception + 60,
    delivery_month = month(date_delivery),
    delivery_year = year(date_delivery),
    conception_month = month(date_conception),
    conception_year = year(date_conception)
  ) %>%
  select(-conception_date_unified)

cat("✓ Calculated temporal windows\n")
```

# Phase 6: Create Placeholders and Quality Flags

```{r phase6-placeholders}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║            PHASE 6: CREATE PLACEHOLDERS AND FLAGS                          ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

environmental_linkage <- environmental_linkage %>%
  mutate(
    # Climate (will be populated)
    env_temp_mean_delivery = NA_real_,
    env_temp_mean_pregnancy = NA_real_,
    env_humidity_delivery = NA_real_,
    env_humidity_pregnancy = NA_real_,
    env_precipitation_delivery = NA_real_,
    env_precipitation_pregnancy = NA_real_,
    env_heat_index_delivery = NA_real_,
    
    # Seasons (will be populated)
    env_season_conception = NA_character_,
    env_season_delivery = NA_character_,
    
    # Air quality (will be populated)
    env_pm25_annual = NA_real_,
    env_pm25_delivery = NA_real_,
    env_pm25_pregnancy = NA_real_,
    
    # Topography (will be populated)
    env_elevation = NA_real_,
    env_slope = NA_real_,
    
    # Vegetation
    env_ndvi_delivery = NA_real_,
    env_ndvi_pregnancy = NA_real_,
    
    # Quality flags
    has_coordinates = !is.na(env_latitude) & !is.na(env_longitude),
    has_delivery_date = !is.na(date_delivery),
    has_conception_date = !is.na(date_conception),
    has_death_date = !is.na(date_death),
    linkage_ready = has_coordinates & has_delivery_date
  )

cat(sprintf("✓ Records with coordinates: %s\n", format(sum(environmental_linkage$has_coordinates), big.mark = ",")))
cat(sprintf("✓ Linkage-ready records: %s\n", format(sum(environmental_linkage$linkage_ready), big.mark = ",")))
```

---

# Part B: Environmental Data Extraction

# Phase 7: Extract Elevation and Slope

```{r phase7-elevation}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                PHASE 7: EXTRACT ELEVATION AND SLOPE                        ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

elev_file <- file.path(DIR_SRTM, "srtm_elevation.tif")
slope_file <- file.path(DIR_SRTM, "srtm_slope.tif")

if (file.exists(elev_file) && file.exists(slope_file)) {
  
  cat("✅ SRTM files found\n")
  
  # Filter to records with coordinates
  coords <- environmental_linkage %>%
    filter(!is.na(env_latitude) & !is.na(env_longitude))
  
  if (nrow(coords) > 0) {
    # Create SpatVector (same approach as working script)
    coord_vect <- vect(coords, geom = c("env_longitude", "env_latitude"), crs = "EPSG:4326")
    
    # Extract elevation
    cat("   Extracting elevation...")
    elev_rast <- rast(elev_file)
    elev_extracted <- terra::extract(elev_rast, coord_vect)
    coords$env_elevation <- elev_extracted[[2]]
    cat(sprintf(" ✅ (%.0f to %.0f m)\n",
                min(coords$env_elevation, na.rm = TRUE),
                max(coords$env_elevation, na.rm = TRUE)))
    
    # Extract slope
    cat("   Extracting slope...")
    slope_rast <- rast(slope_file)
    slope_extracted <- terra::extract(slope_rast, coord_vect)
    coords$env_slope <- slope_extracted[[2]]
    cat(sprintf(" ✅ (%.1f to %.1f degrees)\n",
                min(coords$env_slope, na.rm = TRUE),
                max(coords$env_slope, na.rm = TRUE)))
    
    # Merge back
    environmental_linkage <- environmental_linkage %>%
      select(-env_elevation, -env_slope) %>%
      left_join(coords %>% select(unified_id, env_elevation, env_slope), by = "unified_id")
  }
  
} else {
  cat("⚠️  SRTM files not found\n")
  cat("   Run the optional SRTM download chunk at the end\n")
}
```

# Phase 8: Extract ERA5 Climate Data

```{r phase8-era5}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                PHASE 8: EXTRACT ERA5 CLIMATE DATA                          ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Check for ERA5 files
if (file.exists(FILE_ERA5_TEMP)) {
  
  cat("✅ ERA5 files found\n")
  
  # Filter to records with coordinates (same approach as working script)
  coords <- environmental_linkage %>%
    filter(!is.na(env_latitude) & !is.na(env_longitude))
  
  cat(sprintf("   Records with coordinates: %s\n\n", format(nrow(coords), big.mark = ",")))
  
  if (nrow(coords) > 0) {
    
    # Extract temperature (t2m) - convert Kelvin to Celsius
    cat("   Extracting temperature (t2m)...")
    coords$env_temp_mean_delivery <- extract_era5_variable(FILE_ERA5_TEMP, "t2m", coords) - 273.15
    cat(sprintf(" ✅ (%.1f to %.1f °C)\n",
                min(coords$env_temp_mean_delivery, na.rm = TRUE),
                max(coords$env_temp_mean_delivery, na.rm = TRUE)))
    
    # Extract dewpoint (d2m) and calculate humidity
    cat("   Extracting dewpoint (d2m) and calculating humidity...")
    dewpoint <- extract_era5_variable(FILE_ERA5_TEMP, "d2m", coords) - 273.15
    temp <- coords$env_temp_mean_delivery
    
    # Magnus formula for relative humidity
    coords$env_humidity_delivery <- 100 * exp((17.625 * dewpoint) / (243.04 + dewpoint)) /
      exp((17.625 * temp) / (243.04 + temp))
    coords$env_humidity_delivery <- pmin(pmax(coords$env_humidity_delivery, 0), 100)
    cat(sprintf(" ✅ (%.1f to %.1f %%)\n",
                min(coords$env_humidity_delivery, na.rm = TRUE),
                max(coords$env_humidity_delivery, na.rm = TRUE)))
    
    # Extract precipitation (tp) - convert m to mm/month
    if (file.exists(FILE_ERA5_PRECIP)) {
      cat("   Extracting precipitation (tp)...")
      coords$env_precipitation_delivery <- extract_era5_variable(FILE_ERA5_PRECIP, "tp", coords) * 1000 * 30
      cat(sprintf(" ✅ (%.1f to %.1f mm/month)\n",
                  min(coords$env_precipitation_delivery, na.rm = TRUE),
                  max(coords$env_precipitation_delivery, na.rm = TRUE)))
    }
    
    # Calculate Heat Index
    cat("   Calculating heat index...")
    T_f <- coords$env_temp_mean_delivery * 9/5 + 32  # Celsius to Fahrenheit
    RH <- coords$env_humidity_delivery
    
    coords$env_heat_index_delivery <- -42.379 + 2.04901523*T_f + 10.14333127*RH -
      0.22475541*T_f*RH - 0.00683783*T_f^2 - 0.05481717*RH^2 +
      0.00122874*T_f^2*RH + 0.00085282*T_f*RH^2 - 0.00000199*T_f^2*RH^2
    
    coords$env_heat_index_delivery <- (coords$env_heat_index_delivery - 32) * 5/9
    cat(sprintf(" ✅ (%.1f to %.1f °C)\n",
                min(coords$env_heat_index_delivery, na.rm = TRUE),
                max(coords$env_heat_index_delivery, na.rm = TRUE)))
    
    # Add pregnancy-period copies
    coords$env_temp_mean_pregnancy <- coords$env_temp_mean_delivery
    coords$env_humidity_pregnancy <- coords$env_humidity_delivery
    coords$env_precipitation_pregnancy <- coords$env_precipitation_delivery
    
    # Merge back
    climate_cols <- c("unified_id", "env_temp_mean_delivery", "env_temp_mean_pregnancy",
                      "env_humidity_delivery", "env_humidity_pregnancy",
                      "env_precipitation_delivery", "env_precipitation_pregnancy",
                      "env_heat_index_delivery")
    
    environmental_linkage <- environmental_linkage %>%
      select(-any_of(climate_cols[climate_cols != "unified_id"])) %>%
      left_join(coords %>% select(any_of(climate_cols)), by = "unified_id")
  }
  
} else {
  cat("⚠️  ERA5 files not found\n")
  cat("   Expected: ", FILE_ERA5_TEMP, "\n")
}
```

# Phase 9: Extract PM2.5 Air Quality

```{r phase9-pm25}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                PHASE 9: EXTRACT PM2.5 AIR QUALITY                          ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

pm25_files <- list.files(DIR_PM25, pattern = "\\.nc$", full.names = TRUE)

if (length(pm25_files) > 0) {
  
  cat(sprintf("✅ Found %d PM2.5 files\n", length(pm25_files)))
  
  # Filter to records with coordinates and dates
  coords <- environmental_linkage %>%
    filter(!is.na(env_latitude) & !is.na(env_longitude) & !is.na(date_delivery))
  
  # Initialize PM2.5 columns in coords
  coords$env_pm25_annual <- NA_real_
  coords$env_pm25_delivery <- NA_real_
  coords$env_pm25_pregnancy <- NA_real_
  
  # Get valid years (1998-2023)
  years_available <- unique(na.omit(coords$delivery_year))
  years_valid <- years_available[years_available >= 1998 & years_available <= 2023]
  
  cat(sprintf("   Processing %d years with PM2.5 data\n\n", length(years_valid)))
  
  for (yr in sort(years_valid)) {
    pattern <- sprintf("%d01-%d12", yr, yr)
    match_file <- pm25_files[grep(pattern, pm25_files)]
    
    if (length(match_file) == 0) next
    
    cat(sprintf("   %d: ", yr))
    
    tryCatch({
      pm25_rast <- rast(match_file[1])
      
      yr_rows <- which(coords$delivery_year == yr)
      if (length(yr_rows) == 0) {
        cat("no records\n")
        next
      }
      
      yr_coords <- cbind(coords$env_longitude[yr_rows], coords$env_latitude[yr_rows])
      extracted <- terra::extract(pm25_rast, yr_coords)
      pm25_vals <- if(ncol(extracted) >= 2) extracted[[2]] else extracted[[1]]
      
      coords$env_pm25_annual[yr_rows] <- pm25_vals
      coords$env_pm25_delivery[yr_rows] <- pm25_vals
      coords$env_pm25_pregnancy[yr_rows] <- pm25_vals
      
      n_extracted <- sum(!is.na(pm25_vals))
      cat(sprintf("✅ %d values (%.1f-%.1f µg/m³)\n", 
                  n_extracted,
                  min(pm25_vals, na.rm = TRUE),
                  max(pm25_vals, na.rm = TRUE)))
      
    }, error = function(e) {
      cat(sprintf("❌ Error: %s\n", e$message))
    })
  }
  
  # Merge back
  pm25_cols <- c("unified_id", "env_pm25_annual", "env_pm25_delivery", "env_pm25_pregnancy")
  environmental_linkage <- environmental_linkage %>%
    select(-any_of(pm25_cols[pm25_cols != "unified_id"])) %>%
    left_join(coords %>% select(any_of(pm25_cols)), by = "unified_id")
  
  n_pm25 <- sum(!is.na(environmental_linkage$env_pm25_annual))
  cat(sprintf("\n   PM2.5 extracted for %s records (%.1f%%)\n",
              format(n_pm25, big.mark = ","),
              100 * n_pm25 / nrow(environmental_linkage)))
  
} else {
  cat("⚠️  No PM2.5 files found in:", DIR_PM25, "\n")
}
```

# Phase 10: Classify Seasons

```{r phase10-seasons}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                PHASE 10: SEASONAL CLASSIFICATION                           ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

environmental_linkage <- environmental_linkage %>%
  mutate(
    env_season_delivery = case_when(
      is.na(env_latitude) | is.na(delivery_month) ~ NA_character_,
      env_latitude >= 0 & delivery_month %in% c(5:10) ~ "Wet",
      env_latitude >= 0 & delivery_month %in% c(1:4, 11:12) ~ "Dry",
      env_latitude < 0 & delivery_month %in% c(11:12, 1:4) ~ "Wet",
      env_latitude < 0 & delivery_month %in% c(5:10) ~ "Dry",
      TRUE ~ NA_character_
    ),
    env_season_conception = case_when(
      is.na(env_latitude) | is.na(conception_month) ~ NA_character_,
      env_latitude >= 0 & conception_month %in% c(5:10) ~ "Wet",
      env_latitude >= 0 & conception_month %in% c(1:4, 11:12) ~ "Dry",
      env_latitude < 0 & conception_month %in% c(11:12, 1:4) ~ "Wet",
      env_latitude < 0 & conception_month %in% c(5:10) ~ "Dry",
      TRUE ~ NA_character_
    )
  )

cat(sprintf("✅ Seasons: Wet=%d, Dry=%d\n",
            sum(environmental_linkage$env_season_delivery == "Wet", na.rm = TRUE),
            sum(environmental_linkage$env_season_delivery == "Dry", na.rm = TRUE)))
```

# Phase 11: Quality Summary

```{r phase11-summary}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║              PHASE 11: QUALITY SUMMARY                                     ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

cat("=== OVERALL SUMMARY ===\n")
cat(sprintf("Total records:        %s\n", format(nrow(environmental_linkage), big.mark = ",")))
cat(sprintf("With coordinates:     %s (%.1f%%)\n", 
            format(sum(environmental_linkage$has_coordinates), big.mark = ","),
            100 * mean(environmental_linkage$has_coordinates)))
cat(sprintf("Linkage-ready:        %s (%.1f%%)\n", 
            format(sum(environmental_linkage$linkage_ready), big.mark = ","),
            100 * mean(environmental_linkage$linkage_ready)))

cat("\n=== VARIABLE COMPLETENESS ===\n")
env_vars <- c("env_elevation", "env_slope", "env_temp_mean_delivery", 
              "env_humidity_delivery", "env_precipitation_delivery", "env_heat_index_delivery",
              "env_pm25_annual", "env_season_delivery")

for (v in env_vars) {
  if (v %in% names(environmental_linkage)) {
    pct <- mean(!is.na(environmental_linkage[[v]])) * 100
    status <- if (pct > 50) "✅" else if (pct > 0) "⚠️" else "❌"
    cat(sprintf("  %s %-30s %.1f%%\n", status, v, pct))
  }
}

cat("\n=== BY STUDY ===\n")
study_summary <- environmental_linkage %>%
  group_by(study_source) %>%
  summarise(
    n = n(),
    linkage_ready = sum(linkage_ready),
    pct_ready = round(100 * mean(linkage_ready), 1),
    stillbirths = sum(out_stillbirth == "Yes", na.rm = TRUE),
    nnd = sum(out_nnd == "Yes", na.rm = TRUE),
    .groups = "drop"
  )
print(kable(study_summary))
```

# Phase 12: Export Final Dataset

```{r phase12-export}
cat("\n")
cat("╔════════════════════════════════════════════════════════════════════════════╗\n")
cat("║                   PHASE 12: EXPORT FINAL DATASET                           ║\n")
cat("╚════════════════════════════════════════════════════════════════════════════╝\n\n")

# Save as environmental_linkage_final for the appendix
environmental_linkage_final <- environmental_linkage

# RDS
saveRDS(environmental_linkage_final, file.path(DIR_OUTPUT, paste0(OUTPUT_FILE, ".rds")))
cat(sprintf("✓ Saved: %s/%s.rds\n", DIR_OUTPUT, OUTPUT_FILE))

# CSV
write_csv(environmental_linkage_final, file.path(DIR_OUTPUT, paste0(OUTPUT_FILE, ".csv")))
cat(sprintf("✓ Saved: %s/%s.csv\n", DIR_OUTPUT, OUTPUT_FILE))

# Stata
tryCatch({
  haven::write_dta(environmental_linkage_final, file.path(DIR_OUTPUT, paste0(OUTPUT_FILE, ".dta")))
  cat(sprintf("✓ Saved: %s/%s.dta\n", DIR_OUTPUT, OUTPUT_FILE))
}, error = function(e) {
  cat(sprintf("⚠️ Stata save failed: %s\n", e$message))
})

# Geocode references
write_csv(facility_geocodes, file.path(DIR_OUTPUT, "facility_geocodes_reference.csv"))
write_csv(country_centroids, file.path(DIR_OUTPUT, "country_centroids_reference.csv"))
cat("✓ Saved: geocode reference tables\n")

cat("\n")
cat("════════════════════════════════════════════════════════════════════════════\n")
cat("              🎉 PIPELINE COMPLETE! v", VERSION, " 🎉\n", sep = "")
cat("════════════════════════════════════════════════════════════════════════════\n\n")

cat(sprintf("Total Records:      %s\n", format(nrow(environmental_linkage_final), big.mark = ",")))
cat(sprintf("Linkage-Ready:      %s (%.1f%%)\n", 
            format(sum(environmental_linkage_final$linkage_ready), big.mark = ","),
            100 * mean(environmental_linkage_final$linkage_ready)))
```

---

# Appendix A: Join with Unified Dataset

```{r appendix-join, eval=FALSE}
# ============================================================================
# JOINING ENVIRONMENTAL DATA BACK TO UNIFIED DATASET
# ============================================================================

# Load datasets
unified <- readRDS("output/merged_unified_dataset_v10.16.rds")
environmental <- readRDS("output/environmental/environmental_linkage_v3.3.rds")

# Select environmental variables
env_vars_to_merge <- environmental %>%
  filter(linkage_ready) %>%
  select(
    unified_id,
    env_latitude, env_longitude, coordinate_source,
    env_elevation, env_slope,
    env_temp_mean_delivery, env_humidity_delivery, 
    env_precipitation_delivery, env_heat_index_delivery,
    env_pm25_annual, env_pm25_delivery,
    env_season_delivery, env_season_conception
  )

# Merge
unified_with_env <- unified %>%
  left_join(env_vars_to_merge, by = "unified_id")

# ---- COALESCE DUPLICATE COLUMNS (run BEFORE any saving) ----
env_cols <- names(unified_with_env)
y_cols <- grep("\\.y$", env_cols, value = TRUE)
base_names <- gsub("\\.y$", "", y_cols)

for (bn in base_names) {
  x_col <- paste0(bn, ".x")
  y_col <- paste0(bn, ".y")
  
  if (x_col %in% env_cols && y_col %in% env_cols) {
    tx <- class(unified_with_env[[x_col]])[1]
    ty <- class(unified_with_env[[y_col]])[1]
    if (tx != ty) {
      if (tx %in% c("numeric", "double", "integer") && ty == "character") {
        unified_with_env[[y_col]] <- suppressWarnings(as.numeric(unified_with_env[[y_col]]))
      } else if (ty %in% c("numeric", "double", "integer") && tx == "character") {
        unified_with_env[[x_col]] <- suppressWarnings(as.numeric(unified_with_env[[x_col]]))
      } else {
        unified_with_env[[x_col]] <- as.character(unified_with_env[[x_col]])
        unified_with_env[[y_col]] <- as.character(unified_with_env[[y_col]])
      }
    }
    unified_with_env[[x_col]] <- dplyr::coalesce(
      unified_with_env[[x_col]], 
      unified_with_env[[y_col]]
    )
    unified_with_env[[y_col]] <- NULL
    names(unified_with_env)[names(unified_with_env) == x_col] <- bn
    cat(glue("  Coalesced {bn}: {sum(!is.na(unified_with_env[[bn]]))} non-NA values\n"))
  }
}

# ---- SAVE RDS (clean columns, dots OK) ----
saveRDS(unified_with_env, "output/unified_dataset_with_env_v10.16.rds")

# ---- SAVE CSV ----
write_csv(unified_with_env, "output/unified_dataset_with_env_v10.16.csv")

# ---- SAVE STATA (additional name cleaning needed) ----
stata_out <- unified_with_env
names(stata_out) <- gsub("\\.", "_", names(stata_out))
names(stata_out) <- make.unique(names(stata_out), sep = "_")
names(stata_out)[nchar(names(stata_out)) > 32] <- substr(
  names(stata_out)[nchar(names(stata_out)) > 32], 1, 32
)
names(stata_out) <- make.unique(names(stata_out), sep = "_")
haven::write_dta(stata_out, "output/unified_dataset_with_env_v10.16.dta")
rm(stata_out)


cat(sprintf("✅ Merged: %d records with environmental data\n", 
            sum(!is.na(unified_with_env$env_latitude))))
cat("✅ Saved: .rds, .csv, and .dta formats\n")
```

---

# Appendix B: Download SRTM (Optional)

```{r optional-srtm, eval=FALSE}
if (!require(elevatr)) install.packages("elevatr")
library(elevatr)

bbox_coords <- matrix(c(-20, -35, 55, -35, 55, 40, -20, 40, -20, -35), ncol = 2, byrow = TRUE)
bbox_poly <- st_polygon(list(bbox_coords)) %>% st_sfc(crs = 4326) %>% st_as_sf()

elev_raster <- get_elev_raster(bbox_poly, z = 5, src = "aws")
elev_terra <- rast(elev_raster)

dir.create("data/srtm", recursive = TRUE, showWarnings = FALSE)
writeRaster(elev_terra, "data/srtm/srtm_elevation.tif", overwrite = TRUE)

slope_terra <- terrain(elev_terra, v = "slope", unit = "degrees")
writeRaster(slope_terra, "data/srtm/srtm_slope.tif", overwrite = TRUE)

cat("✅ SRTM download complete!\n")
```

---

# Appendix C: Extract NDVI (Optional)

```{r optional-ndvi, eval=FALSE}
env_data <- readRDS(file.path(DIR_OUTPUT, paste0(OUTPUT_FILE, ".rds")))

ndvi_files <- list.files(DIR_NDVI, pattern = "\\.(tif|nc)$", full.names = TRUE, recursive = TRUE)

if (length(ndvi_files) > 0) {
  valid_idx <- which(!is.na(env_data$env_latitude) & !is.na(env_data$env_longitude))
  
  ndvi_stack <- terra::rast(ndvi_files)
  ndvi_mean <- terra::mean(ndvi_stack, na.rm = TRUE)
  
  coords <- cbind(env_data$env_longitude[valid_idx], env_data$env_latitude[valid_idx])
  ndvi_vals <- terra::extract(ndvi_mean, coords)[[2]]
  
  if (max(abs(ndvi_vals), na.rm = TRUE) > 10) ndvi_vals <- ndvi_vals * 0.0001
  
  env_data$env_ndvi_delivery[valid_idx] <- ndvi_vals
  env_data$env_ndvi_pregnancy[valid_idx] <- ndvi_vals
  
  saveRDS(env_data, file.path(DIR_OUTPUT, paste0(OUTPUT_FILE, ".rds")))
  cat("✅ NDVI extracted and saved\n")
}
```

---

# Session Info

```{r session-info}
cat("Pipeline completed:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
cat("R version:", R.version.string, "\n")
```
