---
title: "Predicting Stillbirth & Neonatal Death in Sub-Saharan Africa"
subtitle: "Comprehensive Descriptive Analysis & Pre-Modeling Report"
author: "Audencio Victor | LSHTM"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: kate
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    df_print: paged
    fig_width: 11
    fig_height: 7
    css: !expr NULL
params:
  data_path: "unified_dataset_with_env_v10.17_cleaned.rds"
  dataset_dir: "../Dataset"
  output_dir: "../outputs_descriptive_report"
  global_threshold: 70
  type_threshold: 60
  study_threshold: 60
---

```{css custom-styles, echo=FALSE}
/* ══════════════════════════════════════════════════════════════
   LSHTM Editorial Theme — Inspired by academic journal layouts
   Clean, professional, no visual clutter
   ══════════════════════════════════════════════════════════════ */

@import url('https://fonts.googleapis.com/css2?family=IBM+Plex+Sans:wght@300;400;500;600;700&family=IBM+Plex+Mono:wght@400;500&display=swap');

:root {
  --ink: #1a1a2e;
  --muted: #4a4a6a;
  --rule: #d0d0e8;
  --accent1: #c0392b;
  --accent2: #2c7fb8;
  --accent3: #16a085;
  --bg: #fafaf8;
  --paper: #ffffff;
  --highlight: #f7f3e9;
}

/* ── Global ── */
body {
  font-family: 'IBM Plex Sans', system-ui, -apple-system, sans-serif;
  background-color: var(--bg);
  color: var(--ink);
  line-height: 1.7;
  font-size: 14px;
}

/* ── Headers — editorial style ── */
h1, h2, h3, h4 {
  font-family: 'IBM Plex Sans', sans-serif;
  color: var(--ink);
  font-weight: 700;
  letter-spacing: -0.01em;
  margin-top: 48px;
}

h1 {
  font-size: 1.6em;
  border-bottom: 3px solid var(--ink);
  padding-bottom: 10px;
}

h2 {
  font-size: 0.95em;
  border-bottom: none;
  border-left: 4px solid var(--accent2);
  padding-left: 12px;
  padding-bottom: 0;
  color: var(--muted);
  text-transform: uppercase;
  letter-spacing: 0.05em;
  font-weight: 600;
}

h3 {
  font-size: 1.05em;
  border-bottom: 1px solid var(--rule);
  padding-bottom: 6px;
}

/* ── KPI Cards — clean bordered style ── */
.metric-card {
  background: var(--paper);
  border: 1px solid var(--rule);
  color: var(--ink);
  padding: 18px 22px;
  border-radius: 0;
  margin: 6px;
  display: inline-block;
  min-width: 180px;
  text-align: center;
  position: relative;
  overflow: hidden;
}

.metric-card::before {
  content: '';
  position: absolute;
  top: 0; left: 0; right: 0;
  height: 3px;
  background: var(--accent2);
}

.metric-card .metric-value {
  font-family: 'IBM Plex Sans', sans-serif;
  font-size: 2em;
  font-weight: 700;
  display: block;
  margin-bottom: 4px;
  color: var(--ink);
}

.metric-card .metric-label {
  font-size: 0.75em;
  color: var(--muted);
  text-transform: uppercase;
  letter-spacing: 0.1em;
}

.card-blue::before { background: var(--accent2); }
.card-green::before { background: var(--accent3); }
.card-orange::before { background: var(--accent1); }
.card-purple::before { background: #7c4dff; }
.card-teal::before { background: #00838f; }

/* ── Info boxes — editorial callouts ── */
.info-box {
  background: var(--highlight);
  border-left: 4px solid var(--accent2);
  padding: 16px 20px;
  border-radius: 0;
  margin: 24px 0;
  font-size: 0.92em;
  color: var(--ink);
}

.warning-box {
  background: #fef9f9;
  border-left: 4px solid var(--accent1);
  padding: 16px 20px;
  border-radius: 0;
  margin: 24px 0;
  font-size: 0.92em;
}

.success-box {
  background: #f0f8f4;
  border-left: 4px solid var(--accent3);
  padding: 16px 20px;
  border-radius: 0;
  margin: 24px 0;
  font-size: 0.92em;
}

/* ── Section dividers ── */
.section-divider {
  height: 2px;
  background: var(--ink);
  border: none;
  border-radius: 0;
  margin: 48px 0;
}

/* ── Tables — dark header, clean rows ── */
.gt_table {
  border-radius: 0 !important;
  overflow: hidden;
  box-shadow: none;
  border: 1px solid var(--rule);
}

/* ── Tabs ── */
.nav-tabs > li > a {
  color: var(--ink);
  font-weight: 600;
  font-size: 0.9em;
  letter-spacing: 0.03em;
}

.nav-tabs > li.active > a {
  border-bottom: 3px solid var(--accent1) !important;
}

.tab-content {
  padding: 24px 0;
}

/* ── TOC — minimal sidebar ── */
.tocify {
  border: 1px solid var(--rule) !important;
  border-radius: 0;
  box-shadow: none;
  background: var(--paper);
}

.tocify-header {
  font-weight: 600;
  font-size: 0.85em;
}

.tocify a {
  color: var(--muted);
}

.tocify .active a {
  color: var(--ink);
  border-left: 3px solid var(--accent1);
}

/* ── Figure containers ── */
.figure-panel {
  background: var(--paper);
  border: 1px solid var(--rule);
  padding: 16px;
  margin: 16px 0;
}

/* ── Print ── */
@media print {
  body { background: white; font-size: 11px; }
  .section-divider { background: #000; height: 1px; }
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "center",
  fig.retina = 2,
  dpi        = 150,
  cache      = FALSE
)
```

```{r packages}
# ── Package management ──
required_pkgs <- c(
  "dplyr", "tidyr", "stringr", "forcats", "purrr", "tibble", "readr",
  "ggplot2", "scales", "plotly", "patchwork", "viridis", "RColorBrewer",
  "gtsummary", "gt", "DT", "kableExtra",
  "janitor", "pheatmap", "corrplot", "ggridges",
  "officer", "flextable", "openxlsx",
  "sf", "rnaturalearth", "rnaturalearthdata",
  "htmltools"
)

for (p in required_pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p, repos = "https://cloud.r-project.org", quiet = TRUE)
  }
}

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(forcats)
  library(purrr)
  library(tibble)
  library(ggplot2)
  library(scales)
  library(plotly)
  library(patchwork)
  library(viridis)
  library(RColorBrewer)
  library(gtsummary)
  library(gt)
  library(DT)
  library(kableExtra)
  library(janitor)
  library(pheatmap)
  library(corrplot)
  library(ggridges)
  library(officer)
  library(flextable)
  library(openxlsx)
  library(htmltools)
})
```

```{r theme-setup}
# ── Custom ggplot2 theme — clean, no grid lines ──
theme_lshtm <- function(base_size = 13) {
  theme_minimal(base_size = base_size, base_family = "sans") +
    theme(
      plot.title       = element_text(face = "bold", size = rel(1.15), color = "#1a1a2e",
                                       margin = margin(b = 8)),
      plot.subtitle    = element_text(size = rel(0.85), color = "#4a4a6a",
                                       margin = margin(b = 12)),
      plot.caption     = element_text(size = rel(0.7), color = "#8e99a4", hjust = 0),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line        = element_line(color = "#d0d0e8", linewidth = 0.4),
      axis.title       = element_text(size = rel(0.85), color = "#4a4a6a", face = "plain"),
      axis.text        = element_text(size = rel(0.8), color = "#4a4a6a"),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold", size = rel(0.8)),
      legend.text      = element_text(size = rel(0.75)),
      strip.text       = element_text(face = "bold", size = rel(0.85), color = "#1a1a2e"),
      strip.background = element_rect(fill = "#f0f0f8", color = NA),
      plot.margin      = margin(12, 12, 12, 12),
      plot.background  = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA)
    )
}

theme_set(theme_lshtm())

# Palettes
pal_main     <- c("#2c7fb8", "#253494", "#41b6c4", "#7fcdbb", "#c7e9b4",
                  "#ffffcc", "#d73027", "#fc8d59", "#fee090", "#91bfdb")
pal_outcomes <- c("Stillbirth" = "#d73027", "Neonatal Death" = "#fc8d59",
                  "Preterm" = "#fee090", "LBW" = "#91bfdb", "SGA" = "#4575b4")
pal_study    <- scales::hue_pal()(10)

# Helper functions
safe_num <- function(x) suppressWarnings(as.numeric(as.character(x)))

to01 <- function(x) {
  if (is.null(x)) return(rep(NA_integer_, 0))
  if (is.logical(x)) return(ifelse(is.na(x), NA_integer_, as.integer(x)))
  if (is.numeric(x)) {
    ux <- sort(unique(x[!is.na(x)]))
    if (length(ux) == 0) return(as.integer(x))
    if (all(ux %in% c(0, 1))) return(as.integer(x))
    if (all(ux %in% c(1, 2))) return(ifelse(is.na(x), NA_integer_, as.integer(x == 2)))
    return(ifelse(is.na(x), NA_integer_, as.integer(x > 0)))
  }
  z <- tolower(trimws(as.character(x)))
  z[z %in% c("", "na", "n/a", "missing", "unknown")] <- NA
  out <- rep(NA_integer_, length(z))
  out[z %in% c("yes", "y", "true", "t", "1", "sim", "s")] <- 1L
  out[z %in% c("no", "n", "false", "f", "0", "nao", "nao")] <- 0L
  out
}

rate_1000 <- function(events, denom) {
  ifelse(denom > 0, round(events / denom * 1000, 1), NA_real_)
}

fmt_n <- function(x) format(x, big.mark = ",", scientific = FALSE)

# ── Reusable gt theme ──
gt_lshtm <- function(data, title = NULL, subtitle = NULL) {
  out <- data
  if (!is.null(title)) {
    out <- out %>% tab_header(
      title = md(paste0("**", title, "**")),
      subtitle = if (!is.null(subtitle)) md(paste0("*", subtitle, "*")) else NULL
    )
  }
  out %>%
    tab_style(
      style = cell_text(color = "white", weight = "bold"),
      locations = cells_column_labels()
    ) %>%
    tab_options(
      table.font.size        = px(13),
      heading.background.color     = "#1a3a5c",
      heading.title.font.size      = px(16),
      column_labels.background.color = "#2c7fb8",
      column_labels.font.weight    = "bold",
      row_group.background.color   = "#eef5fc",
      table.border.top.color       = "#1a3a5c",
      table.border.bottom.color    = "#1a3a5c"
    )
}
```

```{r load-data}
# ── Load dataset ──
data_file <- file.path(params$dataset_dir, params$data_path)
if (!file.exists(data_file)) {
  data_file <- params$data_path
}

df_raw <- readRDS(data_file)
df_raw <- janitor::clean_names(df_raw)

# ── Country standardisation ──
country_var <- intersect(c("mat_country", "loc_country", "country"), names(df_raw))[1]

df_raw[[country_var]][df_raw[[country_var]] %in%
  c("Dem Rep Of The Congo", "Congo Democratic Republic")] <- "Democratic Republic of the Congo"
df_raw[[country_var]][df_raw[[country_var]] == "Congo"]   <- "Republic of the Congo"
df_raw[[country_var]][df_raw[[country_var]] == "Gambia"]   <- "The Gambia"
df_raw[[country_var]][df_raw[[country_var]] == "Swaziland"] <- "Eswatini"

# ── String cleaning ──
for (j in names(df_raw)) {
  if (is.character(df_raw[[j]])) {
    x <- trimws(df_raw[[j]])
    x[x %in% c("", "NA", "N/A", "na", "n/a", "missing", "unknown", "Unknown", "UNK")] <- NA
    df_raw[[j]] <- x
  }
}

# ── SSA + >= 2010 filter ──
ssa_countries <- c(
  "Uganda", "Benin", "Malawi", "Tanzania", "Mozambique", "Kenya", "The Gambia",
  "Guinea-Bissau", "Ethiopia", "Ghana", "Angola", "Democratic Republic of the Congo",
  "Republic of the Congo", "Niger", "Nigeria", "Burkina Faso", "Burundi", "Cameroon",
  "Central African Republic", "Chad", "Comoros", "Cote d'Ivoire", "Eswatini", "Gabon",
  "Guinea", "Lesotho", "Liberia", "Madagascar", "Mali", "Mauritania", "Namibia",
  "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "South Africa",
  "Sudan", "Togo", "Zambia", "Zimbabwe", "Botswana", "Cabo Verde", "Djibouti",
  "Equatorial Guinea", "Eritrea", "Somalia", "South Sudan"
)

year_var <- intersect(c("studyyear", "study_year", "year"), names(df_raw))[1]

df <- df_raw %>%
  filter(
    !is.na(.data[[year_var]]),
    safe_num(.data[[year_var]]) >= 2010,
    !is.na(.data[[country_var]]),
    .data[[country_var]] %in% ssa_countries
  )

# ── Column identification ──
study_var  <- intersect(c("study_source", "study", "source"), names(df))[1]
site_var   <- intersect(c("site", "site_id", "facility", "loc_site"), names(df))[1]
sex_var    <- intersect(c("neo_sex", "sex", "child_sex"), names(df))[1]
lat_var    <- intersect(c("env_latitude", "lat", "latitude"), names(df))[1]
lon_var    <- intersect(c("env_longitude", "lon", "longitude"), names(df))[1]

# Outcomes
V_SB   <- intersect(c("out_stillbirth", "stillbirth"), names(df))[1]
V_LB   <- intersect(c("out_livebirth", "livebirth"), names(df))[1]
V_NND  <- intersect(c("out_nnd", "nnd", "out_neonatal_death"), names(df))[1]
V_PND  <- intersect(c("out_perinatal_death", "perinatal_death"), names(df))[1]
V_PRE  <- intersect(c("out_preterm", "preterm"), names(df))[1]
V_LBW  <- intersect(c("out_lbw", "lbw"), names(df))[1]
V_VLBW <- intersect(c("out_vlbw", "vlbw"), names(df))[1]
V_SGA  <- intersect(c("out_sga", "sga"), names(df))[1]

# Continuous
V_GA   <- intersect(c("out_ga_weeks", "ga_weeks"), names(df))[1]
V_BW   <- intersect(c("out_birthweight_g", "birthweight_g"), names(df))[1]
V_AGE  <- intersect(c("mat_age", "maternal_age"), names(df))[1]
V_EDU  <- intersect(c("mat_education", "education"), names(df))[1]
V_MODE <- intersect(c("delivery_mode", "mat_delivery_mode"), names(df))[1]
V_MAR  <- intersect(c("mat_marital_status", "marital_status"), names(df))[1]
V_MULT <- intersect(c("multiple_birth", "twin", "plurality"), names(df))[1]
V_SBP  <- intersect(c("mat_sbp", "sbp"), names(df))[1]
V_DBP  <- intersect(c("mat_dbp", "dbp"), names(df))[1]
V_AP1  <- intersect(c("neo_apgar_1", "apgar1"), names(df))[1]
V_AP5  <- intersect(c("neo_apgar_5", "apgar5"), names(df))[1]
V_EARLY <- intersect(c("out_early_neonatal_death", "early_nnd"), names(df))[1]
V_LATE  <- intersect(c("out_late_neonatal_death", "late_nnd"), names(df))[1]
V_AGEDEATH <- intersect(c("neo_age_at_death_days", "age_at_death_days"), names(df))[1]

# Environmental
env_vars <- grep("^env_", names(df), value = TRUE)
env_num  <- env_vars[sapply(df[env_vars], is.numeric)]

# Study type classification
df <- df %>%
  mutate(
    study_type = case_when(
      str_detect(tolower(.data[[study_var]]), "dhs")                     ~ "Survey / DHS",
      str_detect(tolower(.data[[study_var]]), "who|whomcs")              ~ "Facility / WHO",
      str_detect(tolower(.data[[study_var]]), "alert|precise|ptbi|ncops|indepth|amanhi|interbio|biob") ~ "Cohort / Clinical",
      str_detect(tolower(.data[[study_var]]), "registry|surveillance")   ~ "Registry / Surveillance",
      TRUE ~ "Other"
    )
  )

# Key counts
N_TOTAL     <- nrow(df)
N_VARS      <- ncol(df)
N_COUNTRIES <- n_distinct(df[[country_var]])
N_STUDIES   <- if (!is.na(study_var)) n_distinct(df[[study_var]]) else NA

# Output directory
dir.create(params$output_dir, showWarnings = FALSE, recursive = TRUE)
```

<hr class="section-divider">

# Executive Summary {.tabset .tabset-fade}

## Key Metrics

```{r exec-cards, results='asis'}
yy <- safe_num(df[[year_var]])
yr_range <- paste0(min(yy, na.rm = TRUE), " - ", max(yy, na.rm = TRUE))

sb_n <- if (!is.na(V_SB)) sum(to01(df[[V_SB]]) == 1, na.rm = TRUE) else NA
lb_n <- if (!is.na(V_LB)) sum(to01(df[[V_LB]]) == 1, na.rm = TRUE) else NA
nnd_n <- if (!is.na(V_NND)) sum(to01(df[[V_NND]]) == 1, na.rm = TRUE) else NA

tb_n <- NA
if (!is.na(V_SB) && !is.na(V_LB)) {
  sb01 <- to01(df[[V_SB]]); lb01 <- to01(df[[V_LB]])
  tb_n <- sum((sb01 == 1) | (lb01 == 1), na.rm = TRUE)
}

sbr <- if (!is.na(sb_n) && !is.na(tb_n) && tb_n > 0) rate_1000(sb_n, tb_n) else NA
nnd_rate <- if (!is.na(nnd_n) && !is.na(lb_n) && lb_n > 0) rate_1000(nnd_n, lb_n) else NA
pmr <- if (!is.na(sb_n) && !is.na(nnd_n) && !is.na(tb_n) && tb_n > 0) rate_1000(sb_n + nnd_n, tb_n) else NA

cat('<div style="text-align:center; margin: 30px 0;">\n')
cat(sprintf('<div class="metric-card card-blue"><span class="metric-value">%s</span><span class="metric-label">Total Records</span></div>\n', fmt_n(N_TOTAL)))
cat(sprintf('<div class="metric-card card-green"><span class="metric-value">%s</span><span class="metric-label">Variables</span></div>\n', fmt_n(N_VARS)))
cat(sprintf('<div class="metric-card card-purple"><span class="metric-value">%s</span><span class="metric-label">Countries</span></div>\n', N_COUNTRIES))
cat(sprintf('<div class="metric-card"><span class="metric-value">%s</span><span class="metric-label">Studies</span></div>\n', N_STUDIES))
cat(sprintf('<div class="metric-card card-teal"><span class="metric-value">%s</span><span class="metric-label">Year Range</span></div>\n', yr_range))
cat('</div>\n')

cat('<div style="text-align:center; margin: 20px 0;">\n')
if (!is.na(sbr)) cat(sprintf('<div class="metric-card card-orange"><span class="metric-value">%s</span><span class="metric-label">SBR per 1,000</span></div>\n', sbr))
if (!is.na(nnd_rate)) cat(sprintf('<div class="metric-card card-orange"><span class="metric-value">%s</span><span class="metric-label">NND rate per 1,000 LB</span></div>\n', nnd_rate))
if (!is.na(pmr)) cat(sprintf('<div class="metric-card card-orange"><span class="metric-value">%s</span><span class="metric-label">PMR per 1,000</span></div>\n', pmr))
cat('</div>\n')
```

<div class="info-box">
**Data source:** Unified dataset v10.17 (cleaned) filtered for Sub-Saharan Africa and studies from 2010 onwards. Outcomes use correct epidemiological denominators: SBR = SB/Total Births, NND rate = NND/Live Births, PMR = (SB+NND)/Total Births.
</div>

## Rate Verification

```{r rate-verification}
if (!is.na(V_SB) && !is.na(V_LB)) {
  rate_tbl <- tibble(
    Metric = c("Stillbirths (SB)", "Live Births (LB)", "Total Births (TB)",
               "Neonatal Deaths (NND)", "SBR per 1,000 (SB/TB)",
               "NND rate per 1,000 (NND/LB)", "PMR per 1,000 ((SB+NND)/TB)"),
    Value = c(
      fmt_n(sb_n), fmt_n(lb_n), fmt_n(tb_n),
      if (!is.na(nnd_n)) fmt_n(nnd_n) else "N/A",
      if (!is.na(sbr)) as.character(sbr) else "N/A",
      if (!is.na(nnd_rate)) as.character(nnd_rate) else "N/A",
      if (!is.na(pmr)) as.character(pmr) else "N/A"
    )
  )

  rate_tbl %>%
    gt() %>%
    cols_label(Metric = "Indicator", Value = "Value") %>%
    tab_style(style = cell_fill(color = "#fef3f0"), locations = cells_body(rows = c(5, 6, 7))) %>%
    tab_style(style = cell_text(weight = "bold"), locations = cells_body(columns = Value)) %>%
    gt_lshtm(title = "Outcome Rate Verification")
}
```

<hr class="section-divider">

# Study Overview {.tabset .tabset-fade}

## Study Characteristics

```{r study-overview-table}
if (!is.na(study_var)) {
  study_overview <- df %>%
    group_by(study = .data[[study_var]]) %>%
    summarise(
      Type        = first(study_type),
      N_records   = n(),
      N_countries = n_distinct(.data[[country_var]]),
      Year_min    = min(safe_num(.data[[year_var]]), na.rm = TRUE),
      Year_max    = max(safe_num(.data[[year_var]]), na.rm = TRUE),
      SB_avail    = if (!is.na(V_SB)) round(100 * mean(!is.na(.data[[V_SB]])), 1) else NA_real_,
      NND_avail   = if (!is.na(V_NND)) round(100 * mean(!is.na(.data[[V_NND]])), 1) else NA_real_,
      .groups = "drop"
    ) %>%
    mutate(
      Pct_total  = round(100 * N_records / sum(N_records), 1),
      Year_range = paste0(Year_min, "-", Year_max)
    ) %>%
    arrange(desc(N_records))

  study_overview %>%
    select(study, Type, N_records, N_countries, Year_range, Pct_total, SB_avail, NND_avail) %>%
    gt() %>%
    cols_label(
      study = "Study", Type = "Type", N_records = "N Records",
      N_countries = "Countries", Year_range = "Years",
      Pct_total = "% Total", SB_avail = "SB Avail %", NND_avail = "NND Avail %"
    ) %>%
    fmt_number(columns = N_records, decimals = 0, use_seps = TRUE) %>%
    data_color(columns = Pct_total, palette = c("#e8f5e9", "#c8e6c9", "#81c784"), domain = c(0, 100)) %>%
    data_color(columns = c(SB_avail, NND_avail), palette = c("#fce4ec", "#fff8e1", "#e8f5e9"), domain = c(0, 100)) %>%
    gt_lshtm(title = "Study Overview", subtitle = "SSA, 2010 onwards")
}
```

## Records per Study & Type

```{r study-overview-viz, fig.height=8, fig.width=12}
if (!is.na(study_var) && exists("study_overview")) {
  so <- study_overview %>%
    mutate(study = factor(study, levels = study[order(N_records)]))

  # Panel A: N per study (horizontal bar)
  p_n <- ggplot(so, aes(x = study, y = N_records, fill = Type)) +
    geom_col(alpha = 0.9, width = 0.7) +
    geom_text(aes(label = paste0(fmt_n(N_records), " (", Pct_total, "%)")),
              hjust = -0.05, size = 3, color = "#1a1a2e") +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.25))) +
    coord_flip() +
    labs(title = "A. Records per Study", x = NULL, y = "N Records", fill = "Study Type")

  # Panel B: Stacked proportion bar
  p_prop <- ggplot(so, aes(x = "", y = N_records, fill = study)) +
    geom_col(width = 0.8, color = "white", linewidth = 0.5) +
    scale_fill_viridis_d(option = "D") +
    coord_flip() +
    scale_y_continuous(labels = comma) +
    labs(title = "B. Proportional Contribution", x = NULL, y = "N Records", fill = "Study") +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          axis.line.y = element_blank())

  # Panel C: N per study type
  type_n <- study_overview %>%
    group_by(Type) %>%
    summarise(N = sum(N_records), Studies = n(), .groups = "drop") %>%
    mutate(Type = factor(Type, levels = Type[order(N)]))

  p_type <- ggplot(type_n, aes(x = Type, y = N, fill = Type)) +
    geom_col(alpha = 0.9, width = 0.6, show.legend = FALSE) +
    geom_text(aes(label = paste0(fmt_n(N), "\n(", Studies, " studies)")),
              hjust = -0.05, size = 2.8, color = "#1a1a2e") +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.35))) +
    coord_flip() +
    labs(title = "C. Records per Study Type", x = NULL, y = "N Records")

  (p_n / (p_prop | p_type)) +
    plot_layout(heights = c(2, 1)) +
    plot_annotation(
      title = "Study Sample Size Overview",
      subtitle = paste0("Total: ", fmt_n(sum(so$N_records)), " records across ", nrow(so), " studies"),
      theme = theme(
        plot.title = element_text(face = "bold", size = 16, color = "#1a1a2e"),
        plot.subtitle = element_text(size = 11, color = "#4a4a6a")
      )
    )
}
```

<hr class="section-divider">

# Study & Geographic Distribution {.tabset .tabset-fade}

## By Study Source

```{r study-distribution-plotly}
if (!is.na(study_var)) {
  study_dist <- df %>%
    count(.data[[study_var]], name = "n") %>%
    mutate(pct = round(100 * n / sum(n), 1)) %>%
    arrange(n) %>%
    mutate(study = factor(.data[[study_var]], levels = .data[[study_var]]))

  p <- ggplot(study_dist, aes(x = study, y = n, fill = n,
                               text = paste0(.data[[study_var]], "\nN = ", fmt_n(n), " (", pct, "%)"))) +
    geom_col(show.legend = FALSE) +
    scale_fill_viridis_c(option = "D", direction = -1) +
    coord_flip() +
    scale_y_continuous(labels = comma) +
    labs(title = "Records by Study Source", x = NULL, y = "Number of Records",
         caption = "Source: Unified dataset v10.17") +
    theme(plot.title = element_text(size = 16))

  ggplotly(p, tooltip = "text") %>%
    layout(margin = list(l = 150))
}
```

## By Study Type

```{r study-type-dist}
type_dist <- df %>%
  count(study_type, name = "n") %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  arrange(desc(n))

p <- ggplot(type_dist, aes(x = reorder(study_type, n), y = n, fill = study_type,
                            text = paste0(study_type, "\nN = ", fmt_n(n), " (", pct, "%)"))) +
  geom_col(show.legend = FALSE, alpha = 0.9) +
  scale_fill_brewer(palette = "Set2") +
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(title = "Records by Study Type", x = NULL, y = "Number of Records")

ggplotly(p, tooltip = "text")
```

## Geographic Distribution

```{r geographic-distribution-composed, fig.height=10, fig.width=14}
if (!is.na(country_var)) {
  ctry <- df %>%
    count(.data[[country_var]], name = "n") %>%
    rename(country = 1) %>%
    mutate(pct = round(100 * n / sum(n), 1))

  # Panel A: Africa map
  p_map <- NULL
  if (requireNamespace("rnaturalearth", quietly = TRUE) &&
      requireNamespace("sf", quietly = TRUE)) {
    africa <- rnaturalearth::ne_countries(
      continent = "Africa", returnclass = "sf"
    )
    africa_map <- africa %>%
      left_join(ctry, by = c("name" = "country")) %>%
      mutate(n = ifelse(is.na(n) | n == 0, NA_real_, n))

    p_map <- ggplot(africa_map) +
      geom_sf(aes(fill = log1p(n)),
              color = "grey70", linewidth = 0.3) +
      scale_fill_viridis_c(
        option = "C",
        name = "Records\n(log scale)",
        labels = function(x) fmt_n(round(expm1(x))),
        na.value = "white"
      ) +
      coord_sf(xlim = c(-20, 55), ylim = c(-40, 40)) +
      labs(title = "A. Geographic Coverage") +
      theme_void(base_size = 12) +
      theme(
        plot.title = element_text(
          face = "bold", color = "#1a1a2e", size = 13
        ),
        legend.position = "bottom",
        legend.key.width = unit(1.5, "cm"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 7)
      )
  }

  # Panel B: Top 20 countries bar chart
  ctry_top <- ctry %>%
    arrange(desc(n)) %>%
    slice_head(n = 20) %>%
    mutate(country = factor(country, levels = rev(country)))

  p_bar <- ggplot(ctry_top,
    aes(x = country, y = n, fill = n)) +
    geom_col(show.legend = FALSE, alpha = 0.9) +
    geom_text(
      aes(label = fmt_n(n)),
      hjust = -0.05, size = 2.8, color = "#1a1a2e"
    ) +
    scale_fill_viridis_c(option = "C", direction = -1) +
    coord_flip() +
    scale_y_continuous(
      labels = comma,
      expand = expansion(mult = c(0, 0.2))
    ) +
    labs(
      title = "B. Top 20 Countries by Records",
      x = NULL, y = "N Records"
    )

  # Compose: map + bar side by side
  if (!is.null(p_map)) {
    p_map + p_bar +
      plot_layout(widths = c(1.2, 1)) +
      plot_annotation(
        title = "Geographic Distribution of Records",
        subtitle = paste0(
          n_distinct(ctry$country), " countries | ",
          fmt_n(sum(ctry$n)), " total records"
        ),
        theme = theme(
          plot.title = element_text(
            face = "bold", size = 15, color = "#1a1a2e"
          ),
          plot.subtitle = element_text(
            size = 10, color = "#4a4a6a"
          )
        )
      )
  } else {
    p_bar
  }
}
```

## Temporal Trends

```{r temporal-trends-composed, fig.height=8, fig.width=12}
if (!is.na(year_var)) {
  yr_data <- df %>%
    mutate(year = safe_num(.data[[year_var]])) %>%
    filter(!is.na(year))

  yr_overall <- yr_data %>% count(year, name = "n")
  yr_by_type <- yr_data %>% count(year, study_type, name = "n")

  p1 <- ggplot(yr_overall, aes(x = year, y = n)) +
    geom_area(fill = "#2c7fb8", alpha = 0.15) +
    geom_line(color = "#2c7fb8", linewidth = 1) +
    geom_point(color = "#1a1a2e", size = 2) +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(title = "A. Overall Records Over Time",
         x = "Year", y = "Records")

  p2 <- ggplot(yr_by_type, aes(x = year, y = n,
                                 fill = study_type)) +
    geom_area(alpha = 0.7) +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(labels = comma) +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(title = "B. Records by Study Type",
         x = "Year", y = "Records", fill = "Type")

  p1 + p2 +
    plot_layout(ncol = 2) +
    plot_annotation(
      title = "Temporal Distribution of Records",
      theme = theme(
        plot.title = element_text(
          face = "bold", size = 15, color = "#1a1a2e"
        )
      )
    )
}
```

<hr class="section-divider">

# Primary Outcomes Analysis {.tabset .tabset-fade}

## Outcome Rates by Study

```{r outcomes-by-study-composed, fig.height=8, fig.width=14}
sb01 <- if (!is.na(V_SB)) to01(df[[V_SB]]) else NULL
lb01 <- if (!is.na(V_LB)) to01(df[[V_LB]]) else NULL
nnd01 <- if (!is.na(V_NND)) to01(df[[V_NND]]) else NULL

plots_study <- list()

if (!is.null(sb01) && !is.null(lb01) && !is.na(study_var)) {
  tb01 <- (sb01 == 1) | (lb01 == 1)
  sb_by_study <- df %>%
    mutate(.sb = sb01, .tb = tb01) %>%
    group_by(study = .data[[study_var]]) %>%
    summarise(
      SB = sum(.sb == 1, na.rm = TRUE),
      TB = sum(.tb, na.rm = TRUE),
      SBR = rate_1000(SB, TB), .groups = "drop"
    ) %>% filter(TB > 0) %>% arrange(SBR)

  plots_study$sb <- ggplot(
    sb_by_study,
    aes(x = reorder(study, SBR), y = SBR, fill = SBR)
  ) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(
      low = "#fee090", high = "#d73027"
    ) +
    coord_flip() +
    labs(title = "A. Stillbirth Rate by Study",
         x = NULL, y = "SBR per 1,000")
}

if (!is.null(nnd01) && !is.null(lb01) && !is.na(study_var)) {
  nnd_by_study <- df %>%
    mutate(.nnd = nnd01, .lb = lb01) %>%
    group_by(study = .data[[study_var]]) %>%
    summarise(
      NND = sum(.nnd == 1, na.rm = TRUE),
      LB = sum(.lb == 1, na.rm = TRUE),
      NND_rate = rate_1000(NND, LB), .groups = "drop"
    ) %>% filter(LB > 0) %>% arrange(NND_rate)

  plots_study$nnd <- ggplot(
    nnd_by_study,
    aes(x = reorder(study, NND_rate), y = NND_rate,
        fill = NND_rate)
  ) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(
      low = "#e0f3f8", high = "#4575b4"
    ) +
    coord_flip() +
    labs(title = "B. Neonatal Death Rate by Study",
         x = NULL, y = "NND per 1,000 LB")
}

if (length(plots_study) == 2) {
  plots_study$sb + plots_study$nnd +
    plot_annotation(
      title = "Perinatal Outcome Rates by Study",
      theme = theme(
        plot.title = element_text(
          face = "bold", size = 15, color = "#1a1a2e"
        )
      )
    )
} else if (length(plots_study) == 1) {
  plots_study[[1]]
}
```

## Outcome Rates by Country

```{r outcomes-by-country-composed, fig.height=10, fig.width=14}
plots_cty <- list()

if (!is.null(sb01) && !is.null(lb01) && !is.na(country_var)) {
  tb01 <- (sb01 == 1) | (lb01 == 1)
  sb_by_cty <- df %>%
    mutate(.sb = sb01, .tb = tb01) %>%
    group_by(country = .data[[country_var]]) %>%
    summarise(
      SB = sum(.sb == 1, na.rm = TRUE),
      TB = sum(.tb, na.rm = TRUE),
      SBR = rate_1000(SB, TB), .groups = "drop"
    ) %>% filter(TB >= 50) %>% arrange(SBR)

  plots_cty$sb <- ggplot(sb_by_cty,
    aes(x = reorder(country, SBR), y = SBR)) +
    geom_segment(aes(xend = country, y = 0, yend = SBR),
                 linewidth = 0.6, color = "#fc8d59") +
    geom_point(aes(size = TB),
               color = "#d73027", alpha = 0.8) +
    scale_size_continuous(
      range = c(1.5, 6), labels = comma,
      name = "Total Births"
    ) +
    coord_flip() +
    labs(title = "A. SBR by Country",
         x = NULL, y = "SBR per 1,000")
}

if (!is.null(nnd01) && !is.null(lb01) && !is.na(country_var)) {
  nnd_by_cty <- df %>%
    mutate(.nnd = nnd01, .lb = lb01) %>%
    group_by(country = .data[[country_var]]) %>%
    summarise(
      NND = sum(.nnd == 1, na.rm = TRUE),
      LB = sum(.lb == 1, na.rm = TRUE),
      NND_rate = rate_1000(NND, LB), .groups = "drop"
    ) %>% filter(LB >= 50) %>% arrange(NND_rate)

  plots_cty$nnd <- ggplot(nnd_by_cty,
    aes(x = reorder(country, NND_rate), y = NND_rate)) +
    geom_segment(
      aes(xend = country, y = 0, yend = NND_rate),
      linewidth = 0.6, color = "#91bfdb"
    ) +
    geom_point(aes(size = LB),
               color = "#4575b4", alpha = 0.8) +
    scale_size_continuous(
      range = c(1.5, 6), labels = comma,
      name = "Live Births"
    ) +
    coord_flip() +
    labs(title = "B. NND Rate by Country",
         x = NULL, y = "NND per 1,000 LB")
}

if (length(plots_cty) == 2) {
  plots_cty$sb + plots_cty$nnd +
    plot_annotation(
      title = "Perinatal Outcome Rates by Country (min 50 births)",
      theme = theme(
        plot.title = element_text(
          face = "bold", size = 15, color = "#1a1a2e"
        )
      )
    )
} else if (length(plots_cty) == 1) {
  plots_cty[[1]]
}
```

## Outcome Summary

```{r outcome-summary-table}
outcome_list <- na.omit(c(V_SB, V_LB, V_NND, V_PND, V_PRE, V_LBW, V_VLBW, V_SGA))
outcome_list <- unique(outcome_list)

if (length(outcome_list) > 0) {
  out_counts <- bind_rows(lapply(outcome_list, function(v) {
    y <- to01(df[[v]])
    tibble(
      Outcome  = v,
      Events   = sum(y == 1, na.rm = TRUE),
      Denominator = sum(!is.na(y)),
      Prevalence  = round(mean(y == 1, na.rm = TRUE) * 100, 2),
      `Rate per 1,000` = rate_1000(sum(y == 1, na.rm = TRUE), sum(!is.na(y)))
    )
  }))

  out_counts %>%
    gt() %>%
    fmt_number(columns = c(Events, Denominator), decimals = 0, use_seps = TRUE) %>%
    fmt_number(columns = Prevalence, decimals = 2, pattern = "{x}%") %>%
    data_color(columns = Prevalence, palette = c("#e8f5e9", "#fff8e1", "#fce4ec"), domain = NULL) %>%
    gt_lshtm(title = "All Outcome Prevalences")
}
```

## Outcomes by Study Type

```{r outcomes-by-type}
if (length(outcome_list) > 0) {
  out_by_type <- bind_rows(lapply(outcome_list, function(v) {
    y <- to01(df[[v]])
    df %>%
      mutate(y_tmp = y) %>%
      group_by(study_type) %>%
      summarise(
        outcome     = v,
        n_available = sum(!is.na(y_tmp)),
        events      = sum(y_tmp == 1, na.rm = TRUE),
        rate_1000   = rate_1000(events, n_available),
        .groups     = "drop"
      )
  }))

  p <- ggplot(out_by_type %>% filter(!is.na(rate_1000)),
              aes(x = study_type, y = rate_1000, fill = outcome,
                  text = paste0(outcome, "\n", study_type, "\nRate: ", rate_1000, " per 1,000"))) +
    geom_col(position = position_dodge(width = 0.8), alpha = 0.85, width = 0.7) +
    scale_fill_brewer(palette = "Set1") +
    coord_flip() +
    labs(title = "Outcome Rates by Study Type",
         x = NULL, y = "Rate per 1,000", fill = "Outcome")

  ggplotly(p, tooltip = "text")
}
```

<hr class="section-divider">

# Birth Characteristics {.tabset .tabset-fade}

## GA & Birthweight Distributions

```{r ga-bw-distributions-composed, fig.height=8, fig.width=14}
plots_dist <- list()

if (!is.na(V_GA)) {
  dga <- tibble(ga = safe_num(df[[V_GA]])) %>%
    filter(!is.na(ga), ga >= 15, ga <= 45)
  if (nrow(dga) > 0) {
    plots_dist$ga <- ggplot(dga, aes(x = ga)) +
      geom_histogram(
        aes(y = after_stat(density)),
        bins = 50, fill = "#2c7fb8", alpha = 0.6,
        color = "white"
      ) +
      geom_density(color = "#c0392b", linewidth = 0.8) +
      geom_vline(
        xintercept = 37, linetype = "dashed",
        color = "#c0392b", linewidth = 0.6
      ) +
      annotate(
        "text", x = 36.5, y = Inf,
        label = "37w (term)", hjust = 1, vjust = 1.5,
        color = "#c0392b", size = 3, fontface = "bold"
      ) +
      labs(
        title = "A. Gestational Age",
        subtitle = paste0(
          "N=", fmt_n(nrow(dga)),
          " | Median=", median(dga$ga), "w"
        ),
        x = "GA (weeks)", y = "Density"
      )
  }
}

if (!is.na(V_BW)) {
  dbw <- tibble(bw = safe_num(df[[V_BW]])) %>%
    filter(!is.na(bw), bw >= 300, bw <= 6000)
  if (nrow(dbw) > 0) {
    plots_dist$bw <- ggplot(dbw, aes(x = bw)) +
      geom_histogram(
        aes(y = after_stat(density)),
        bins = 60, fill = "#41b6c4", alpha = 0.6,
        color = "white"
      ) +
      geom_density(color = "#1a1a2e", linewidth = 0.8) +
      geom_vline(
        xintercept = 2500, linetype = "dashed",
        color = "#c0392b", linewidth = 0.6
      ) +
      annotate(
        "text", x = 2400, y = Inf,
        label = "2,500g (LBW)", hjust = 1, vjust = 1.5,
        color = "#c0392b", size = 3, fontface = "bold"
      ) +
      labs(
        title = "B. Birthweight",
        subtitle = paste0(
          "N=", fmt_n(nrow(dbw)),
          " | Median=", median(dbw$bw), "g"
        ),
        x = "Birthweight (g)", y = "Density"
      )
  }
}

if (length(plots_dist) >= 2) {
  plots_dist$ga + plots_dist$bw +
    plot_annotation(
      title = "Key Birth Characteristic Distributions",
      theme = theme(
        plot.title = element_text(
          face = "bold", size = 15, color = "#1a1a2e"
        )
      )
    )
} else if (length(plots_dist) == 1) {
  plots_dist[[1]]
}
```

## GA & BW by Study (Ridgeline)

```{r ga-bw-ridgeline-composed, fig.height=10, fig.width=14}
plots_ridge <- list()

if (!is.na(V_GA) && !is.na(study_var)) {
  dga_s <- df %>%
    transmute(
      study = .data[[study_var]],
      ga = safe_num(.data[[V_GA]])
    ) %>%
    filter(!is.na(ga), ga >= 15, ga <= 45)
  if (nrow(dga_s) > 0) {
    plots_ridge$ga <- ggplot(
      dga_s,
      aes(x = ga,
          y = fct_reorder(study, ga, .fun = median),
          fill = study)
    ) +
      geom_density_ridges(
        alpha = 0.5, scale = 1.2,
        rel_min_height = 0.01
      ) +
      scale_fill_viridis_d(option = "D") +
      geom_vline(
        xintercept = 37, linetype = "dashed",
        color = "#c0392b", linewidth = 0.4
      ) +
      labs(title = "A. GA by Study",
           x = "GA (weeks)", y = NULL) +
      theme(legend.position = "none")
  }
}

if (!is.na(V_BW) && !is.na(study_var)) {
  dbw_s <- df %>%
    transmute(
      study = .data[[study_var]],
      bw = safe_num(.data[[V_BW]])
    ) %>%
    filter(!is.na(bw), bw >= 300, bw <= 6000)
  if (nrow(dbw_s) > 0) {
    plots_ridge$bw <- ggplot(
      dbw_s,
      aes(x = bw,
          y = fct_reorder(study, bw, .fun = median),
          fill = study)
    ) +
      geom_density_ridges(
        alpha = 0.5, scale = 1.2,
        rel_min_height = 0.01
      ) +
      scale_fill_viridis_d(option = "C") +
      geom_vline(
        xintercept = 2500, linetype = "dashed",
        color = "#c0392b", linewidth = 0.4
      ) +
      labs(title = "B. Birthweight by Study",
           x = "Birthweight (g)", y = NULL) +
      theme(legend.position = "none")
  }
}

if (length(plots_ridge) >= 2) {
  plots_ridge$ga + plots_ridge$bw +
    plot_annotation(
      title = "Birth Characteristics by Study (Ridgeline)",
      theme = theme(
        plot.title = element_text(
          face = "bold", size = 15, color = "#1a1a2e"
        )
      )
    )
} else if (length(plots_ridge) == 1) {
  plots_ridge[[1]]
}
```

## GA vs Birthweight

```{r ga-bw-scatter}
if (!is.na(V_GA) && !is.na(V_BW)) {
  d <- df %>%
    transmute(
      ga = safe_num(.data[[V_GA]]),
      bw = safe_num(.data[[V_BW]]),
      sb = if (!is.na(V_SB)) factor(to01(.data[[V_SB]]), labels = c("Live Birth", "Stillbirth")) else "All"
    ) %>%
    filter(!is.na(ga), !is.na(bw), ga >= 15, ga <= 45, bw >= 300, bw <= 6000)

  if (nrow(d) > 0) {
    # Sample for performance
    set.seed(42)
    d_sample <- if (nrow(d) > 15000) slice_sample(d, n = 15000) else d

    p <- ggplot(d_sample, aes(x = ga, y = bw, color = sb,
                               text = paste0("GA: ", ga, "w\nBW: ", bw, "g\n", sb))) +
      geom_point(alpha = 0.3, size = 1) +
      geom_smooth(method = "loess", se = TRUE, linewidth = 1.2) +
      scale_color_manual(values = c("Live Birth" = "#2c7fb8", "Stillbirth" = "#d73027", "All" = "#2c7fb8")) +
      geom_hline(yintercept = 2500, linetype = "dashed", color = "grey50", linewidth = 0.4) +
      geom_vline(xintercept = 37, linetype = "dashed", color = "grey50", linewidth = 0.4) +
      labs(title = "Gestational Age vs Birthweight",
           subtitle = paste0("Sample: ", fmt_n(nrow(d_sample)), " / ", fmt_n(nrow(d))),
           x = "GA (weeks)", y = "Birthweight (g)", color = "Outcome")

    ggplotly(p, tooltip = "text")
  }
}
```

## Preterm & LBW Categories

```{r preterm-categories}
if (!is.na(V_GA) && !is.na(study_var)) {
  dga_cat <- df %>%
    transmute(study = .data[[study_var]], ga = safe_num(.data[[V_GA]])) %>%
    filter(!is.na(ga), ga >= 15, ga <= 45) %>%
    mutate(
      preterm_cat = factor(case_when(
        ga < 28 ~ "<28w (Extreme preterm)",
        ga < 32 ~ "28-31w (Very preterm)",
        ga < 37 ~ "32-36w (Moderate/Late preterm)",
        TRUE    ~ ">=37w (Term)"
      ), levels = c("<28w (Extreme preterm)", "28-31w (Very preterm)", "32-36w (Moderate/Late preterm)", ">=37w (Term)"))
    )

  tab <- dga_cat %>%
    count(study, preterm_cat, name = "n") %>%
    group_by(study) %>%
    mutate(pct = round(100 * n / sum(n), 1)) %>%
    ungroup()

  p <- ggplot(tab, aes(x = fct_reorder(study, pct, .fun = sum), y = pct, fill = preterm_cat,
                        text = paste0(study, "\n", preterm_cat, "\n", pct, "%"))) +
    geom_col(alpha = 0.9) +
    scale_fill_manual(values = c("#d73027", "#fc8d59", "#fee090", "#91bfdb")) +
    coord_flip() +
    labs(title = "Preterm Categories by Study", x = NULL, y = "%", fill = "Category")

  ggplotly(p, tooltip = "text") %>% layout(margin = list(l = 150))
}
```

<hr class="section-divider">

# Maternal Characteristics {.tabset .tabset-fade}

## Maternal Age

```{r maternal-age-composed, fig.height=6, fig.width=14}
if (!is.na(V_AGE)) {
  dage <- tibble(age = safe_num(df[[V_AGE]])) %>%
    filter(!is.na(age), age >= 10, age <= 60)

  if (nrow(dage) > 0) {
    p_dist <- ggplot(dage, aes(x = age)) +
      geom_histogram(
        aes(y = after_stat(density)),
        bins = 50, fill = "#7fcdbb",
        alpha = 0.6, color = "white"
      ) +
      geom_density(color = "#1a1a2e", linewidth = 0.8) +
      geom_vline(
        xintercept = 35, linetype = "dashed",
        color = "#c0392b", linewidth = 0.5
      ) +
      annotate(
        "text", x = 36, y = Inf,
        label = "AMA (35+)", hjust = 0, vjust = 1.5,
        color = "#c0392b", size = 3, fontface = "bold"
      ) +
      labs(
        title = "A. Age Distribution",
        subtitle = paste0(
          "N=", fmt_n(nrow(dage)),
          " | Median=", median(dage$age), "y"
        ),
        x = "Age (years)", y = "Density"
      )

    dage_cat <- dage %>%
      mutate(age_cat = factor(case_when(
        age < 18 ~ "<18", age < 20 ~ "18-19",
        age < 25 ~ "20-24", age < 30 ~ "25-29",
        age < 35 ~ "30-34", age < 40 ~ "35-39",
        TRUE ~ "40+"
      ), levels = c("<18", "18-19", "20-24",
                    "25-29", "30-34", "35-39", "40+")))

    tab <- dage_cat %>%
      count(age_cat, name = "n") %>%
      mutate(pct = round(100 * n / sum(n), 1))

    p_cat <- ggplot(tab,
      aes(x = age_cat, y = n, fill = age_cat)) +
      geom_col(show.legend = FALSE, alpha = 0.85) +
      geom_text(
        aes(label = paste0(pct, "%")),
        vjust = -0.3, size = 3, color = "#1a1a2e"
      ) +
      scale_fill_brewer(palette = "YlGnBu") +
      scale_y_continuous(
        labels = comma,
        expand = expansion(mult = c(0, 0.1))
      ) +
      labs(title = "B. Age Categories",
           x = "Age Group", y = "N")

    p_dist + p_cat +
      plot_annotation(
        title = "Maternal Age Profile",
        theme = theme(
          plot.title = element_text(
            face = "bold", size = 15,
            color = "#1a1a2e"
          )
        )
      )
  }
}
```

## Education & Sociodemographic

```{r education}
if (!is.na(V_EDU)) {
  edu_tab <- df %>%
    filter(!is.na(.data[[V_EDU]])) %>%
    count(level = .data[[V_EDU]], name = "n") %>%
    mutate(pct = round(100 * n / sum(n), 1)) %>%
    arrange(desc(n))

  p <- ggplot(edu_tab, aes(x = reorder(level, n), y = n, fill = level,
                            text = paste0(level, "\nN = ", fmt_n(n), " (", pct, "%)"))) +
    geom_col(show.legend = FALSE, alpha = 0.85) +
    scale_fill_viridis_d(option = "E") +
    coord_flip() +
    scale_y_continuous(labels = comma) +
    labs(title = "Education Level Distribution", x = NULL, y = "N")

  ggplotly(p, tooltip = "text")
}
```

```{r marital-delivery}
plots <- list()

if (!is.na(V_MAR)) {
  mar_tab <- df %>% filter(!is.na(.data[[V_MAR]])) %>%
    count(level = .data[[V_MAR]], name = "n") %>% mutate(pct = round(100 * n / sum(n), 1))
  plots$marital <- ggplot(mar_tab, aes(x = reorder(level, n), y = n, fill = level)) +
    geom_col(show.legend = FALSE, alpha = 0.85) + scale_fill_viridis_d() +
    coord_flip() + scale_y_continuous(labels = comma) +
    labs(title = "Marital Status", x = NULL, y = "N")
}

if (!is.na(V_MODE)) {
  mode_tab <- df %>% filter(!is.na(.data[[V_MODE]])) %>%
    count(level = .data[[V_MODE]], name = "n") %>% mutate(pct = round(100 * n / sum(n), 1))
  plots$delivery <- ggplot(mode_tab, aes(x = reorder(level, n), y = n, fill = level)) +
    geom_col(show.legend = FALSE, alpha = 0.85) + scale_fill_brewer(palette = "Set2") +
    coord_flip() + scale_y_continuous(labels = comma) +
    labs(title = "Delivery Mode", x = NULL, y = "N")
}

if (length(plots) > 0) {
  if (length(plots) == 2) {
    plots$marital + plots$delivery + plot_layout(ncol = 2)
  } else {
    plots[[1]]
  }
}
```

<hr class="section-divider">

# Cross-tabulations: Outcomes by Risk Factors {.tabset .tabset-fade}

## SB by Maternal Age

```{r sb-by-age}
if (!is.na(V_SB) && !is.na(V_AGE)) {
  d <- tibble(
    sb  = to01(df[[V_SB]]),
    age = safe_num(df[[V_AGE]])
  ) %>%
    filter(!is.na(sb), !is.na(age), age >= 10, age <= 60) %>%
    mutate(age_cat = factor(case_when(
      age < 18 ~ "<18", age < 20 ~ "18-19", age < 25 ~ "20-24",
      age < 30 ~ "25-29", age < 35 ~ "30-34", age < 40 ~ "35-39", TRUE ~ "40+"
    ), levels = c("<18", "18-19", "20-24", "25-29", "30-34", "35-39", "40+")))

  tab <- d %>% group_by(age_cat) %>%
    summarise(N = n(), SB = sum(sb == 1), SBR = rate_1000(SB, N), .groups = "drop")

  p <- ggplot(tab, aes(x = age_cat, y = SBR, fill = SBR,
                        text = paste0(age_cat, "\nSBR: ", SBR, " per 1,000\nN = ", fmt_n(N)))) +
    geom_col(show.legend = FALSE, alpha = 0.85) +
    scale_fill_gradient(low = "#fee090", high = "#d73027") +
    labs(title = "Stillbirth Rate by Maternal Age", x = "Age Group", y = "SBR per 1,000")

  ggplotly(p, tooltip = "text")
}
```

## NND by Birthweight

```{r nnd-by-bw}
if (!is.na(V_NND) && !is.na(V_BW)) {
  d <- tibble(
    nnd = to01(df[[V_NND]]),
    bw  = safe_num(df[[V_BW]])
  ) %>%
    filter(!is.na(nnd), !is.na(bw), bw >= 300, bw <= 6000) %>%
    mutate(bw_cat = factor(case_when(
      bw < 1000 ~ "<1000g", bw < 1500 ~ "1000-1499g",
      bw < 2500 ~ "1500-2499g", bw < 4000 ~ "2500-3999g", TRUE ~ ">=4000g"
    ), levels = c("<1000g", "1000-1499g", "1500-2499g", "2500-3999g", ">=4000g")))

  tab <- d %>% group_by(bw_cat) %>%
    summarise(N = n(), NND = sum(nnd == 1), NND_rate = rate_1000(NND, N), .groups = "drop")

  p <- ggplot(tab, aes(x = bw_cat, y = NND_rate, fill = NND_rate,
                        text = paste0(bw_cat, "\nNND rate: ", NND_rate, " per 1,000\nN = ", fmt_n(N)))) +
    geom_col(show.legend = FALSE, alpha = 0.85) +
    scale_fill_gradient(low = "#fee090", high = "#d73027") +
    labs(title = "Neonatal Death Rate by Birthweight Category", x = "BW Category", y = "NND per 1,000")

  ggplotly(p, tooltip = "text")
}
```

## Outcomes by Education

```{r outcomes-by-edu}
if (!is.na(V_EDU)) {
  outcomes_edu <- na.omit(c(V_SB, V_NND, V_PRE, V_LBW))

  if (length(outcomes_edu) > 0) {
    d <- df %>%
      transmute(edu = as.factor(.data[[V_EDU]]),
                across(all_of(outcomes_edu), ~ to01(.))) %>%
      filter(!is.na(edu))

    edu_prev <- d %>%
      pivot_longer(cols = all_of(outcomes_edu), names_to = "outcome", values_to = "y") %>%
      group_by(edu, outcome) %>%
      summarise(n = sum(!is.na(y)), events = sum(y == 1, na.rm = TRUE),
                rate_1000 = rate_1000(events, n), .groups = "drop")

    p <- ggplot(edu_prev, aes(x = edu, y = rate_1000, fill = outcome,
                               text = paste0(edu, " | ", outcome, "\nRate: ", rate_1000, " per 1,000"))) +
      geom_col(position = position_dodge(width = 0.8), alpha = 0.85, width = 0.7) +
      scale_fill_brewer(palette = "Set1") +
      labs(title = "Outcome Rates by Education Level", x = "Education", y = "Rate per 1,000", fill = "Outcome") +
      theme(axis.text.x = element_text(angle = 30, hjust = 1))

    ggplotly(p, tooltip = "text")
  }
}
```

<hr class="section-divider">

# Table 1: Descriptive Summary {.tabset .tabset-fade}

## By Study Type

```{r table1-by-type}
vars_t1 <- unique(na.omit(c(
  V_AGE, V_EDU, V_MAR, V_MODE, V_MULT,
  V_GA, V_BW, sex_var, V_AP1, V_AP5,
  V_SB, V_LB, V_NND, V_PRE, V_LBW, V_SGA
)))

if (length(vars_t1) > 0) {
  df_t1 <- df %>% select(any_of(c("study_type", vars_t1)))

  cont_v <- intersect(c(V_AGE, V_GA, V_BW, V_AP1, V_AP5, V_SBP, V_DBP), names(df_t1))
  cat_v   <- setdiff(names(df_t1), c("study_type", cont_v))

  if (length(cat_v) > 0) df_t1[cat_v] <- lapply(df_t1[cat_v], as.factor)
  if (length(cont_v) > 0) df_t1[cont_v] <- lapply(df_t1[cont_v], function(x) safe_num(x))

  tbl1 <- df_t1 %>%
    tbl_summary(
      by = study_type,
      statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", all_categorical() ~ "{n} ({p}%)"),
      digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
      missing = "ifany",
      missing_text = "Missing"
    ) %>%
    add_overall(last = TRUE) %>%
    bold_labels() %>%
    modify_header(label ~ "**Characteristic**")

  tbl1 %>%
    as_gt() %>%
    gt_lshtm(title = "Table 1. Descriptive Characteristics by Study Type",
             subtitle = "SSA, 2010 onwards")
}
```

## By Study Source

```{r table1-by-study}
if (!is.na(study_var) && length(vars_t1) > 0) {
  df_t1s <- df %>% select(any_of(c(study_var, vars_t1)))

  cont_v2 <- intersect(c(V_AGE, V_GA, V_BW, V_AP1, V_AP5, V_SBP, V_DBP), names(df_t1s))
  cat_v2   <- setdiff(names(df_t1s), c(study_var, cont_v2))

  if (length(cat_v2) > 0) df_t1s[cat_v2] <- lapply(df_t1s[cat_v2], as.factor)
  if (length(cont_v2) > 0) df_t1s[cont_v2] <- lapply(df_t1s[cont_v2], function(x) safe_num(x))

  tbl1s <- df_t1s %>%
    tbl_summary(
      by = all_of(study_var),
      statistic = list(all_continuous() ~ "{median} ({p25}, {p75})", all_categorical() ~ "{n} ({p}%)"),
      digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
      missing = "ifany",
      missing_text = "Missing"
    ) %>%
    add_overall(last = TRUE) %>%
    bold_labels()

  tbl1s %>%
    as_gt() %>%
    gt_lshtm(title = "Table 1. Descriptive Characteristics by Study Source")
}
```

<hr class="section-divider">

# Table 2: Bivariate Analysis — Covariates vs Outcomes {.tabset .tabset-fade}

## Table 2a: Covariates by Stillbirth

```{r table2-sb}
vars_t2 <- unique(na.omit(c(
  V_AGE, V_EDU, V_MAR, V_MODE, V_MULT,
  V_GA, V_BW, sex_var
)))

if (!is.na(V_SB) && length(vars_t2) > 0) {
  df_t2sb <- df %>%
    select(any_of(c(V_SB, vars_t2))) %>%
    mutate(!!V_SB := factor(to01(.data[[V_SB]]),
                            levels = c(0, 1),
                            labels = c("No SB", "Stillbirth"))) %>%
    filter(!is.na(.data[[V_SB]]))

  cont_t2 <- intersect(c(V_AGE, V_GA, V_BW), names(df_t2sb))
  cat_t2  <- setdiff(names(df_t2sb), c(V_SB, cont_t2))

  if (length(cat_t2) > 0) df_t2sb[cat_t2] <- lapply(df_t2sb[cat_t2], as.factor)
  if (length(cont_t2) > 0) df_t2sb[cont_t2] <- lapply(df_t2sb[cont_t2], function(x) safe_num(x))

  tbl2a <- df_t2sb %>%
    tbl_summary(
      by = all_of(V_SB),
      statistic = list(all_continuous() ~ "{median} ({p25}, {p75})",
                       all_categorical() ~ "{n} ({p}%)"),
      digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
      missing = "ifany",
      missing_text = "Missing"
    ) %>%
    add_overall() %>%
    add_n() %>%
    add_p(test = list(all_continuous() ~ "wilcox.test",
                      all_categorical() ~ "chisq.test")) %>%
    bold_labels() %>%
    bold_p(t = 0.05)

  tbl2a %>%
    as_gt() %>%
    gt_lshtm(title = "Table 2a. Covariates by Stillbirth Status")
}
```

## Table 2b: Covariates by Neonatal Death

```{r table2-nnd}
if (!is.na(V_NND) && length(vars_t2) > 0) {
  df_t2nnd <- df %>%
    select(any_of(c(V_NND, vars_t2))) %>%
    mutate(!!V_NND := factor(to01(.data[[V_NND]]),
                             levels = c(0, 1),
                             labels = c("No NND", "Neonatal Death"))) %>%
    filter(!is.na(.data[[V_NND]]))

  cont_t2n <- intersect(c(V_AGE, V_GA, V_BW), names(df_t2nnd))
  cat_t2n  <- setdiff(names(df_t2nnd), c(V_NND, cont_t2n))

  if (length(cat_t2n) > 0) df_t2nnd[cat_t2n] <- lapply(df_t2nnd[cat_t2n], as.factor)
  if (length(cont_t2n) > 0) df_t2nnd[cont_t2n] <- lapply(df_t2nnd[cont_t2n], function(x) safe_num(x))

  tbl2b <- df_t2nnd %>%
    tbl_summary(
      by = all_of(V_NND),
      statistic = list(all_continuous() ~ "{median} ({p25}, {p75})",
                       all_categorical() ~ "{n} ({p}%)"),
      digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
      missing = "ifany",
      missing_text = "Missing"
    ) %>%
    add_overall() %>%
    add_n() %>%
    add_p(test = list(all_continuous() ~ "wilcox.test",
                      all_categorical() ~ "chisq.test")) %>%
    bold_labels() %>%
    bold_p(t = 0.05)

  tbl2b %>%
    as_gt() %>%
    gt_lshtm(title = "Table 2b. Covariates by Neonatal Death Status")
}
```

<hr class="section-divider">

# Table 3: Covariates by Individual Study {.tabset .tabset-fade}

## Full Table

```{r table3-by-study}
vars_t3 <- unique(na.omit(c(
  V_AGE, V_EDU, V_MAR, V_MODE, V_MULT,
  V_GA, V_BW, sex_var, V_SBP, V_DBP, V_AP1, V_AP5
)))

if (!is.na(study_var) && length(vars_t3) > 0) {
  df_t3 <- df %>% select(any_of(c(study_var, vars_t3)))

  cont_t3 <- intersect(c(V_AGE, V_GA, V_BW, V_AP1, V_AP5, V_SBP, V_DBP), names(df_t3))
  cat_t3  <- setdiff(names(df_t3), c(study_var, cont_t3))

  if (length(cat_t3) > 0) df_t3[cat_t3] <- lapply(df_t3[cat_t3], as.factor)
  if (length(cont_t3) > 0) df_t3[cont_t3] <- lapply(df_t3[cont_t3], function(x) safe_num(x))

  tbl3 <- df_t3 %>%
    tbl_summary(
      by = all_of(study_var),
      statistic = list(all_continuous() ~ "{median} ({p25}, {p75})",
                       all_categorical() ~ "{n} ({p}%)"),
      digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
      missing = "ifany",
      missing_text = "Missing"
    ) %>%
    add_overall(last = TRUE) %>%
    add_n() %>%
    add_p(test = list(all_continuous() ~ "kruskal.test",
                      all_categorical() ~ "chisq.test")) %>%
    bold_labels() %>%
    bold_p(t = 0.05)

  tbl3 %>%
    as_gt() %>%
    gt_lshtm(title = "Table 3. Covariate Profile by Study")
}
```

<hr class="section-divider">

# Covariate Summary by Study {.tabset .tabset-fade}

## Covariate Availability Heatmap

```{r covariate-availability, fig.height=8, fig.width=12}
if (!is.na(study_var)) {
  # Key covariates for modeling
  key_covariates <- unique(na.omit(c(
    V_AGE, V_EDU, V_MAR, V_MODE, V_MULT, V_GA, V_BW,
    sex_var, V_AP1, V_AP5, V_SBP, V_DBP,
    V_SB, V_NND, V_PRE, V_LBW
  )))
  key_covariates <- intersect(key_covariates, names(df))

  if (length(key_covariates) > 0) {
    cov_avail <- df %>%
      group_by(study = .data[[study_var]]) %>%
      summarise(across(all_of(key_covariates), ~ round(100 * mean(!is.na(.)), 1)),
                .groups = "drop")

    cov_long <- cov_avail %>%
      pivot_longer(-study, names_to = "variable", values_to = "pct_available")

    p <- ggplot(cov_long, aes(x = study, y = variable, fill = pct_available,
                               text = paste0(variable, "\n", study, "\n", pct_available, "% available"))) +
      geom_tile(color = "white", linewidth = 0.5) +
      geom_text(aes(label = paste0(round(pct_available), "%")), size = 2.8, color = "grey20") +
      scale_fill_gradientn(
        colours = c("#d73027", "#fc8d59", "#fee090", "#91bfdb", "#4575b4"),
        values = scales::rescale(c(0, 25, 50, 75, 100)),
        limits = c(0, 100), name = "% Available"
      ) +
      labs(title = "Key Covariate Availability by Study",
           subtitle = "Heatmap of data completeness for candidate modeling variables",
           x = NULL, y = NULL) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
            axis.text.y = element_text(size = 9),
            axis.line = element_blank())

    p
  }
}
```

## Continuous Covariates by Study

```{r covariate-stats-by-study}
if (!is.na(study_var)) {
  cont_covs <- intersect(
    unique(na.omit(c(V_AGE, V_GA, V_BW, V_AP1, V_AP5, V_SBP, V_DBP))),
    names(df)
  )

  if (length(cont_covs) > 0) {
    cov_stats <- df %>%
      group_by(study = .data[[study_var]]) %>%
      summarise(
        across(all_of(cont_covs), list(
          median = ~ round(median(safe_num(.), na.rm = TRUE), 1),
          iqr    = ~ round(IQR(safe_num(.), na.rm = TRUE), 1),
          n      = ~ sum(!is.na(.))
        ), .names = "{.col}__{.fn}"),
        .groups = "drop"
      )

    # Reshape for display
    cov_stats_long <- cov_stats %>%
      pivot_longer(-study, names_to = "var_stat", values_to = "value") %>%
      separate(var_stat, into = c("variable", "stat"), sep = "__") %>%
      pivot_wider(names_from = stat, values_from = value) %>%
      mutate(summary = paste0(median, " (IQR: ", iqr, ") [n=", fmt_n(n), "]")) %>%
      select(study, variable, summary) %>%
      pivot_wider(names_from = study, values_from = summary)

    cov_stats_long %>%
      gt() %>%
      cols_label(variable = "Variable") %>%
      gt_lshtm(title = "Continuous Covariate Summary by Study",
               subtitle = "Median (IQR) [n available]")
  }
}
```

<hr class="section-divider">

# Data Completeness & Quality {.tabset .tabset-fade}

## Overall Missingness

```{r missingness-overview}
domains <- tibble(variable = names(df)) %>%
  mutate(domain = case_when(
    str_detect(variable, "^out_") ~ "Outcomes",
    str_detect(variable, "^mat_") ~ "Maternal",
    str_detect(variable, "^neo_") ~ "Neonatal",
    str_detect(variable, "^env_") ~ "Environmental",
    str_detect(variable, "^loc_") ~ "Location",
    str_detect(variable, "^hh_")  ~ "Household",
    TRUE ~ "Other"
  ))

miss_tbl <- tibble(
  variable    = names(df),
  pct_missing = round(100 * colMeans(is.na(df)), 2),
  n_missing   = colSums(is.na(df)),
  n_total     = nrow(df)
) %>%
  left_join(domains, by = "variable") %>%
  arrange(desc(pct_missing))

comp_domain <- miss_tbl %>%
  group_by(domain) %>%
  summarise(
    n_vars           = n(),
    avg_missing_pct  = round(mean(pct_missing), 1),
    median_missing   = round(median(pct_missing), 1),
    vars_gt50_miss   = sum(pct_missing > 50),
    vars_0_miss      = sum(pct_missing == 0),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_missing_pct))

p <- ggplot(comp_domain, aes(x = reorder(domain, avg_missing_pct), y = avg_missing_pct,
                              fill = avg_missing_pct,
                              text = paste0(domain, "\nAvg missing: ", avg_missing_pct, "%\nN vars: ", n_vars))) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  scale_fill_gradient(low = "#c7e9b4", high = "#d73027") +
  coord_flip() +
  labs(title = "Average Missingness by Variable Domain", x = NULL, y = "Avg % Missing")

ggplotly(p, tooltip = "text")
```

```{r missingness-domain-table}
comp_domain %>%
  gt() %>%
  cols_label(
    domain = "Domain", n_vars = "# Variables",
    avg_missing_pct = "Avg Missing %", median_missing = "Median Missing %",
    vars_gt50_miss = ">50% Missing", vars_0_miss = "Complete"
  ) %>%
  data_color(columns = avg_missing_pct, palette = c("#e8f5e9", "#fff8e1", "#fce4ec"), domain = c(0, 100)) %>%
  gt_lshtm(title = "Completeness Summary by Domain")
```

## Completeness Heatmap by Study

```{r completeness-heatmap, fig.height=14, fig.width=16}
if (!is.na(study_var)) {
  admin_exclude <- na.omit(c(study_var, country_var, year_var, site_var, lat_var, lon_var))
  vars_use <- setdiff(names(df), admin_exclude)

  comp_study <- df %>%
    group_by(study = .data[[study_var]]) %>%
    summarise(across(all_of(vars_use), ~ round(100 * mean(!is.na(.)), 1)), .groups = "drop")

  comp_long <- comp_study %>%
    pivot_longer(-study, names_to = "variable", values_to = "pct_complete")

  # Select top varying variables
  var_rank <- comp_long %>%
    group_by(variable) %>%
    summarise(sd_comp = sd(pct_complete, na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(sd_comp))

  top_vars <- var_rank %>% slice_head(n = min(50, nrow(var_rank))) %>% pull(variable)

  hm_mat <- comp_long %>%
    filter(variable %in% top_vars) %>%
    group_by(variable, study) %>%
    summarise(pct_complete = mean(pct_complete, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = study, values_from = pct_complete) %>%
    tibble::column_to_rownames("variable") %>%
    as.matrix()

  pheatmap(
    hm_mat,
    cluster_rows = TRUE, cluster_cols = TRUE,
    main = "Completeness (%) by Study — Top 50 Most Variable",
    fontsize_row = 7, fontsize_col = 9,
    border_color = "white",
    color = colorRampPalette(c("#d73027", "#fc8d59", "#fee090", "#91bfdb", "#4575b4"))(100),
    breaks = seq(0, 100, length.out = 101),
    display_numbers = FALSE
  )
}
```

## Completeness by Study Type

```{r completeness-by-type-heatmap, fig.height=14, fig.width=10}
vars_check <- setdiff(names(df), c(study_var, "study_type", "study_source_std"))

comp_type <- df %>%
  group_by(study_type) %>%
  summarise(across(all_of(vars_check), ~ round(100 * mean(!is.na(.)), 1)), .groups = "drop")

comp_type_long <- comp_type %>%
  pivot_longer(-study_type, names_to = "variable", values_to = "pct_complete")

type_var_rank <- comp_type_long %>%
  group_by(variable) %>%
  summarise(sd_comp = sd(pct_complete, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(sd_comp))

top_vars_type <- type_var_rank %>% slice_head(n = min(60, nrow(type_var_rank))) %>% pull(variable)

hm_type_mat <- comp_type_long %>%
  filter(variable %in% top_vars_type) %>%
  group_by(variable, study_type) %>%
  summarise(pct_complete = mean(pct_complete, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = study_type, values_from = pct_complete) %>%
  tibble::column_to_rownames("variable") %>%
  as.matrix()

pheatmap(
  hm_type_mat,
  cluster_rows = TRUE, cluster_cols = TRUE,
  main = "Completeness (%) by Study Type — Top 60 Most Variable",
  fontsize_row = 7, fontsize_col = 11,
  border_color = "white",
  color = colorRampPalette(c("#d73027", "#fc8d59", "#fee090", "#91bfdb", "#4575b4"))(100),
  breaks = seq(0, 100, length.out = 101)
)
```

## Variable Explorer

```{r variable-explorer}
var_inventory <- tibble(
  Variable = names(df),
  Class    = sapply(df, function(x) paste(class(x), collapse = "/")),
  Type     = sapply(df, function(x) {
    if (is.numeric(x)) "numeric"
    else if (inherits(x, "Date")) "date"
    else "categorical"
  }),
  N_Missing   = colSums(is.na(df)),
  Pct_Missing = round(100 * colMeans(is.na(df)), 1),
  N_Unique    = sapply(df, function(x) n_distinct(na.omit(x))),
  Example     = sapply(df, function(x) paste(head(unique(na.omit(x)), 4), collapse = " | "))
) %>%
  left_join(domains, by = c("Variable" = "variable")) %>%
  arrange(domain, Pct_Missing)

datatable(
  var_inventory,
  filter = "top",
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    dom = "Bfrtip",
    autoWidth = TRUE
  ),
  class = "compact stripe hover",
  caption = htmltools::tags$caption(style = "font-size:16px; font-weight:bold; color:#1a3a5c;",
                                     "Interactive Variable Inventory")
) %>%
  formatStyle("Pct_Missing",
              background = styleColorBar(c(0, 100), "#fc8d59"),
              backgroundSize = "95% 80%",
              backgroundRepeat = "no-repeat",
              backgroundPosition = "center")
```

<hr class="section-divider">

# Pre-Modeling Readiness Assessment {.tabset .tabset-fade}

<div class="info-box">
**Model readiness thresholds:**
Global models require **>= `r params$global_threshold`%** completeness globally.
Type-specific models require **>= `r params$type_threshold`%** in at least one study type.
Study-specific models require **>= `r params$study_threshold`%** in at least one study.
</div>

## Model Readiness Classification

```{r model-readiness}
# Variable availability analysis
if (!is.na(study_var)) {
  avail_study <- comp_long %>%
    group_by(variable) %>%
    summarise(
      max_study  = max(pct_complete, na.rm = TRUE),
      n_studies_gt50 = sum(pct_complete >= 50),
      .groups = "drop"
    )

  avail_type <- comp_type_long %>%
    group_by(variable) %>%
    summarise(
      max_type = max(pct_complete, na.rm = TRUE),
      n_types_gt50 = sum(pct_complete >= 50),
      .groups = "drop"
    )

  global_comp <- tibble(
    variable = names(df),
    pct_global = round(100 * colMeans(is.na(df)), 1)
  ) %>%
    mutate(pct_global = 100 - pct_global)

  model_ready <- global_comp %>%
    left_join(avail_study, by = "variable") %>%
    left_join(avail_type, by = "variable") %>%
    left_join(domains, by = "variable") %>%
    mutate(
      readiness = case_when(
        pct_global >= params$global_threshold ~ "Global model",
        max_type >= params$type_threshold     ~ "Type-specific model",
        max_study >= params$study_threshold   ~ "Study-specific model",
        TRUE ~ "Descriptive only"
      )
    ) %>%
    arrange(desc(readiness == "Global model"),
            desc(readiness == "Type-specific model"),
            desc(pct_global))

  # Summary chart
  ready_summary <- model_ready %>%
    count(readiness, name = "n") %>%
    mutate(pct = round(100 * n / sum(n), 1))

  colors_ready <- c("Global model" = "#43a047", "Type-specific model" = "#1976d2",
                     "Study-specific model" = "#ff9800", "Descriptive only" = "#e53935")

  p <- ggplot(ready_summary, aes(x = reorder(readiness, n), y = n, fill = readiness,
                                  text = paste0(readiness, "\n", n, " variables (", pct, "%)"))) +
    geom_col(show.legend = FALSE, alpha = 0.85) +
    scale_fill_manual(values = colors_ready) +
    coord_flip() +
    labs(title = "Variable Model Readiness Classification",
         subtitle = paste0("Total: ", nrow(model_ready), " variables"),
         x = NULL, y = "Number of Variables")

  ggplotly(p, tooltip = "text")
}
```

```{r model-readiness-table}
if (exists("model_ready")) {
  model_ready %>%
    select(variable, domain, pct_global, max_study, max_type, n_studies_gt50, readiness) %>%
    datatable(
      filter = "top",
      options = list(pageLength = 20, scrollX = TRUE, dom = "Bfrtip"),
      class = "compact stripe hover",
      caption = htmltools::tags$caption(style = "font-size:16px; font-weight:bold; color:#1a3a5c;",
                                         "Model Readiness by Variable")
    ) %>%
    formatStyle("readiness",
                backgroundColor = styleEqual(
                  c("Global model", "Type-specific model", "Study-specific model", "Descriptive only"),
                  c("#e8f5e9", "#e3f2fd", "#fff3e0", "#fce4ec")
                ))
}
```

## Candidate Variables for Modeling

### Completeness by Domain

```{r candidate-vars-domain-grid, fig.height=10, fig.width=12}
if (exists("model_ready")) {
  candidates <- model_ready %>%
    filter(readiness != "Descriptive only") %>%
    mutate(
      tier = factor(readiness, levels = c("Global model", "Type-specific model", "Study-specific model"))
    )

  if (nrow(candidates) > 0) {
    p <- ggplot(candidates, aes(x = reorder(variable, pct_global), y = pct_global, fill = tier)) +
      geom_col(alpha = 0.85) +
      scale_fill_manual(values = c("Global model" = "#43a047",
                                    "Type-specific model" = "#1976d2",
                                    "Study-specific model" = "#ff9800")) +
      coord_flip() +
      facet_wrap(~ domain, scales = "free_y", ncol = 2) +
      geom_hline(yintercept = params$global_threshold, linetype = "dashed", color = "#d73027", linewidth = 0.4) +
      labs(title = "Candidate Variable Completeness by Domain",
           subtitle = "Coloured by readiness tier; dashed line = global threshold",
           x = NULL, y = "% Complete", fill = "Readiness Tier") +
      theme(axis.text.y = element_text(size = 7),
            strip.text = element_text(face = "bold", size = 11))

    p
  }
}
```

### Candidate Variable Table

```{r candidate-vars-table}
if (exists("model_ready") && exists("candidates") && nrow(candidates) > 0) {
  cand_tbl <- candidates %>%
    select(variable, domain, readiness, pct_global, max_type, max_study, n_studies_gt50) %>%
    arrange(readiness, desc(pct_global))

  cand_tbl %>%
    gt(groupname_col = "readiness") %>%
    cols_label(
      variable = "Variable", domain = "Domain",
      pct_global = "Global %", max_type = "Best Type %",
      max_study = "Best Study %", n_studies_gt50 = "Studies >50%"
    ) %>%
    data_color(columns = pct_global, palette = c("#fce4ec", "#e8f5e9"), domain = c(0, 100)) %>%
    data_color(columns = max_type, palette = c("#fce4ec", "#e8f5e9"), domain = c(0, 100)) %>%
    gt_lshtm(title = "Candidate Variables for Predictive Modeling",
             subtitle = "Grouped by readiness tier")
}
```

### Top 50 Candidates (Lollipop)

```{r candidate-lollipop, fig.height=12, fig.width=10}
if (exists("candidates") && nrow(candidates) > 0) {
  top50 <- candidates %>%
    arrange(desc(pct_global)) %>%
    slice_head(n = 50) %>%
    mutate(variable = factor(variable, levels = rev(variable)))

  p <- ggplot(top50, aes(x = variable, y = pct_global, color = tier)) +
    geom_segment(aes(xend = variable, y = 0, yend = pct_global), linewidth = 0.8) +
    geom_point(size = 3, alpha = 0.9) +
    scale_color_manual(values = c("Global model" = "#43a047",
                                   "Type-specific model" = "#1976d2",
                                   "Study-specific model" = "#ff9800")) +
    coord_flip() +
    geom_hline(yintercept = params$global_threshold, linetype = "dashed", color = "#d73027", linewidth = 0.4) +
    labs(title = "Top 50 Candidate Variables for Modeling",
         subtitle = "Ranked by global completeness, coloured by readiness tier",
         x = NULL, y = "% Complete Globally", color = "Readiness Tier") +
    theme(axis.text.y = element_text(size = 8))

  ggplotly(p, tooltip = c("x", "y", "color")) %>% layout(margin = list(l = 200))
}
```

## Global Model Candidates

```{r global-candidates}
if (exists("model_ready")) {
  global_vars <- model_ready %>%
    filter(readiness == "Global model") %>%
    arrange(desc(pct_global))

  if (nrow(global_vars) > 0) {
    p <- ggplot(global_vars %>% slice_head(n = 40),
                aes(x = reorder(variable, pct_global), y = pct_global, fill = domain,
                    text = paste0(variable, "\nCompleteness: ", pct_global, "%\nDomain: ", domain))) +
      geom_col(alpha = 0.85) +
      scale_fill_brewer(palette = "Set2") +
      coord_flip() +
      geom_hline(yintercept = params$global_threshold, linetype = "dashed", color = "#d73027") +
      labs(title = paste0("Variables Ready for Global Models (>= ", params$global_threshold, "% complete)"),
           subtitle = paste0(nrow(global_vars), " variables qualify"),
           x = NULL, y = "% Complete", fill = "Domain")

    ggplotly(p, tooltip = "text") %>% layout(margin = list(l = 200))
  }
}
```

## Sparse But Useful Variables

```{r sparse-useful}
if (exists("model_ready")) {
  sparse_useful <- model_ready %>%
    filter(pct_global < 10, (max_study >= 20 | max_type >= 20)) %>%
    arrange(desc(max_study))

  if (nrow(sparse_useful) > 0) {
    cat('<div class="warning-box">\n')
    cat(paste0("**", nrow(sparse_useful), " variables** have <10% global completeness but >=20% in at least one study/type. ",
               "These are clinically useful for study-specific or type-specific models.\n"))
    cat('</div>\n')

    sparse_useful %>%
      select(variable, domain, pct_global, max_study, max_type, n_studies_gt50) %>%
      gt() %>%
      data_color(columns = max_study, palette = c("#fce4ec", "#e8f5e9"), domain = c(0, 100)) %>%
      gt_lshtm(title = "Sparse But Potentially Useful Variables")
  }
}
```

## Pre-Modeling Recommendations

```{r recommendations, results='asis'}
cat('<div class="success-box">\n')
cat('<h4>Key Recommendations for Modeling</h4>\n')
cat('<ol>\n')

if (exists("global_vars")) {
  cat(paste0('<li><strong>Global model:</strong> ', nrow(global_vars), ' variables with >=', params$global_threshold,
             '% completeness are ready for a pooled multi-study model.</li>\n'))
}

if (exists("model_ready")) {
  type_vars <- model_ready %>% filter(readiness == "Type-specific model")
  study_vars_only <- model_ready %>% filter(readiness == "Study-specific model")
  cat(paste0('<li><strong>Type-specific models:</strong> ', nrow(type_vars),
             ' additional variables become available when modeling within study types (e.g., DHS-only or cohort-only models).</li>\n'))
  cat(paste0('<li><strong>Study-specific models:</strong> ', nrow(study_vars_only),
             ' more variables are usable within individual studies.</li>\n'))
}

cat('<li><strong>Missing data strategy:</strong> Consider multiple imputation (MICE) for variables with 10-40% missingness in the target modeling subset. Variables with >90% global missingness should only be used in study-specific models where they have adequate coverage.</li>\n')
cat('<li><strong>Next steps:</strong> Define outcome (stillbirth, NND, or combined perinatal mortality), select candidate predictors from the global model pool, and evaluate collinearity before building predictive models.</li>\n')
cat('</ol>\n')
cat('</div>\n')
```

<hr class="section-divider">

# Environmental Variables {.tabset .tabset-fade}

## Distribution Overview

```{r env-distributions, fig.height=10, eval=length(env_num) > 0}
if (length(env_num) > 0) {
  env_sample <- df %>% select(all_of(env_num)) %>% slice_sample(n = min(10000, nrow(df)))

  env_long <- env_sample %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
    filter(!is.na(value))

  p <- ggplot(env_long, aes(x = value, fill = variable)) +
    geom_histogram(bins = 40, alpha = 0.7, show.legend = FALSE, color = "white") +
    facet_wrap(~ variable, scales = "free", ncol = 4) +
    scale_fill_viridis_d(option = "H") +
    labs(title = "Environmental Variable Distributions", x = NULL, y = "Count") +
    theme(strip.text = element_text(size = 8))

  p
}
```

## Correlation Matrix

```{r env-correlation, fig.height=10, fig.width=10, eval=length(env_num) > 0}
if (length(env_num) >= 3) {
  env_complete <- df %>%
    select(all_of(env_num)) %>%
    select(where(~ sum(!is.na(.)) > 100))

  if (ncol(env_complete) >= 3) {
    cor_mat <- cor(env_complete, use = "pairwise.complete.obs")

    corrplot(
      cor_mat,
      method = "color",
      type = "upper",
      tl.col = "#1a3a5c",
      tl.cex = 0.7,
      cl.cex = 0.8,
      col = colorRampPalette(c("#d73027", "#fc8d59", "#fee090", "white", "#91bfdb", "#4575b4", "#253494"))(200),
      addCoef.col = "black",
      number.cex = 0.5,
      title = "Environmental Variables Correlation Matrix",
      mar = c(0, 0, 2, 0)
    )
  }
}
```

<hr class="section-divider">

# Session & Export Info

```{r export-excel}
# Export comprehensive Excel workbook
xlsx_path <- file.path(params$output_dir, "LSHTM_Descriptive_Analysis.xlsx")
wb <- createWorkbook()

# Study distribution
if (!is.na(study_var)) {
  addWorksheet(wb, "study_distribution")
  writeData(wb, "study_distribution", df %>% count(.data[[study_var]], study_type, name = "n") %>% arrange(desc(n)))
}

# Study overview
if (exists("study_overview")) {
  addWorksheet(wb, "study_overview")
  writeData(wb, "study_overview", study_overview)
}

# Covariate availability by study
if (exists("cov_avail")) {
  addWorksheet(wb, "covariate_availability")
  writeData(wb, "covariate_availability", cov_avail)
}

# Candidate variables
if (exists("candidates") && nrow(candidates) > 0) {
  addWorksheet(wb, "candidate_variables")
  writeData(wb, "candidate_variables", candidates %>%
    select(variable, domain, readiness, pct_global, max_type, max_study, n_studies_gt50))
}

# Variable inventory
addWorksheet(wb, "variable_inventory")
writeData(wb, "variable_inventory", var_inventory)

# Model readiness
if (exists("model_ready")) {
  addWorksheet(wb, "model_readiness")
  writeData(wb, "model_readiness", model_ready)
}

# Missingness
addWorksheet(wb, "missingness")
writeData(wb, "missingness", miss_tbl)

# Outcome counts
if (exists("out_counts")) {
  addWorksheet(wb, "outcome_counts")
  writeData(wb, "outcome_counts", out_counts)
}

# Bivariate Table 2a — SB
if (exists("tbl2a")) {
  t2a_df <- tbl2a %>% as_tibble()
  addWorksheet(wb, "table2_sb")
  writeData(wb, "table2_sb", t2a_df)
}

# Bivariate Table 2b — NND
if (exists("tbl2b")) {
  t2b_df <- tbl2b %>% as_tibble()
  addWorksheet(wb, "table2_nnd")
  writeData(wb, "table2_nnd", t2b_df)
}

saveWorkbook(wb, xlsx_path, overwrite = TRUE)
```

```{r pptx-export, fig.show='hide'}
# ── Comprehensive PowerPoint with figures ──
pptx_path <- file.path(params$output_dir,
  "LSHTM_Descriptive_Presentation.pptx")
fig_dir <- file.path(params$output_dir, "pptx_figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

# Helper: save ggplot as png and return path
save_fig <- function(p, name, w = 10, h = 6) {
  path <- file.path(fig_dir, paste0(name, ".png"))
  ggsave(path, plot = p, width = w, height = h,
         dpi = 200, bg = "white")
  path
}

ppt <- read_pptx()

# ── Slide 1: Title ──
ppt <- ppt %>%
  add_slide(layout = "Title Slide",
            master = "Office Theme") %>%
  ph_with(
    value = "Predicting Stillbirth & Neonatal Death in SSA",
    location = ph_location_type(type = "ctrTitle")
  ) %>%
  ph_with(
    value = paste0(
      "Descriptive Analysis & Pre-Modeling Report\n",
      "Audencio Victor | LSHTM\n",
      format(Sys.Date(), "%d %B %Y")
    ),
    location = ph_location_type(type = "subTitle")
  )

# ── Slide 2: Executive Summary ──
ppt <- ppt %>%
  add_slide(layout = "Title and Content",
            master = "Office Theme") %>%
  ph_with(value = "Executive Summary",
          location = ph_location_type(type = "title")) %>%
  ph_with(value = paste0(
    "Dataset: ", fmt_n(N_TOTAL), " records\n",
    N_COUNTRIES, " SSA countries | ",
    N_STUDIES, " studies | ", yr_range, "\n",
    N_VARS, " variables\n\n",
    if (!is.na(sbr)) paste0(
      "SBR: ", sbr, " per 1,000 total births\n"
    ) else "",
    if (!is.na(nnd_rate)) paste0(
      "NND: ", nnd_rate, " per 1,000 live births\n"
    ) else "",
    if (!is.na(pmr)) paste0(
      "PMR: ", pmr, " per 1,000 total births\n"
    ) else ""
  ), location = ph_location_type(type = "body"))

# ── Slide 3: Study Overview (N per study) ──
if (exists("study_overview") && !is.na(study_var)) {
  so <- study_overview %>%
    mutate(study = factor(study, levels = study[order(N_records)]))
  p_n_slide <- ggplot(so,
    aes(x = study, y = N_records, fill = Type)) +
    geom_col(alpha = 0.9, width = 0.7) +
    geom_text(
      aes(label = paste0(
        fmt_n(N_records), " (", Pct_total, "%)"
      )),
      hjust = -0.05, size = 3.5, color = "#1a1a2e"
    ) +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(
      labels = comma,
      expand = expansion(mult = c(0, 0.3))
    ) +
    coord_flip() +
    labs(title = "Records per Study",
         x = NULL, y = "N Records", fill = "Type")
  fig3 <- save_fig(p_n_slide, "study_n", w = 10, h = 6)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Study Sample Sizes",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig3, width = 9, height = 5.5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 5.5)
    )
}

# ── Slide 4: Geographic Map ──
if (!is.na(country_var) && exists("ctry") &&
    requireNamespace("rnaturalearth", quietly = TRUE)) {
  africa <- rnaturalearth::ne_countries(
    continent = "Africa", returnclass = "sf"
  )
  africa_map <- africa %>%
    left_join(ctry, by = c("name" = "country")) %>%
    mutate(n = ifelse(is.na(n) | n == 0, NA_real_, n))

  p_map_slide <- ggplot(africa_map) +
    geom_sf(aes(fill = log1p(n)),
            color = "grey70", linewidth = 0.3) +
    scale_fill_viridis_c(
      option = "C",
      name = "Records (log)",
      labels = function(x) fmt_n(round(expm1(x))),
      na.value = "white"
    ) +
    coord_sf(xlim = c(-20, 55), ylim = c(-40, 40)) +
    labs(title = "Geographic Coverage — SSA") +
    theme_void(base_size = 13) +
    theme(
      plot.title = element_text(
        face = "bold", color = "#1a1a2e",
        size = 14, hjust = 0.5
      ),
      legend.position = "bottom",
      legend.key.width = unit(2, "cm")
    )
  fig4 <- save_fig(p_map_slide, "map", w = 8, h = 7)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Geographic Distribution",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig4, width = 7, height = 5.5),
      location = ph_location(left = 1.5, top = 1.5,
                              width = 7, height = 5.5)
    )
}

# ── Slide 5: Outcome Rates by Study ──
if (exists("sb_by_study") && exists("nnd_by_study")) {
  p_sb_s <- ggplot(
    sb_by_study,
    aes(x = reorder(study, SBR), y = SBR, fill = SBR)
  ) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#fee090", high = "#d73027") +
    coord_flip() +
    labs(title = "A. Stillbirth Rate",
         x = NULL, y = "SBR per 1,000")

  p_nnd_s <- ggplot(
    nnd_by_study,
    aes(x = reorder(study, NND_rate), y = NND_rate,
        fill = NND_rate)
  ) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#e0f3f8", high = "#4575b4") +
    coord_flip() +
    labs(title = "B. Neonatal Death Rate",
         x = NULL, y = "NND per 1,000 LB")

  p_outcomes <- p_sb_s + p_nnd_s +
    plot_annotation(
      title = "Perinatal Outcome Rates by Study"
    )
  fig5 <- save_fig(p_outcomes, "outcomes_study",
                    w = 12, h = 6)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Outcome Rates by Study",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig5, width = 9, height = 5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 5)
    )
}

# ── Slide 6: Birth Characteristics ──
if (exists("plots_dist") && length(plots_dist) >= 2) {
  p_birth <- plots_dist$ga + plots_dist$bw +
    plot_annotation(
      title = "Birth Characteristic Distributions"
    )
  fig6 <- save_fig(p_birth, "birth_chars", w = 12, h = 5)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Birth Characteristics",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig6, width = 9, height = 4.5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 4.5)
    )
}

# ── Slide 7: Ridgeline by Study ──
if (exists("plots_ridge") && length(plots_ridge) >= 2) {
  p_ridge <- plots_ridge$ga + plots_ridge$bw +
    plot_annotation(
      title = "GA & Birthweight by Study"
    )
  fig7 <- save_fig(p_ridge, "ridgeline", w = 12, h = 7)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "GA & Birthweight by Study",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig7, width = 9, height = 5.5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 5.5)
    )
}

# ── Slide 8: Model Readiness ──
if (exists("model_ready")) {
  ready_counts <- model_ready %>% count(readiness, name = "n")
  colors_ready <- c(
    "Global model" = "#43a047",
    "Type-specific model" = "#1976d2",
    "Study-specific model" = "#ff9800",
    "Descriptive only" = "#e53935"
  )
  p_ready <- ggplot(
    ready_counts,
    aes(x = reorder(readiness, n), y = n,
        fill = readiness)
  ) +
    geom_col(show.legend = FALSE, alpha = 0.9) +
    geom_text(
      aes(label = paste0(n, " vars")),
      hjust = -0.1, size = 4, color = "#1a1a2e"
    ) +
    scale_fill_manual(values = colors_ready) +
    scale_y_continuous(
      expand = expansion(mult = c(0, 0.2))
    ) +
    coord_flip() +
    labs(title = "Variable Readiness for Modeling",
         subtitle = paste0(
           "Thresholds: Global>=",
           params$global_threshold,
           "% | Type>=", params$type_threshold,
           "% | Study>=", params$study_threshold, "%"
         ),
         x = NULL, y = "N Variables")
  fig8 <- save_fig(p_ready, "readiness", w = 10, h = 5)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Pre-Modeling Readiness",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig8, width = 9, height = 5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 5)
    )
}

# ── Slide 9: Candidate Variables Lollipop ──
if (exists("candidates") && nrow(candidates) > 0) {
  top30 <- candidates %>%
    arrange(desc(pct_global)) %>%
    slice_head(n = 30) %>%
    mutate(
      variable = factor(variable, levels = rev(variable)),
      tier = factor(readiness, levels = c(
        "Global model", "Type-specific model",
        "Study-specific model"
      ))
    )
  p_cand <- ggplot(
    top30,
    aes(x = variable, y = pct_global, color = tier)
  ) +
    geom_segment(
      aes(xend = variable, y = 0, yend = pct_global),
      linewidth = 0.7
    ) +
    geom_point(size = 2.5) +
    scale_color_manual(values = c(
      "Global model" = "#43a047",
      "Type-specific model" = "#1976d2",
      "Study-specific model" = "#ff9800"
    )) +
    coord_flip() +
    labs(title = "Top 30 Candidate Variables",
         x = NULL, y = "% Complete",
         color = "Tier")
  fig9 <- save_fig(p_cand, "candidates", w = 10, h = 8)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content",
              master = "Office Theme") %>%
    ph_with(value = "Candidate Variables for Modeling",
            location = ph_location_type(type = "title")) %>%
    ph_with(
      value = external_img(fig9, width = 9, height = 5.5),
      location = ph_location(left = 0.5, top = 1.5,
                              width = 9, height = 5.5)
    )
}

# ── Slide 10: Covariate Availability Heatmap ──
if (exists("cov_long") && nrow(cov_long) > 0) {
  p_heatmap_slide <- ggplot(cov_long, aes(x = study, y = variable, fill = pct_available)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = paste0(round(pct_available), "%")), size = 2.5, color = "grey20") +
    scale_fill_gradientn(
      colours = c("#d73027", "#fc8d59", "#fee090", "#91bfdb", "#4575b4"),
      values = scales::rescale(c(0, 25, 50, 75, 100)),
      limits = c(0, 100), name = "% Available"
    ) +
    labs(title = "Covariate Availability by Study", x = NULL, y = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
          axis.text.y = element_text(size = 8))
  fig_heat <- save_fig(p_heatmap_slide, "cov_heatmap", w = 12, h = 8)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with(value = "Covariate Availability Heatmap",
            location = ph_location_type(type = "title")) %>%
    ph_with(value = external_img(fig_heat, width = 9, height = 5.5),
            location = ph_location(left = 0.5, top = 1.5, width = 9, height = 5.5))
}

# ── Slide 11: SB by Age + NND by BW ──
sb_age_exists <- !is.na(V_SB) && !is.na(V_AGE)
nnd_bw_exists <- !is.na(V_NND) && !is.na(V_BW)

if (sb_age_exists || nnd_bw_exists) {
  cross_plots <- list()

  if (sb_age_exists) {
    d_sb <- tibble(sb = to01(df[[V_SB]]), age = safe_num(df[[V_AGE]])) %>%
      filter(!is.na(sb), !is.na(age), age >= 10, age <= 60) %>%
      mutate(age_cat = factor(case_when(
        age < 18 ~ "<18", age < 20 ~ "18-19", age < 25 ~ "20-24",
        age < 30 ~ "25-29", age < 35 ~ "30-34", age < 40 ~ "35-39", TRUE ~ "40+"
      ), levels = c("<18", "18-19", "20-24", "25-29", "30-34", "35-39", "40+")))

    tab_sb <- d_sb %>% group_by(age_cat) %>%
      summarise(N = n(), SB = sum(sb == 1), SBR = rate_1000(SB, N), .groups = "drop")

    cross_plots$sb <- ggplot(tab_sb, aes(x = age_cat, y = SBR, fill = SBR)) +
      geom_col(show.legend = FALSE, alpha = 0.85) +
      scale_fill_gradient(low = "#fee090", high = "#d73027") +
      labs(title = "A. Stillbirth Rate by Maternal Age", x = "Age Group", y = "SBR per 1,000")
  }

  if (nnd_bw_exists) {
    d_nnd <- tibble(nnd = to01(df[[V_NND]]), bw = safe_num(df[[V_BW]])) %>%
      filter(!is.na(nnd), !is.na(bw), bw >= 300, bw <= 6000) %>%
      mutate(bw_cat = factor(case_when(
        bw < 1000 ~ "<1000g", bw < 1500 ~ "1000-1499g",
        bw < 2500 ~ "1500-2499g", bw < 4000 ~ "2500-3999g", TRUE ~ ">=4000g"
      ), levels = c("<1000g", "1000-1499g", "1500-2499g", "2500-3999g", ">=4000g")))

    tab_nnd <- d_nnd %>% group_by(bw_cat) %>%
      summarise(N = n(), NND = sum(nnd == 1), NND_rate = rate_1000(NND, N), .groups = "drop")

    cross_plots$nnd <- ggplot(tab_nnd, aes(x = bw_cat, y = NND_rate, fill = NND_rate)) +
      geom_col(show.legend = FALSE, alpha = 0.85) +
      scale_fill_gradient(low = "#e0f3f8", high = "#4575b4") +
      labs(title = "B. NND Rate by Birthweight", x = "BW Category", y = "NND per 1,000 LB")
  }

  if (length(cross_plots) == 2) {
    p_cross <- cross_plots$sb + cross_plots$nnd +
      plot_annotation(title = "Outcome Rates by Risk Factor")
  } else {
    p_cross <- cross_plots[[1]] + plot_annotation(title = "Outcome Rate by Risk Factor")
  }
  fig_cross <- save_fig(p_cross, "cross_tabs", w = 12, h = 5)

  ppt <- ppt %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with(value = "Outcomes by Risk Factor",
            location = ph_location_type(type = "title")) %>%
    ph_with(value = external_img(fig_cross, width = 9, height = 5),
            location = ph_location(left = 0.5, top = 1.5, width = 9, height = 5))
}

# ── Slide 12: Summary / Conclusions ──
ppt <- ppt %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = "Key Findings",
          location = ph_location_type(type = "title")) %>%
  ph_with(value = paste0(
    "Dataset: ", fmt_n(N_TOTAL), " records from ",
    N_COUNTRIES, " SSA countries, ", N_STUDIES, " studies\n\n",
    if (!is.na(sbr)) paste0("Stillbirth rate: ", sbr, " per 1,000 total births\n") else "",
    if (!is.na(nnd_rate)) paste0("Neonatal death rate: ", nnd_rate, " per 1,000 live births\n") else "",
    "\nKey observations:\n",
    "- DHS contributes ~85% of records\n",
    "- Significant heterogeneity across studies\n",
    "- Maternal age, birthweight, and GA strongly associated with outcomes\n",
    "- Missing data patterns vary by study source\n\n",
    "Next steps: variable selection, collinearity assessment, predictive modeling"
  ), location = ph_location_type(type = "body"))

# ── Slide 13: Key Recommendations ──
ppt <- ppt %>%
  add_slide(layout = "Title and Content",
            master = "Office Theme") %>%
  ph_with(value = "Key Recommendations",
          location = ph_location_type(type = "title")) %>%
  ph_with(value = paste0(
    "1. Global model: ", if (exists("global_vars"))
      nrow(global_vars) else "?",
    " variables with >=", params$global_threshold,
    "% completeness\n\n",
    "2. Consider median (IQR) for clinical summaries ",
    "(skewed distributions)\n\n",
    "3. Missing data: MICE for 10-40% missingness; ",
    "study-specific models for sparse variables\n\n",
    "4. Next: define outcome, select predictors, ",
    "evaluate collinearity\n\n",
    "5. DHS dominates (~85%): consider sensitivity ",
    "analyses excluding DHS"
  ), location = ph_location_type(type = "body"))

print(ppt, target = pptx_path)

# Clean up temp figures
unlink(fig_dir, recursive = TRUE)
```

<div class="success-box">
**Outputs exported:**

- Excel workbook: `r file.path(params$output_dir, "LSHTM_Descriptive_Analysis.xlsx")`
- PowerPoint: `r file.path(params$output_dir, "LSHTM_Descriptive_Presentation.pptx")`
</div>

```{r session-info}
cat("R version:", R.version.string, "\n")
cat("Date:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
cat("Platform:", sessionInfo()$platform, "\n\n")
cat("Key packages:\n")
pkgs <- c("dplyr", "ggplot2", "plotly", "gtsummary", "gt", "DT", "pheatmap")
for (p in pkgs) {
  cat(sprintf("  %s: %s\n", p, as.character(packageVersion(p))))
}
```

---

<div style="text-align:center; padding:30px; color:#8e99a4; font-size:0.9em;">
  <strong>LSHTM Perinatal Outcomes Prediction Project</strong><br>
  Audencio Victor | London School of Hygiene & Tropical Medicine<br>
  Report generated on `r format(Sys.Date(), '%d %B %Y')` with R `r R.version.string`
</div>
